---
title: "ggplot Tips"
author: "Sebastian DiGeronimo"
date: '2022-11-11'
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create list of ggplot elements
You can use this to create a base set of options to be used in 
whatever plot you want

Also -
This can be used to select between a few different option if creating
a function. 
```{r}
# ---- ggplot create ggproto ----
# you can make a list of element objects
library("ggplot2")
library("magrittr")

test_gg <- list(geom_point(), theme_bw())
(mtcars %>% 
  ggplot(aes(x = hp, y = mpg, linetype = as.factor(cyl))) +
  geom_line() +
  test_gg[1] +
  labs(title = "Test") + 
  test_gg[2]) %>%
  print()

# useful in a function
test_plt <- function(.x, type) {
  test_gg <- list(geom_point(), theme_bw())
  .x  %>% 
    ggplot(aes(x = hp, y = mpg, linetype = as.factor(cyl))) +
    geom_line() +
    labs(title = "Test function") + 
    test_gg[type]
}
test_plt(mtcars, 1) %>% print()
test_plt(mtcars, 2) %>% print()

rm(test_plt, test_gg)
pacman::p_unload("all")


```

## Get components from a plot
This can be used to see what components exist in your plot.
Then you can select which to extract when plotting multiple things
on top of each other

Seems like you first need to create a plot where everything is on
the same scale, then create other plots where they are on their 
respective scales and are just extracting that single element
```{r}
library("ggplot2")
library("cowplot")
library("patchwork")

dat <- iris
dat$Sepal.Length <- dat$Sepal.Length * 100
dat$Petal.Width <- dat$Petal.Width * 1000

p1 <- ggplot(dat, aes(Sepal.Width, Sepal.Length)) + 
  geom_point() + theme(axis.line = element_line())
p2 <- ggplot(dat, aes(Sepal.Width, Petal.Width)) + geom_point() +
  theme(axis.line = element_line())

p3 <- ggplot(dat, aes(Sepal.Width, Petal.Length)) + 
  geom_point(aes(color = "Petal.Length")) +
  geom_point(aes(y = Sepal.Length/100, color = "Sepal.Length")) +
  geom_point(aes(y = Petal.Width / 1000, color = "Petal.Width")) +
  theme(axis.line = element_line(),
        plot.margin = margin(10, 10, 10, 30))


p1
p2
p3
wrap_elements(get_plot_component(p1, "ylab-l")) +
  wrap_elements(get_y_axis(p1)) +
  wrap_elements(get_plot_component(p2, "ylab-l")) +
  wrap_elements(get_y_axis(p2)) +
  p3 + 
  plot_layout(widths = c(3, 1, 3, 1, 40))

pacman::p_unload("all")
```


## Plot a cirlc around a point
Base Plot and ggplot

Calculate points around a circle from 0 - 360 degrees at 1 degree increments

$$
(x_{coord}, y_{coord}) = \text{location} \\
r = \text{radius,} \\
\theta = \text{degrees,} \\
rads = \theta \times \frac{\pi}{180} \\
x_{component} = x_{coord} + r \times cos(rads) \\
y_{component} = y_{coord} + r \times cos(rads) \\

plot(x_{component}, y_{component}) \text{ for a circle}
$$


```{r, fig.height=15, fig.width=15}
# center location of circle
x_loc <-  data.frame(x = 5, y = 5)

# radius
sd <- 6
bearing <- seq(0, 360, by = 1) * pi/180
circle <- 
  cbind(
    cx = x_loc[,1] + sd * cos(bearing), 
    cy = x_loc[,2] + sd * sin(bearing)
    )

plot(
  x_loc,
  type = "p",
  xlim = range(circle[,1]),
  ylim = range(circle[,2])
  )
lines(circle, col = "red", lwd = 2)


library(ggplot2) 

ggplot() +
  geom_point(
    data  = x_loc, 
    aes(x, y),
    color = "green", 
    size  = 10
    ) +
  geom_path(
    data      = as.data.frame(circle),
    aes(cx, cy),
    color     = "red",
    linewidth = 1
    ) +
  geom_point(
    data  = as.data.frame(circle[seq(1, nrow(circle)) %% 2 == 1, ]),
    aes(cx, cy), 
    color = "orange",
    size  = 4
    ) +
  labs(title = paste0(
    "A circle for you, center = (", 
    x_loc[1,1], 
    ",",
    x_loc[1,2], 
    "), radius = ",
    sd),
    x = NULL,
    y = NULL) +
  theme_classic() +
  theme(title = element_text(size = 30))

pacman::p_unload("all")
```

# Add log ticks to x or y axes
```{r add-log-tick}
library(ggplot2)

ggplot(mtcars, 
       aes(
         x = disp,
         y = mpg
       )
       ) +
  geom_point() +
  labs(title = "Raw",
       x = "Raw disp",
       y = "Raw mpg")

ggplot(mtcars, 
       aes(
         disp, y = log(mpg))
       ) +
  geom_point() +
  labs(title = "Log Transformed Y",
       x = "Raw disp",
       y = "Log mpg")

ggplot(mtcars, 
       aes(x = disp, 
           y = log(mpg))
       ) +
  geom_point() +
  annotation_logticks(side = "l") +
  ylim(0, 5) +
  labs(title = "Log Transformed Y with Log Axis Ticks",
       x = "Raw disp",
       y = "Log mpg")

ggplot(mtcars, 
       aes(x = log(disp), 
           y = log(mpg))
       ) +
  geom_point() +
  annotation_logticks(side = "lb") +
  ylim(0, 5) +
  xlim(0, 7) +
  labs(title = "Log Transformed X and Y with Log Axis Ticks",
       x = "Log disp",
       y = "Log mpg")

pacman::p_unload("all")
```
# Fun: `stat_function` and `geom_stat`
```{r stat-geom-func}
library("ggplot2")
ggplot() +
  xlim(-10, 10) +
  stat_function(fun = dnorm, geom = "point", n = 20, color = "red") +
  geom_function(fun = dnorm, n = 100) +
  geom_function(fun = dnorm, n = 100, args = list(mean = 4, sd = 1), 
                xlim = c(0, 10))

pacman::p_unload("all")
```