---
title: "ggplot Tips"
author: "Sebastian DiGeronimo"
date: '2022-11-11'
output: html_document
---

```{r}
if (!require("librarian")) {
    install.packages("librarian")
}

librarian::shelf(
  quiet = TRUE,
  librarian, ggplot2, magrittr, pacman,
  cowplot, patchwork, ggh4x
  
  )
```



## Create list of ggplot elements
You can use this to create a base set of options to be used in 
whatever plot you want

Also -
This can be used to select between a few different option if creating
a function. 
```{r}
# ---- ggplot create ggproto ---- #
# you can make a list of element objects
library("ggplot2")
library("magrittr")

test_gg <- list(geom_point(), theme_bw())
(mtcars %>% 
  ggplot(aes(x = hp, y = mpg, linetype = as.factor(cyl))) +
  geom_line() +
  test_gg[1] +
  labs(title = "Test") + 
  test_gg[2]) %>%
  print()

# useful in a function
test_plt <- function(.x, type) {
  test_gg <- list(geom_point(), theme_bw())
  .x  %>% 
    ggplot(aes(x = hp, y = mpg, linetype = as.factor(cyl))) +
    geom_line() +
    labs(title = "Test function") + 
    test_gg[type]
}
test_plt(mtcars, 1) %>% print()
test_plt(mtcars, 2) %>% print()

rm(test_plt, test_gg)
pacman::p_unload("all")


```

## Get components from a plot
This can be used to see what components exist in your plot.
Then you can select which to extract when plotting multiple things
on top of each other

Seems like you first need to create a plot where everything is on
the same scale, then create other plots where they are on their 
respective scales and are just extracting that single element
```{r}
library("ggplot2")
library("cowplot")
library("patchwork")

dat <- iris
dat$Sepal.Length <- dat$Sepal.Length * 100
dat$Petal.Width <- dat$Petal.Width * 1000

p1 <- ggplot(dat, aes(Sepal.Width, Sepal.Length)) + 
  geom_point() + theme(axis.line = element_line())
p2 <- ggplot(dat, aes(Sepal.Width, Petal.Width)) + geom_point() +
  theme(axis.line = element_line())

p3 <- ggplot(dat, aes(Sepal.Width, Petal.Length)) + 
  geom_point(aes(color = "Petal.Length")) +
  geom_point(aes(y = Sepal.Length/100, color = "Sepal.Length")) +
  geom_point(aes(y = Petal.Width / 1000, color = "Petal.Width")) +
  theme(axis.line = element_line(),
        plot.margin = margin(10, 10, 10, 30))


p1
p2
p3
wrap_elements(get_plot_component(p1, "ylab-l")) +
  wrap_elements(get_y_axis(p1)) +
  wrap_elements(get_plot_component(p2, "ylab-l")) +
  wrap_elements(get_y_axis(p2)) +
  p3 + 
  plot_layout(widths = c(3, 1, 3, 1, 40))

pacman::p_unload("all")
```


## Plot a circle around a point
Base Plot and ggplot

Calculate points around a circle from 0 - 360 degrees at 1 degree increments

$$
(x_{coord}, y_{coord}) = \text{location} \\
r = \text{radius,} \\
\theta = \text{degrees,} \\
rads = \theta \times \frac{\pi}{180} \\
x_{component} = x_{coord} + r \times cos(rads) \\
y_{component} = y_{coord} + r \times cos(rads) \\

plot(x_{component}, y_{component}) \text{ for a circle}
$$


```{r, fig.height=15, fig.width=15}
# center location of circle
x_loc <-  data.frame(x = 5, y = 5)

# radius
sd <- 6
bearing <- seq(0, 360, by = 1) * pi/180
circle <- 
  cbind(
    cx = x_loc[,1] + sd * cos(bearing), 
    cy = x_loc[,2] + sd * sin(bearing)
    )

plot(
  x_loc,
  type = "p",
  xlim = range(circle[,1]),
  ylim = range(circle[,2])
  )
lines(circle, col = "red", lwd = 2)


library(ggplot2) 

ggplot() +
  geom_point(
    data  = x_loc, 
    aes(x, y),
    color = "green", 
    size  = 10
    ) +
  geom_path(
    data      = as.data.frame(circle),
    aes(cx, cy),
    color     = "red",
    linewidth = 1
    ) +
  geom_point(
    data  = as.data.frame(circle[seq(1, nrow(circle)) %% 2 == 1, ]),
    aes(cx, cy), 
    color = "orange",
    size  = 4
    ) +
  labs(title = paste0(
    "A circle for you, center = (", 
    x_loc[1,1], 
    ",",
    x_loc[1,2], 
    "), radius = ",
    sd),
    x = NULL,
    y = NULL) +
  theme_classic() +
  theme(title = element_text(size = 30))

pacman::p_unload("all")
```

# Add log ticks to x or y axes
```{r add-log-tick}
library(ggplot2)

ggplot(mtcars, 
       aes(
         x = disp,
         y = mpg
       )
       ) +
  geom_point() +
  labs(title = "Raw",
       x = "Raw disp",
       y = "Raw mpg")

ggplot(mtcars, 
       aes(
         disp, y = log(mpg))
       ) +
  geom_point() +
  labs(title = "Log Transformed Y",
       x = "Raw disp",
       y = "Log mpg")

ggplot(mtcars, 
       aes(x = disp, 
           y = log(mpg))
       ) +
  geom_point() +
  annotation_logticks(side = "l") +
  ylim(0, 5) +
  labs(title = "Log Transformed Y with Log Axis Ticks",
       x = "Raw disp",
       y = "Log mpg")

ggplot(mtcars, 
       aes(x = log(disp), 
           y = log(mpg))
       ) +
  geom_point() +
  annotation_logticks(side = "lb") +
  ylim(0, 5) +
  xlim(0, 7) +
  labs(title = "Log Transformed X and Y with Log Axis Ticks",
       x = "Log disp",
       y = "Log mpg")

pacman::p_unload("all")
```


# Fun: `stat_function` and `geom_stat`
```{r stat-geom-func}
library("ggplot2")
ggplot() +
  xlim(-10, 10) +
  stat_function(fun = dnorm, geom = "point", n = 20, color = "red") +
  geom_function(fun = dnorm, n = 100) +
  geom_function(fun = dnorm, n = 100, args = list(mean = 4, sd = 1), 
                xlim = c(0, 10))

pacman::p_unload("all")
```


# Test `earowang/sugrrants` Package

vignette: [sugrrants](https://pkg.earo.me/sugrrants/articles/frame-calendar.html)

Stands for "SUpporting GRaphs with R for ANalysing Time Series"

Uses `ggplot2` and `cowplot` to make graphs in a sort of "calendar" view.

This may be useful when have daily data and want to see how it changes over time.


```{r ex-sugrrants}
librarian::shelf(
  quiet = TRUE,
  earowang/sugrrants,
  dplyr,
  magrittr)

# plot example of `sugrrants` using hourly data
hourly_peds %>%
  filter(Date < as.Date("2016-05-01")) %T>% print() %>% 
  ggplot(aes(x = Time, y = Hourly_Counts, colour = Sensor_Name)) +
  geom_line() +
  facet_calendar(~ Date) + # a variable contains dates
  theme_bw() +
  theme(legend.position = "bottom")


# plot temporal autocorrelation
hourly_peds %>%
  filter(Sensor_ID == 13)  %T>% print() %>%
  ggplot(aes(x = after_stat(lag), y = Hourly_Counts)) +
  geom_acf()

pacman::p_unload("all")

```
# Select the Key Legend Glyph

Since `ggplot2 v3.2.0`

Examples From: https://emilhvitfeldt.com/post/changing-glyph-in-ggplot2/

Use `key_glyph` within a `geom_*` to select the specific legend key glyph

Start with `draw_key_*` to find others

```{r key-lengend-glyphs}
librarian::shelf(
  ggplot2
)
mtcars |>
  tibble::as_tibble(rownames = "car") |>
  ggplot(
    aes(
      x    = as.character(cyl), 
      y    = mpg, 
      fill = as.character(cyl)
      )
         ) +
  geom_boxplot(key_glyph = "rect")
mtcars |>
  tibble::as_tibble(rownames = "car") |>
  ggplot(
    aes(
      x    = as.character(cyl), 
      y    = mpg, 
      fill = as.character(cyl)
      )
         ) +
  geom_boxplot(key_glyph = draw_key_rect)
mtcars |>
  tibble::as_tibble(rownames = "car") |>
  ggplot(
    aes(
      x    = as.character(cyl), 
      y    = mpg, 
      color = as.character(cyl),
      fill = as.character(cyl)
      )
         ) +
  geom_boxplot(key_glyph = draw_key_timeseries)
```

# Add Zebra Striping Pattern on Boxplot

```{r zebra-stripe}
librarian::shelf(magrittr, ggplot2, dplyr, lubridate)
dat <- 
  palmerpenguins::penguins %>%
  
  dplyr::mutate(
    plot_lab = 
      forcats::fct_reorder(
        species, 
        bill_length_mm,
        .desc = TRUE,
        .na_rm = TRUE
    )
  )

# extract every other name to add grey box
rec_factor <-
  dat %>%
  .$species %>%
  levels() %>%
  .[seq(1, length(.), 2)] %>%
  paste0(collapse = "|") %T>% 
  print()

# start plot
dat %>%
  ggplot(
    aes(
      x     = species,
      y     = bill_depth_mm,
      fill  = sex,
      color = after_scale(prismatic::clr_darken(fill, 0.3))
    )
  ) +
  
  # here is where add zebra strips
  geom_tile(
    data =
      ~ .x %>%
      distinct(species, island) %>%
      tidyr::complete(island, species) %>%
      mutate(y = if_else(stringr::str_detect(species, rec_factor), 1, NA)),
    aes(
      x = species,
      y = y,
      height = Inf,
    ),
    fill = "grey",
    alpha = 0.5,
    color = NA,
    na.rm = TRUE
  ) +

  geom_boxplot(
    data = ~ .x %>%
      # distinct(species, island) %>%
      tidyr::complete(island, species, sex, fill = list(bill_depth_mm = -1)),
    width         = 1,
    linewidth     = 0.2,
    outlier.color = "gray70",
    outlier.alpha = 50,
    outlier.size  = 0.05,
    outlier.shape = 20,
    position      = position_dodge2(
      width    = 1,
      preserve = "single",
      padding = 0.5
    ),
    key_glyph    = draw_key_rect,
    na.rm        = TRUE
  ) +
  facet_wrap(~ island, nrow = 3) +
  coord_cartesian(ylim = c(0, NA), expand = FALSE) +
  theme_bw()

# basically need facet column and a separator column
# facet = island
# sep = species
# y == 1 for the grey zebra stripe
dat %>%
  distinct(species, island) %>%
  tidyr::complete(island, species) %>%
  mutate(y = if_else(stringr::str_detect(species, rec_factor), 1, NA))



rm(dat, rec_Factor)

# add stripes to time series
## create tibble with dates covering timeframe
 stripe <-
   expand.grid(2016:2021, paste0("-", c(12, 3, 6, 9), "-01")) %>%
   mutate(
     date = paste0(Var1, Var2),
     date = as_date(date)
   ) %>%
   mutate(season = rep(c("Winter", "Spring", "Summer", "Fall"), each = 6))
 
  ggplot() +
  geom_rect(
    data = stripe,
    aes(xmin = date, xmax = date + months(3), ymin = 0, ymax = Inf,
        fill = season),
    alpha = 0.5
  ) +
  scale_fill_manual(values = c("grey", "white", "grey", "white")) +
  geom_line(
    data =
      paste0("2020-0", c(1:6), "-15") %>%
        as_date() %>%
        tibble(date = .) %>%
        mutate(dat = 1:6),
    aes(x = date, y = dat)
  ) +
  geom_point(
    data =
      paste0("2020-0", c(1:6), "-15") %>%
        as_date() %>%
        tibble(date = .) %>%
        mutate(dat = 1:6),
    aes(x = date, y = dat)
  ) +
  theme_bw()
  
pacman::p_unload("all")
```


# Leaflet Marker 

This shows how to add a marker with multiple rows. Can be modified to include
data with different rows

example here:
<http://www.mysmu.edu/faculty/jwwang/post/how-to-add-multiple-lines-label-on-a-leaflet-map/>

```{r leaflet-marker}
library("leaflet")
long <- c(147.8, 147.6, 147.8, 147., 147.7, 147.76)
lat <- c(-36.852, -36.852, -36.852, -36.852, -36.852, -36.852)
# label <- paste(
#   sep = "<br/>", 
#   "long label1 = ", 
#   "long label2", 
#   "long label3", 
#   "long label4", 
#   "long label5", 
#   "long label6"
#   )

# add placement of external data
x = 1
label <- glue::glue(
  .sep = "<br/>", 
  "long label1 = {x}", 
  "long label2", 
  "long label3", 
  "long label4", 
  "long label5", 
  "long label6"
  )

markers <- data.frame(lat,long, label = as.character(label))


leaflet() %>%
  addTiles() %>%
  addMarkers(
    lng   = markers$long, 
    lat   = markers$lat,
    popup = markers$label,
  )

pacman::p_unload("all")
```

# `sf` Object

```{r create-sf-obj}
librarian::shelf(
  quiet = TRUE,
  librarian,
  ggplot2,
  sf
  )

# if start with data frame or tibble
dat <- data.frame(
  lon = c(-78, -80, -82),
  lat = c(34, 35.5, 35.5),
  val = c(1, 2, 4)
)

# convert data.frame object into sf object 
dat_sf <- st_as_sf(dat, coords = c("lon", "lat"), crs = 4326)
class(dat_sf)

# example shapefile
nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

# plot
ggplot() +
  geom_sf(data = nc, aes(fill = AREA)) +
  
  # ---- if using sf object ---- #
  geom_sf(
    data  = dat_sf, 
    size  = 3, 
    color = "red"
    ) +
  
  # ---- if using data.frame object ---- #
  geom_point(
    data  = dat, 
    aes(x = lon + 0.5, y = lat + 0.5), 
    color = "yellow"
    )


pacman::p_unload("all")
```
# Test `geom_col` with position and `scale_y_reverse`

```{r col-yreverse}
librarian::shelf(
  quiet = TRUE,
  librarian,
  ggplot2, lubridate, tibble, dplyr
  )

test_data <- 
  tibble(
  date    = c(rep("2020-01-01", 20), rep("2020-01-10", 10)), 
  station = rep(c("stn1", "stn2", "stn3"), each = 10), 
  value   = rep(1:10, 3)
) |>
  mutate(
    date = as_date(date)
  )
  
ggplot(test_data, aes(x = date, y = value, fill = station)) +
  geom_col()

ggplot(test_data, aes(x = date, y = value, fill = station)) +
  geom_col(position = position_dodge(preserve = "single")) +
  scale_y_reverse()


```



# `Marmap` Package

Pante and Simon Bouhet - marmap package

Figures 1 and 2

<https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0073051#s4>

Figure 1C - 3D benthic map

```{r marmap}
### LOAD MARMAP
library(marmap)
library(lattice)

# my attempt
fl_benthic <- 
  getNOAA.bathy(lon1 = -85, lon2 = -80, lat1 = 24, lat2 = 29, resolution = 0.5) 

plot(
  fl_benthic,
  image     = TRUE, 
  # bpal      = blues(100), 
  # xlim      = c(141,155),
  deep      = -100, 
  shallow   = 0,
  step      = 25,,
  # col       = c("lightgrey", "darkgrey", "black"), 
  # lwd       = c(0.3, 0.3, 0.6), 
  lty       = c(1, 1, 1),
  drawlabel = c(TRUE, TRUE, TRUE)
)

pacman::p_unload("all")
```


```{r marmap}
### LOAD MARMAP
library(marmap)
library(lattice)

### LOAD DATA
data(nw.atlantic)
data(hawaii)
data(hawaii.sites)
atl <- as.bathy(nw.atlantic) 
papoue <- 
  getNOAA.bathy(lon1 = 140, lon2 = 155, lat1 = -13, lat2 = 0, resolution = 1)

### Figure 1A (top left)
# ---------------------------------------------------------------------------- #

blues <- colorRampPalette(c("purple","lightblue","cadetblue2","cadetblue1","white"))
plot(
  atl,
  image      = TRUE,
  bpal       = blues(100),
  deep       = c(-6500, 0),
  shallow    = c(-50, 0),
  step       = c(500, 0),
  lwd        = c(0.3, 0.6),
  lty        = c(1, 1),
  col        = c("black", "black"),
  drawlabels = c(FALSE, FALSE)
)
scaleBathy(atl, deg = 3, y = 42.5, x = -74)

### Figure 1B (middle left, 2D)
# ---------------------------------------------------------------------------- #

test <- 
  get.transect(
    atl, 
    x1  = -67.9, 
    y1  = 41.2, 
    x2  = -55.2, 
    y2  = 33.6, 
    loc = FALSE, 
    dis = TRUE)
points(
  test$lon, 
  test$lat, 
  type = "l", 
  col  = col2alpha("blue", alpha = 0.5), 
  lwd  = 2, 
  lty  = 1)
points(min(test$lon), max(test$lat), col = "blue")
points(max(test$lon), min(test$lat), col = "blue")
plotProfile(test)


### Figure 1C (middle left, 3D)
# ---------------------------------------------------------------------------- #

out <-
  get.box(
    atl,
    x1    = -67.9,
    y1    = 41.2,
    x2    = -55.2,
    y2    = 33.6,
    width = 2,
    col   = 2
  )
wireframe(
  out,
  shade        = TRUE, 
  zoom         = 1.1,
  aspect       = c(1 / 4, 0.1),
  screen       = list(z = -20, x = -50),
  par.settings = list(axis.line = list(col = "transparent")),
  par.box      = c(col = rgb(0, 0, 0, 0.1))
)

### Figure 1D (bottom left, 3D)
# ---------------------------------------------------------------------------- #

wireframe(
  unclass(atl),
  shade        = TRUE, 
  aspect       = c(1 / 2, 0.1),
  screen       = list(z = 0, x = -50),
  par.settings = list(axis.line = list(col = "transparent")),
  par.box      = c(col = rgb(0, 0, 0, 0.1))
)

### Figure 1E (top right)
# ---------------------------------------------------------------------------- #

blues <- colorRampPalette(c("red", "purple", "blue", "cadetblue1", "white")) 
plot(papoue,
  image     = TRUE, 
  bpal      = blues(100), 
  # xlim      = c(141,155),
  deep      = c(-9000, -3000, 0), 
  shallow   = c(-3000, -10, 0), 
  step      = c(1000, 1000, 0),
  col       = c("lightgrey", "darkgrey", "black"), 
  lwd       = c(0.3, 0.3, 0.6), 
  lty       = c(1, 1, 1),
  drawlabel = c(FALSE, FALSE, FALSE)
)

### Figure 1F (middle right)
# ---------------------------------------------------------------------------- #

sites <- hawaii.sites[-c(1, 4), ]
rownames(sites) <- 1:4

trans1 <- trans.mat(hawaii, min.depth = -1000)
trans2 <- trans.mat(hawaii, min.depth = -4000)

out1 <- lc.dist(trans1, sites, res = "path")
out2 <- lc.dist(trans2, sites, res = "path")

plot(
  hawaii,
  xlim = c(-161, -154), 
  ylim = c(18, 23),
  deep = c(-4000, -1000, 0), 
  shallow = c(-4000, -1000, 0),
  col  = c("grey80", "grey40", "black"), 
  step = c(1, 1, 1),
  lty  = c(1, 1, 1), 
  lwd = c(0.6, 0.6, 1.2), 
  draw = c(FALSE, FALSE, FALSE)
)

points(
  sites, 
  pch = 21, 
  col = "blue", 
  bg  = col2alpha("blue", .9), 
  cex = 1.2)
text(
  sites[, 1], 
  sites[, 2], 
  lab = rownames(sites), 
  pos = c(3, 4, 1, 2), 
  col = "blue")
dummy <- lapply(out2, lines, col = col2alpha("blue", alpha = 0.3), lwd = 1, lty = 1)
dummy <- lapply(out1, lines, col = col2alpha("red", alpha = 0.3), lwd = 1, lty = 1)


### Figure 1G (bottom right)
# ---------------------------------------------------------------------------- #

plot(hawaii, lwd = 0.2)

bathyal <- get.area(hawaii, level.inf = -4000, level.sup = -1000) 
abyssal <- get.area(hawaii, level.inf = min(hawaii), level.sup = -4000)

image(
  bathyal$Lon, 
  bathyal$Lat, 
  bathyal$Area, 
  col = c(rgb(0, 0, 0, 0), rgb(.7, 0, 0, .3)), 
  add = TRUE)
image(
  abyssal$Lon, 
  abyssal$Lat, 
  abyssal$Area, 
  col = c(rgb(0, 0, 0, 0), rgb(.7, .7, 0.3, .3)), 
  add = TRUE)

round(bathyal$Square.Km, 0) -> ba
round(abyssal$Square.Km, 0) -> ab

legend(
  x = "bottomleft", 
  legend = c(paste("bathyal:", ba, "km2"), paste("abyssal:", ab, "km2")),
  bty = "n", 
  fill = c(rgb(.7, 0, 0, .3), rgb(.7, .7, 0, .3))
)

### Figure 2: projected Europe
# ---------------------------------------------------------------------------- #

# bath <- getNOAA.bathy(-20, 20, 30, 80, res = 10, keep = FALSE)
# 
# r1         <- as.raster(bath)
# projection <- "+proj=ortho"
# r2         <- projectRaster(r1, crs = projection)
# bath2      <- as.bathy(r2)
# 
# pal <- colorRampPalette(c("#4F503E", "#3073BB", "#4698CD", "#C6D7D5", "#93856B"))
# 
# # tiff(
# #   filename = "Europe_projected.tiff",
# #   width = 17, height = 17, unit = "cm", res = 650, pointsize = 12,
# #   compression = "lzw", bg = "white", type = c("cairo", "Xlib", "quartz")
# # )
# plot.bathy(
#   bath2, 
#   lwd  = .03, 
#   n    = 1000, 
#   land = TRUE, 
#   col  = pal(1000), 
#   xlab = "", 
#   ylab = "", 
#   axes = FALSE)
# plot.bathy(
#   bath2, 
#   lwd     = .5, 
#   deep    = 0, 
#   shallow = 0, 
#   step    = 0, 
#   col     = pal(1), 
#   add     = TRUE)
# # dev.off()

pacman::p_unload("all")
```



# Adding Multiple Breaks/Labels when Faceting

This helps by allowing specificatino to the breaks or levels for each
facet instead of using `scales = "free_"` or `scales = "free_y"`

Original Code from:
<https://riffomonas.org/code_club/2024-10-31-faceted-boxplots>

Using `ggh4x` to scale each y faceted axes separately as suggested by:
<https://stackoverflow.com/questions/72494323/how-can-i-change-the-number-of-breaks-of-x-axis-for-each-graph-individually-for>

`ggh4x::facetted_pos_scales()`

```{r add multiple-breaks}
library(tidyverse)
library(ggtext)

# ---- original code ---- #
set.seed(19760620)
coral_physiology <- tibble(
  fragment    = 1:72,
  temperature = rep(c(25, 32), each = 36),
  density     = c(
    rnorm(n = 36, mean = 32, sd = 10),
    rnorm(n = 36, mean = 5, sd = 1)
  ),
  photosynthesis = abs(c(
    rnorm(n = 36, mean = 4.5, sd = 1),
    rnorm(n = 36, mean = 2, sd = 1)
  )),
  respiration = abs(c(
    rnorm(n = 36, mean = -2, sd = 0.75),
    rnorm(n = 36, mean = -1, sd = 0.75)
  ))
)

pretty_names <- c(
  "Symbionts (x10<sup>5</sup> cm<sup>-2</sup>)",
  "Photosynthesis<br>(&mu;mol O<sub>2</sub> cm<sup>-2</sup> h<sup>-1</sup>)",
  "Respiration<br>(-&mu;mol O<sub>2</sub> cm<sup>-2</sup> h<sup>-1</sup>)"
)
names(pretty_names) <- c("density", "photosynthesis", "respiration")

sig_table <- tibble(
  name = c("density", "photosynthesis", "respiration"),
  x    = 1.1, xend = 1.9,
  y    = c(55, 8, 4),
  yend = y
)

point_table <- tibble(
  name  = c("density", "photosynthesis", "respiration"),
  x     = 1.5, xend = 1.9,
  y     = c(55, 8, 4) * 1.02,
  label = "*"
)

coral_physiology %>%
  pivot_longer(cols = c(density, photosynthesis, respiration)) %>%
  ggplot(aes(x = as.character(temperature), y = value)) +
  geom_boxplot() +
  geom_segment(
    data = sig_table, 
    aes(x = x, y = y, xend = xend, yend = yend)
    ) +
  geom_text(
    data = point_table, 
    aes(x = x, y = y, label = label),
    size = 5, 
    fontface = "bold"
    ) +
  geom_hline(yintercept = 0) +
  facet_wrap(
    facets         = ~name, 
    scales         = "free_y", 
    nrow           = 3,
    strip.position = "left",
    labeller       = labeller(name = pretty_names)
    ) +
  scale_y_continuous(limits = c(0, NA), expand = c(0,0)) +
  coord_cartesian(clip = "off") +
  labs(x = "Temperature (&deg;C)", y = NULL) +
  theme_classic() +
  theme(
    strip.placement  = "outside",
    strip.background = element_blank(),
    strip.text       = element_markdown(),
    axis.title.x     = element_markdown(),
    panel.spacing    = unit(15, units = "pt")
  )

# ---- No Absolute Values ---- #
set.seed(19760620)

sig_table <- tibble(
  name = c("density", "photosynthesis", "respiration"),
  x = 1.1, xend = 1.9,
  # y = c(55, 8, 4),
  y = c(55, 8, 1),
  yend = y
)

point_table <- tibble(
  name  = c("density", "photosynthesis", "respiration"),
  x     = 1.5, xend = 1.9,
  # y     = c(55, 8, 4) * 1.02,
  y     = c(55, 8, 1) * 1.02,
  label = "*"
)

coral_physiology <- tibble(
  fragment = 1:72,
  temperature = rep(c(25, 32), each = 36),
  density = c(
    rnorm(n = 36, mean = 32, sd = 10),
    rnorm(n = 36, mean = 5, sd = 1)
  ),
  photosynthesis = abs(c(
    rnorm(n = 36, mean = 4.5, sd = 1),
    rnorm(n = 36, mean = 2, sd = 1)
  )),
  
  # remove `abs()` function
  respiration = c(
    rnorm(n = 36, mean = -2, sd = 0.75),
    rnorm(n = 36, mean = -1, sd = 0.75)
  )
)

coral_physiology %>%
  pivot_longer(cols = c(density, photosynthesis, respiration)) %>%
  ggplot(aes(x = as.character(temperature), y = value)) +
  geom_boxplot() +
  geom_segment(data = sig_table, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_text(data = point_table, aes(x = x, y = y, label = label),
            size = 5, fontface = "bold") +
  geom_hline(yintercept = 0) +
  facet_wrap(
    facets         = ~name, 
    scales         = "free_y", 
    nrow           = 3,
    strip.position = "left",
    labeller       = labeller(name = pretty_names)
    ) +
  
  # remove limits arguments
  # scale_y_continuous(limits = c(0, NA), expand = c(0,0)) +
  
  coord_cartesian(clip = "off") +
  labs(x = "Temperature (&deg;C)", y = NULL) +
  theme_classic() +
  theme(
    strip.placement  = "outside",
    strip.background = element_blank(),
    strip.text       = element_markdown(),
    axis.title.x     = element_markdown(),
    panel.spacing    = unit(15, units = "pt")
  )

# ---- No Absolute Values and y-scaled ---- #
library(ggh4x)

coral_physiology %>%
  pivot_longer(cols = c(density, photosynthesis, respiration)) %>%
  ggplot(aes(x = as.character(temperature), y = value)) +
  geom_boxplot() +
  geom_segment(data = sig_table, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_text(data = point_table, aes(x = x, y = y, label = label),
            size = 5, fontface = "bold") +
  geom_hline(yintercept = 0) +
  facet_wrap(
    facets         = ~name, 
    scales         = "free_y", 
    nrow           = 3,
    strip.position = "left",
    labeller       = labeller(name = pretty_names)
    ) +
  
  # remove limits arguments
  # scale_y_continuous(limits = c(0, NA), expand = c(0,0)) +
  
  # add `ggh4x::facetted_pos_scales()` to specify each limit and break
  ggh4x::facetted_pos_scales(
    y = list(
      scale_y_continuous(limits = c(0, NA), expand = c(0, 0)),
      scale_y_continuous(limits = c(0, NA), expand = c(0, 0)),
      scale_y_continuous(limits = c(-3, 3), expand = c(0, 0))
    )
  ) +
  coord_cartesian(clip = "off") +
  labs(x = "Temperature (&deg;C)", y = NULL) +
  theme_classic() +
  theme(
    strip.placement  = "outside",
    strip.background = element_blank(),
    strip.text       = element_markdown(),
    axis.title.x     = element_markdown(),
    panel.spacing    = unit(15, units = "pt")
  )
```


# Add Labels to ggplot2 using a vector instead of adding a new column to dataframe

Jan 08, 2025

Use a vector to change the labels of a plot instead of adding a new
column

first method uses vector for label and facet label:
c("og label" = "new label")

second method uses a two column tibble to create labels
tibble <- tibble(
  "og_label"  = "label1",
  "new_label" = "new_label"
)
plot_label <- tibble$new_label
names(plot_label) <- tibble$og_label

```{r plot-labels-using-vector}
library(ggplot2)
mtcars

car_labs <- 
  c(
    "6" = "6 cylinder",
    "8" = "8 cylinder",
    "4" = "4 cylinder"
  )

# used in facetting with `labeller`
ggplot(mtcars) +
  geom_point(aes(x = mpg, y = hp, color = as.factor(cyl)),
             show.legend = FALSE) +
  facet_grid(~cyl, labeller = labeller(cyl = car_labs)) +
  labs(title = "Used with facet and vector")

# used in labeling 
ggplot(mtcars) +
  geom_point(aes(x = mpg, y = hp, color = as.factor(cyl))) +
  scale_color_discrete(labels = car_labs) +
  labs(
    color = "Cylinder",
    title = "Used as label with vector")



# used in labeling if labels are a tibble instead of vector
car_labs2 <-
tibble::tibble(
  num_cyl = c(
    "8",
    "6",
    "4"
  ),
  car_lab = c(
    "8 cylinder",
    "6 cylinder",
    "4 cylinder"
  )
)

# what you want the labels to be goes first
# then the original names of those labels
cyl_labels <- car_labs2$car_lab 
names(cyl_labels) <- car_labs2$num_cyl

ggplot(mtcars) +
  geom_point(aes(x = mpg, y = hp, color = as.factor(cyl))) +
  scale_color_discrete(labels = cyl_labels) +
  labs(
    color = "Cylinder",
    title = "Used with label and tibble")


# above tip turn into a function 
plot_labels <- function(og_label, new_label) {
  
  label_vec <- new_label
  names(label_vec) <- og_label
  
  return(label_vec)
    # ---- end of function plot_labels
}

cyl_labels2 <- plot_labels(car_labs2$num_cyl, car_labs2$car_lab)

# with var for labels
ggplot(mtcars) +
  geom_point(aes(x = mpg, y = hp, color = as.factor(cyl))) +
  scale_color_discrete(labels = cyl_labels2) +
  labs(
    color = "Cylinder",
    title = "Used with label as tibble, and function and var")

# using function within labels
ggplot(mtcars) +
  geom_point(aes(x = mpg, y = hp, color = as.factor(cyl))) +
  scale_color_discrete(
    labels = plot_labels(car_labs2$num_cyl, car_labs2$car_lab)
    ) +
  labs(
    color = "Cylinder",
    title = "Used with label as tibble and function in place")


pacman::p_unload("all")
```

