---
title: "Belize Mangrove Cover"
author: "Sebastian Di Geronimo"
date: 2025-12-18
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

Extracting mangrove area from `Mapping National Mangrove Cover for Belize Using Google Earth Engine and Sentinel-2 Imagery`
Cissel et al., 2021

Then cropping to bounding box of Palancia


Help from: <https://sarahlistabarth.github.io/blog/posts/opening-kmz-files/>

# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  fs, magrittr, here,
  
  # additional
  sf, units, rstudioapi, archive, rnaturalearth
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select(),
  magrittr::extract()
  )
  

```

# 2.0 Placencia Bounding Box

The bounding box was created from GEE:

var roi = ee.Geometry.Polygon([
  [
    [-88.45002432342531, 16.692982197956695],
    [-88.29913965385717, 16.692982197956695],
    [-88.29913965385717, 16.497023893909766],
    [-88.45002432342531, 16.497023893909766],
    [-88.45002432342531, 16.692982197956695]
  ]
]);

```{r placencia-bbox}
placencia_bbox <-
  tribble(
    ~lon,               ~lat,
    -88.45002432342531, 16.692982197956695,
    -88.29913965385717, 16.692982197956695,
    -88.29913965385717, 16.497023893909766,
    -88.45002432342531, 16.497023893909766,
    -88.45002432342531, 16.692982197956695
  ) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_combine() %>%
  st_cast("POLYGON") %>%
  # convert to projection
  st_transform(crs = "epsg:32616") %T>%
  print()

# bounding box 
bbox <- st_bbox(placencia_bbox)
```

# 3.0 Locating and unzipping `kmz` file

```{r kmz-to-kml}
# path to kmz file
mangrove_kmz <- rstudioapi::selectFile() %T>% 
  print()

# unzip kmz file
archive_extract(
  archive = mangrove_kmz,
  dir = dirname(mangrove_kmz)
  )

# path to unzipped `kml` file
mangrove_kml <- 
  dirname(mangrove_kmz) %>%
  dir_ls(regexp = "doc.kml") %T>% 
  print()
```

## 3.1 Load `kml` file containing whole mangrove polygons for Belize

```{r load-kml-file}
# load kml file
belize_mg_kml <- st_read(mangrove_kml) 

# convert CRS to EPSG of Belize: EPSG:32616
mg_trans <- st_transform(belize_mg_kml, crs = "epsg:32616")

full_mg_bbox <- st_bbox(mg_trans)

# print CRS
cat("\n-----\n\n")
st_crs(belize_mg_kml) # old: WGS84
cat("\n-----\n\n")
st_crs(mg_trans)      # new: projected UTM zone 16N
```

# 4.0 Plot Old and New CRS

```{r plot-area}
# old CRS 
ggplot() +
  geom_sf(data = belize_mg_kml, color = "blue") +
  labs(title = paste("CRS:",st_crs(belize_mg_kml)$input)) +
  theme_bw()

# new CRS
ggplot() +
  geom_sf(data = mg_trans, color = "blue") +
  labs(title = paste("CRS:",st_crs(mg_trans)$input)) +
  theme_bw()
```



# 5.0 Plot Belize Mangrove with Placencia

```{r placencia}
# plot mangrove area on top of placencia
ggplot() +
  geom_sf(
    data = rnaturalearth::ne_countries(returnclass = "sf", scale = "large") %>%
  st_transform(crs = "epsg:32616")
  ) +
  geom_sf(data = mg_trans, fill = "blue", color = "blue") +
  geom_sf(data = placencia_bbox, color = "red", fill = NA) +
  # crop to Belize mangrove
  coord_sf(
    ylim = full_mg_bbox[c(2,4)],
    xlim = full_mg_bbox[c(1,3)],
    )

ggplot() +
  geom_sf(
    data = rnaturalearth::ne_countries(returnclass = "sf", scale = "large") %>%
    st_transform(crs = "epsg:32616")
  ) +
  geom_sf(data = mg_trans, fill = "blue") +
  geom_sf(data = placencia_bbox, color = "red", fill = NA) +
  # crop to Placencia bounding box
  coord_sf(
    ylim = bbox[c(2,4)],
    xlim = bbox[c(1,3)],
    )
```

## 5.1 Crop and Calculate Placencia Area

Area of Placencia: 23.5378682228298 km^2

```{r placencia-crop-area}
# crop to bounding box
placencia_mangrove <- 
  st_crop(mg_trans,  st_bbox(placencia_bbox)) %>%
  mutate(
    area    = st_area(geometry),
    area_km = set_units(area, "km2"),
    area_km = drop_units(area_km)
    ) %T>% 
  print()

# print area in km^2
cat(paste("Area of Placencia:", placencia_mangrove$area_km, "km^2"))

# final area used in area calculation
ggplot() +
  geom_sf(
    data = rnaturalearth::ne_countries(returnclass = "sf", scale = "large") %>%
    st_transform(crs = "epsg:32616")
  ) +
  geom_sf(data = placencia_mangrove, fill = "blue") +
  geom_sf(data = placencia_bbox, color = "red", fill = NA) +
    # crop to Placencia bounding box
  coord_sf(
    ylim = bbox[c(2,4)],
    xlim = bbox[c(1,3)],
    )
```




```{r}
maptiles::get_tiles(belize_mg_kml, zoom = 9)

nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
nc_osm <- get_tiles(nc, crop = TRUE, zoom = 6)
plot_tiles(nc_osm)
```


# 6.0 Calculate Placencia Mangrove Area using GEE Model

Layers trained and modeled by Claudia using GEE
- image based ground truth points
- community ground truth points

```{r file-path}
shelf(terra)
mang_tiff <- 
  here("data", "raw", "placencia") %>%
  dir_ls(regexp = "tif") %T>% 
  print()
```


```{r file-path}
# load mangrove geotiff
mg_load <- 
  terra::rast(mang_tiff[1]) %T>% 
  print()
# names(mg_load) <- c("co", "ib")
mg_load
plot(mg_load)

# calc area of each pixel
mg_cell_size <- 
  cellSize(mg_load, unit = "km", transform = F) 
mg_area <- zonal(mg_cell_size, mg_load, "sum", na.rm = TRUE)

mg_area

# load mangrove geotiff
mg_load <- 
  terra::rast(mang_tiff[2]) %T>% 
  print()
# names(mg_load) <- c("co", "ib")
mg_load
plot(mg_load)

# calc area of each pixel
mg_cell_size <- 
  cellSize(mg_load, unit = "km", transform = F) 
mg_area <- zonal(mg_cell_size, mg_load, "sum", na.rm = TRUE)

mg_area

```


```{r calc-area}
mg_load <- 
  terra::rast(mang_tiff) %>%
  terra::crop(placencia_bbox) %T>% 
  print()
names(mg_load) <- c("co", "ib")
NAflag(mg_load) <- 0

# calc area of each pixel
mg_cell_size <- 
  cellSize(mg_load, unit = "km", transform = F) %>%
  as.data.frame()

as.data.frame(mg_load, na.rm = FALSE) %>%
  bind_cols(mg_cell_size) %>%
   pivot_longer(
     data      = .,
     cols      = c(co, ib),     # columns to pivot long,
     names_to  = "type",  # desired name for category column
     values_to = "class", # desired name for value column
     ) %>%
  summarise(
    .by = c(type, class),
    area = sum(area)
  )

```

```{r}
ggplot() +
  tidyterra::geom_spatraster(
    data = mutate(mg_load, co = as.factor(co), ib = as.factor(ib)) %>%
  terra::crop(placencia_bbox)
    ) +
  geom_sf(data = placencia_bbox, fill = NA, color = "red") +
  scale_fill_discrete(na.value = "transparent") +
  guides(fill = "none") +
  facet_wrap(~lyr) +
  theme_bw()

mutate(mg_load, co = as.factor(co), ib = as.factor(ib)) %>%
  terra::crop(placencia_bbox)
```


# 7.0 Use of public Earth observation data for tracking progress in sustainable management of coastal forest ecosystems in Belize, Central America

Cherrington et al., 2020
<https://www.sciencedirect.com/science/article/pii/S0034425720301681#s0010>

Data Link: 
<https://data.mendeley.com/datasets/2fzbnpnsdp/1>

Calculating area of mangroves change cropped to Placencia

```{r}
cherr_dat <- 
  here("data", "raw", "bz_mangrove_servir_1900_2017") %>%
  dir_ls(regexp = "shp$") %>%
  read_sf() %T>% 
  print()

st_crop(cherr_dat, placencia_bbox) %>%
  plot()


ggplot() +
  geom_sf(
    data = rnaturalearth::ne_countries(returnclass = "sf", scale = "large") %>%
    st_transform(crs = "epsg:32616") %>%
    st_crop(placencia_bbox)
  ) +
  geom_sf(
    data = st_crop(cherr_dat, placencia_bbox),
    aes(fill = TYPE),
    color = "red",
    linewidth = 0.2
  ) +
  facet_wrap(~DESCR, ncol = 6)  +
  theme_bw()

st_crop(cherr_dat, placencia_bbox) %>%
  mutate(
    area_km2 = st_area(geometry),
    area_km2 = set_units(area_km2, "km2"),
    area_ha  = set_units(area_km2, "ha"),
    area_km2 = drop_units(area_km2),
    area_ha  = drop_units(area_ha)
  ) %>%
  arrange(DESCR) %T>% 
  print()
```
18.8604


```{r}

ggplot() +
  geom_sf(
    data = rnaturalearth::ne_countries(returnclass = "sf", scale = "large") %>%
    st_transform(crs = "epsg:32616") %>%
    st_crop(placencia_bbox)
  ) +
  geom_sf(
    data = st_crop(cherr_dat, placencia_bbox)  %>%
    filter(DESCR == "Mangrove 1980-2017"),
    aes(fill = TYPE),
    linewidth = 0.2
  ) +
  facet_wrap(~TYPE, ncol = 6)  +
  theme_bw()
```

