---
title: "Calculate Biomass for Mangrove Area in Placencia, Belize"
author: "Sebastian Di Geronimo"
date: 2026-04-14
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

This document is to run an inverse distance weighted interpolation on a mangrove
biomass data set. 

2 main inputs:
- mangrove tiff file
- mangrove height data
  - location of each measured point
  - average total biomass for each sampled site (above and below ground)
  



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, purrr, dplyr,
  fs, magrittr, here, stringr,
  
  # additional packages, if from GitHub, include username
  sf, terra, geodata, gstat, ggrepel, grid,
  patchwork, tidyterra,
  readxl, janitor, viridisLite, scales
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
```

## 2.2 IDW Defaults and Gaussian Kernel Function

```{r defaults}
# default: local IDW = tighter, less spread, more visually refined
idw_options <-
  list(
    "idp_pow"   = 2.0,  # idw power parameter
    "maxdist_m" = 2000, # max distance from sampled points (in meters)
    "nmin_nb"   = 3,    # min number of data points to use
    "nmax_nb"   = 6    # maximum number of data points
  )

# gaussian kernel for smoothing
# smooths over values with weights
# size - number of cells from center pixel (i.e. 5 = 11 x 11 box)
# sigma - control smoothness
gaussian_kernel = function(size, sigma) {
    x <- seq(-size, size, length.out = 2 * size + 1)
    kernel <- outer(x, x, function(x, y) exp(-(x^2 + y^2) / (2 * sigma^2)))
    return(kernel / sum(kernel)) # Normalize the kernel
}


# region colors and lighten
# region_col <- c("Upper" = "#8566ff", "Lower" = "#ff84ac",
#                 # "Mango Creek" = "#fab474")
#                 "Main Creeks" = "#fab474")
region_col <- c(
  "Upper" = "#c2b3ff", 
  "Lower" = "#ff9ebd", 
  "Main Creeks" = "#f89943")
# region_col <-
#   prismatic::clr_alpha(region_col, 0.75) %>%
#   as.vector() %>%
#   magrittr::set_names(names(region_col)) %T>%
#   print()

# print without degree symbol or N/S/E/W
plain_coord <- function(x) sprintf("%.2f", x)
```


## 2.3 Placencia Bounding Box

The bounding box was created from GEE:

var roi = ee.Geometry.Polygon([
  [
    [-88.45002432342531, 16.692982197956695],
    [-88.29913965385717, 16.692982197956695],
    [-88.29913965385717, 16.497023893909766],
    [-88.45002432342531, 16.497023893909766],
    [-88.45002432342531, 16.692982197956695]
  ]
]);


From Claudia (May 27, 2026)
xmin = -88.440
ymin = 16.5
xmax = -88.303
ymax = 16.685

```{r placencia-bbox}
# projection
wgs_crs     <- 4326 # WGS 84
belize_epsg <- 32616 # for Belize


placencia_bbox <-
#   tribble(
#     ~lon,               ~lat,
#     -88.45002432342531, 16.692982197956695,
#     -88.29913965385717, 16.692982197956695,
#     -88.29913965385717, 16.497023893909766,
#     -88.45002432342531, 16.497023893909766,
#     -88.45002432342531, 16.692982197956695
#   ) %>%
  tribble(
    ~lon,               ~lat,
    -88.440, 16.5,
    -88.440, 16.685,
    -88.303, 16.5,
    -88.303, 16.685
  ) %>%
  st_as_sf(coords = c("lon", "lat"), crs = wgs_crs) %>%
  st_combine() %>%
  st_cast("POLYGON") %>%
  # convert to projection in meters
  st_transform(crs = belize_epsg) %T>%
  print()

# bounding box 
bbox <- st_bbox(placencia_bbox) %T>% 
  print()

# convert to wgs
st_transform(bbox, crs = wgs_crs)

```

## 2.4 Extract Belize Map from `geodata`

```{r belize-map}
# belize map object from geodata then projected
belize_sf <- 
  geodata::gadm(country = "Belize", level = 0, path = tempdir()) %>%
  st_as_sf() %>%
  st_transform(belize_epsg) %>% 

  # crop belize to mangrove bounds
  st_crop(bbox)

belize_sf
plot(belize_sf)

# dark basemap
shelf(maptiles)

belize_tile <- 
  st_transform(bbox, crs = wgs_crs) %>% # convert to WGS
  # get_tiles(provider = "CartoDB.DarkMatterNoLabels", zoom = 14) %>%
  get_tiles(provider = "Esri.WorldImagery", zoom = 14) %>%
  project(paste0("EPSG:", belize_epsg)) # project to Belize UTM

ggplot() +
  geom_spatraster_rgb(data = belize_tile)
```



## 2.5 Path for Mangrove Map and Data

```{r path-data}
# mangrove tiff file
mang_tiff <- 
  here("data", "raw", "placencia") %>%
  dir_ls(regexp = "tif") %>%
  str_subset("~", negate = TRUE) %>% 
  str_subset("2025") %T>% 
  print()

# biomass excel file
biomass_path <- 
  here("data", "raw", "placencia") %>%
  dir_ls(regexp = "Placencia") %>% 
  str_subset("update2") %>% 
  str_subset("~", negate = TRUE) %T>% 
  print()
```



# 3.0 Load Placencia Mangrove tiff

```{r load-tiff}
# load mangrove geotiff
mang_load <- 
  rast(mang_tiff[1]) %T>% 
  print()

mang_load
plot(mang_load)

# name layer
names(mang_load) <- c("co")

# set values of 0 to NA
NAflag(mang_load) <- 0

ggplot() +
  tidyterra::geom_spatraster(
    data = mutate(mang_load, co = as.factor(co))
    ) + 
  guides(fill = "none") +
  
  ggspatial::annotation_north_arrow(
    location = "br", 
    # pad_x = unit(0.4, "in"), 
    pad_y = unit(0.4, "in"),
    style  = ggspatial::north_arrow_orienteering(fill = c("white", "white"))
    ) + 
  ggspatial::annotation_scale(location = "br") +
  
  scale_fill_discrete(na.value = "transparent") +
  theme_bw()

# crop to Placencia bounding box
mang_crop <- 
  mang_load %>%
  crop(bbox)
```



## 3.1 Plot Mangrove on Belize Map

```{r placencia map}
placencia_map <-
  ggplot() +
  geom_sf(data = belize_sf, fill = "grey", alpha = 0.5) +
  tidyterra::geom_spatraster(
    data = mang_crop, 
    aes(fill = co),
    show.legend = FALSE,
    maxcell     = Inf
  ) +
  
  ggspatial::annotation_north_arrow(
    location = "br", 
    # pad_x = unit(0.4, "in"), 
    pad_y = unit(0.4, "in"),
    style  = ggspatial::north_arrow_orienteering(fill = c("white", "white"))
    ) + 
  ggspatial::annotation_scale(location = "br") +
  
  labs(
    x = NULL,
    y = NULL
  ) +
  
  scale_fill_continuous(na.value = "transparent") +
  scale_y_continuous(breaks = seq(16.52, 16.68, 0.02), labels = plain_coord,
                     expand = expansion()) +
  scale_x_continuous(breaks = seq(-88.44, -88.32, 0.02), labels = plain_coord
                     ) +
  
  coord_sf(
    expand = FALSE,
    ylim = bbox[c(2,4)],
    xlim = bbox[c(1,3)],
    label_axes = "ENEN"
  ) +
  theme_bw()

placencia_map
```
## 3.2 Save Mangrove Belize Map

```{r save-base-map}
# set location and name to save plot
plt_save <- 
  here("data", "raw", "placencia",
  paste0("placencia_mangrove_map_", Sys.Date(), ".jpeg")
  ) %T>% 
  print()

if (FALSE) {
  ggsave(
    filename = plt_save,
    plot = placencia_map,
    device   = "jpeg",
    dpi      = 600,
    width    = 8.5,
    height   = 9,
    units    = "in",
    bg       = "white"
  )
} else {
  message("Not saving magrove map!")
}
```


# 4.0 Load Data from the Field

```{r load-sample-data}
# read mangrove biomass dataset
biomass_dat <- 
  readxl::read_xlsx(biomass_path) %>%
  janitor::clean_names() %>%
  select(-x5) %>%
  janitor::remove_empty("cols") %>%
  janitor::remove_empty("rows") %>%
  mutate(
    upper_lower = if_else(str_detect(upper_lower, "Mango"), "Main Creeks", upper_lower)
  ) %T>% 
  print()

# locations as dataframe
biomass_loc <- 
  biomass_dat %>%
  select(
    # "point" = samplingpoint_1,
    station_id,
    "point" = samplingpoint,
    "lat"   = lat,
    # "lat"   = lat_16,
    # "long"  = long_17,
    "long"  = long,
    upper_lower
    ) %>%
  distinct() %>%
  drop_na(lat, long) %>%
  mutate(lat = as.numeric(lat)) %T>% 
  print()

# biomass values
biom_val <- 
  biomass_dat %>%
  # select(samplingpoint_21:last_col()) %>%
  # rename("point" = samplingpoint_21) %>%
  select(sampling_point:total_biomass_per_point) %>%
  rename("point" = sampling_point) %>%
  drop_na() %T>% 
  # right_join(
  #   biomass_sf
  # ) %T>%
  print() 

# location merged with biomass converted to meters projection
biom_df <-
  biomass_loc %>%
  st_as_sf(coords = c("long", "lat"), remove = FALSE, crs = wgs_crs) %>%
  # convert to projection in meters
  st_transform(crs = belize_epsg) %>%
  mutate(
    x = st_coordinates(geometry)[, 1],
    y = st_coordinates(geometry)[, 2],
  ) %>%
  filter(
    between(y, bbox[2], bbox[4]),
    between(x, bbox[1], bbox[3]),
  ) %>%
  left_join(
    biom_val
  ) %>%
  as.data.frame() %T>%
  print() %>%
  left_join(
    tribble(
      ~upper_lower, ~density_per_0_1_ha,
      "Upper", 470,
      "Lower", 733,
      "Main Creeks", 189
      # "Mango Creek", 189
    ) %>%
      mutate(density_100_sq_m = density_per_0_1_ha / 10)
  )  %>%
  mutate(
    total_biomass_per_100_sq_m = total_biomass_per_point * density_100_sq_m
  ) %T>%
  print()


if (FALSE) {
  biom_df %>%
  write.csv(here("data", "raw", "placencia", "placencia_biomass_density.csv"))
}
```
```{r}
biomass_select <- 
  biomass_dat %>%
  filter(str_detect(upper_lower,  "Main Creeks|Upper|Lower")) %>%
  select(
    samplingpoint, 
    station_id, 
    quarternumber, 
    species, 
    morphotype,
    upper_lower,
    east_west,
    location,
    dbh_cm
    ) %>%
  mutate(
    across(morphotype:location, ~ as.factor(.x)),
    upper_lower = factor(upper_lower, levels = c("Upper", "Lower", "Main Creeks"))
         ) %T>% 
  print()

biomass_select %>%
  # filter(!is.na(east_west)) %>%
  WRS2::t1way(
  data    = .,
  dbh_cm  ~ morphotype,
  # dep     = "dbh_cm",       # dependent variable (numeric)
  # factors = c("morphotype"), # independent variables (factors)
  # method  = "trim",                  # "trim" (default), "median", or "boot"
  # ph      = TRUE,                    # post-hoc tests
  tr      = 0.2                      # trim proportion (20%)
  )

biomass_select %>%
  filter(!is.na(east_west)) %>%
  walrus::ranova(
  data    = .,
  dep     = "dbh_cm",       # dependent variable (numeric)
  factors = c("morphotype", "east_west"), # independent variables (factors)
  method  = "trim",                  # "trim" (default), "median", or "boot"
  ph      = TRUE,                    # post-hoc tests
  tr      = 0.2                      # trim proportion (20%)
  )

biomass_select %>%
  walrus::ranova(
  data    = .,
  dep     = "dbh_cm",       # dependent variable (numeric)
  factors = c("morphotype", "upper_lower"), # independent variables (factors)
  method  = "trim",                  # "trim" (default), "median", or "boot"
  ph      = TRUE,                    # post-hoc tests
  tr      = 0.2                      # trim proportion (20%)
  )

morph_loc <- biomass_select %>%
  walrus::ranova(
  data    = .,
  dep     = "dbh_cm",       # dependent variable (numeric)
  factors = c("morphotype", "location"), # independent variables (factors)
  method  = "trim",                  # "trim" (default), "median", or "boot"
  ph      = TRUE,                    # post-hoc tests
  tr      = 0.2                      # trim proportion (20%)
  )

morph_loc$phs

WRS2::t2way(dbh_cm ~ morphotype*upper_lower, data = biomass_select)
WRS2::mcp2atm(dbh_cm ~ morphotype*upper_lower, data = biomass_select)


# set location and name to save plot
box_save <- 
  here("data", "raw", "placencia",
  paste0("placencia_box_", c("upper_lower_", "east_west_", "loc_"), Sys.Date(), ".jpeg")
  ) %T>% 
  print()
shell.exec(dirname(box_save)[1])

biomass_select %>%
  ggplot(aes(x = morphotype, y = dbh_cm)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, 
           color = "black", fill = NA, linewidth = rel(1.1)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.5) +
  facet_wrap(~upper_lower) +
  labs(
    x = "Morphotype",
    y = "dbh (cm)"
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(size = rel(1.5)),
    axis.text = element_text(size = rel(1.5)),
    axis.line = element_blank(),
    strip.text = element_text(size = rel(2))
  )

if (FALSE) {
  ggsave(
    filename = box_save[1],
    plot     = last_plot(),
    device   = "jpeg",
    dpi      = 600,
    width    = 9,
    height   = 6,
    units    = "in",
    bg       = "white"
  )
}

biomass_select %>%
  filter(!is.na(east_west)) %>%
  ggplot(aes(x = morphotype, y = dbh_cm)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, 
           color = "black", fill = NA, linewidth = rel(1.1)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.5) +
  facet_wrap(~east_west) +
  labs(
    x = "Morphotype",
    y = "dbh (cm)"
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(size = rel(1.5)),
    axis.text = element_text(size = rel(1.5)),
    axis.line = element_blank(),
    strip.text = element_text(size = rel(2))
  )

if (FALSE) {
  ggsave(
      filename = box_save[2],
      plot = last_plot(),
      device   = "jpeg",
      dpi      = 600,
      width    = 9,
      height   = 6,
      units    = "in",
      bg       = "white"
    )
}

biomass_select %>%
  ggplot(aes(x = morphotype, y = dbh_cm)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, 
           color = "black", fill = NA, linewidth = rel(1.1)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.5) +
  facet_wrap(~location) +
  labs(
    x = "Morphotype",
    y = "dbh (cm)"
  ) +
  theme_classic() +
  theme(
    axis.title = element_text(size = rel(1.5)),
    axis.text = element_text(size = rel(1.5)),
    axis.line = element_blank(),
    strip.text = element_text(size = rel(2))
  )

if (FALSE) {
  ggsave(
      filename = box_save[3],
      plot = last_plot(),
      device   = "jpeg",
      dpi      = 600,
      width    = 9,
      height   = 6,
      units    = "in",
      bg       = "white"
    )
}


```

```{r}
# re_station <- 
#   biomass_dat %>%
#   select(
#     # "point" = samplingpoint_1,
#     "point" = samplingpoint,
#     "lat"   = lat,
#     # "lat"   = lat_16,
#     # "long"  = long_17,
#     "long"  = long,
#     upper_lower
#     ) %>%
#   filter(!is.na(lat) & !is.na(long)) %>%
#   # distinct(upper_lower) %$% upper_lower %>%
#   # cat(sep = "', '")
#   mutate(
#     .by = point,
#     up_low = forcats::fct(
#       upper_lower, 
#       levels = c('Upper', 'Lower', 'Main Creeks', #'Mango Creek', 
#                  'Laughing Bird Caye', 
#                  'Baker\'s Caye', 'Bugle Caye', 'Little Bugle Caye', 
#                  'Scipio Caye', 'Jeremy Caye', 'Dolphin Caye', 'Whipray Caye', 
#                  'Loggerhead Caye', 'Saddle Caye', 'Abigail Caye')
#       )
#   ) %>%
#   arrange(up_low, desc(lat)) %>%
#   mutate(station_id = row_number(), .before = point) %T>% 
#   print()
#   
# if (FALSE) {
#   
#   biomass_dat %>%
#     left_join(select(re_station, station_id, "samplingpoint" = point)) %>%
#     relocate(station_id, .after = 1) %>%
#     select(1,2, upper_lower) %>%
#     writexl::write_xlsx(here(dirname(biomass_path), "station_id.xlsx"))
# }
# 
# biom_df <- 
#   biom_df %>%
#   left_join(select(re_station, station_id, point)) %>%
#   relocate(station_id, .after = 1)
# 
biomass_dat %>%
  select(1,5,aboveground_biomass_kg_per_tree, belowground_biomass_kg_per_tree, total_biomass_per_tree) %T>% 
  print() %>%
  summarise(
    .by = c(samplingpoint),
    across(c(aboveground_biomass_kg_per_tree:total_biomass_per_tree),
           list(avg = ~mean(.x, na.rm = TRUE)))
  ) %>%
  left_join(biom_val, by = c( "samplingpoint" = "point")) %>%
  filter(!is.na(average_agb_per_point)) %>%
  mutate(
    across(aboveground_biomass_kg_per_tree_avg:last_col(),
           ~ round(.x, 6))
  ) %>%
  mutate(
    all_equal = aboveground_biomass_kg_per_tree_avg == average_agb_per_point 
    & belowground_biomass_kg_per_tree_avg == average_bgb_per_point 
    & total_biomass_per_tree_avg == total_biomass_per_point,
    
    aba_equal = aboveground_biomass_kg_per_tree_avg == average_agb_per_point,
    bgb_eg    =  belowground_biomass_kg_per_tree_avg == average_bgb_per_point, 
    tot_eq    = total_biomass_per_tree_avg == total_biomass_per_point
  )


```

## 4.1 Set Value Breaks from Histogram
```{r breaks}
hist(
  # biom_df$total_biomass_kg, 
  biom_df$total_biomass_per_point, 
  breaks = 50, 
  main   = paste("Histogram of Total Biomass (kg)"),
  xlab   = "Total Biomass (kg)"
  )

hist(
  biom_df$total_biomass_per_100_sq_m, 
  breaks = 100, 
  main   = paste("Histogram of Total Biomass per 100 sq. m (kg 100 m^-2)"),
  xlab   = "Total Biomass per 100 sq. m (kg 100 m^-2)"
  )
axis(side = 1, at = seq(0, 11000, 250))

plot(ecdf(biom_df$total_biomass_per_100_sq_m))

# class int
brk_num <- 15 # edit number of breaks

# classInt::classIntervals(biom_df$total_biomass_per_100_sq_m, 15, style = "bclust") %T>% 
#   print() %>%
#   plot(pal = viridisLite::viridis(15), main = "ECDF Breaks 15")

breaks_8 <- 
  classInt::classIntervals(biom_df$total_biomass_per_100_sq_m, brk_num, style = "quantile") %T>% 
  print() 

plot(breaks_8, pal = viridisLite::viridis(brk_num), main = "ECDF Breaks 8")

break_df <- data.frame(
  xmin = head(breaks_8$brks, -1),
  xmax = tail(breaks_8$brks, -1),
  y = 0,
  cls = factor(seq_len(length(breaks_8$brks) - 1))
)

ecdf_df <- biom_df %>%
  select("x" = total_biomass_per_100_sq_m) %>%
  arrange(x) %>%
  mutate(
    .keep = "used",
    ecdf = seq_along(x) / length(x)
    )

ggplot() +
  geom_rect(
    data = break_df,
    aes(xmin = xmin, xmax = xmax, ymin = -0.06, ymax = -0.02, fill = cls),
    color = "grey40"
  ) +
  geom_step(data = ecdf_df, aes(x = x, y = ecdf), color = "black") +
  geom_vline(xintercept = breaks_8$brks, color = "grey", linewidth = 0.4,
             linetype = "dashed") +
  labs(
    title = paste("ECDF with", brk_num, "Breaks"),
    x = NA,
    y = "F(n)"
  ) +
  scale_fill_manual(values = viridisLite::viridis(brk_num), guide = "none") +
  scale_x_continuous(breaks = scales::breaks_width(2000), expand = expansion()) +
  scale_y_continuous(expand = expansion()) +
  theme_classic() +
  theme(
    axis.title.x       = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title.y       = element_text(angle = 90),
    axis.text.y        = element_text(angle = 90),
  )


breaks_8$brks
breaks_8 <- c(0, round(breaks_8$brks / 50) * 50, 11000)
breaks_8 <- c(0, 50, 200, 400, 800, 1600, 2400, 2400, 5700, 11000)

# rm(break_df, ecdf_df)
```

## 4.2 Plot Data Point on Placencia Map

```{r plt-points-map}
placencia_map +
  # ggspatial::geom_spatial_rect(
  #   aes(
  #     xmin = bbox[1],
  #     ymin = bbox[2],
  #     xmax = bbox[3],
  #     ymax = bbox[4],
  #     ),
  #   color = "red",
  #   fill = NA,
  #   crs = belize_epsg,
  #   ) +
  geom_sf(
    data = biom_df,
    aes(geometry = geometry, color = upper_lower),
    alpha = 0.75,
    size  = rel(2),
    # color = "red4",
    shape = 16,
    # show.legend = FALSE
  ) +
  scale_color_manual(
    name = "Region",
    # values = c("Upper" = "#8566ff", "Lower" = "#ff84ac", "Mango Creek" = "#fab474")
    values = region_col,
    breaks = names(region_col)
    ) +
  ggrepel::geom_text_repel(
    data = biom_df,
    # aes(x = x, y = y, label = point, color = upper_lower),
    aes(x = x, y = y, label = station_id, color = upper_lower),
    min.segment.length = 0,
    max.overlaps       = 100,
    # color              = "white",
    bg.color           = "black",
    segment.color      = "black", 
    alpha              = 0.75,
    show.legend = FALSE
  ) +
  coord_sf(
    expand = FALSE,
    ylim = bbox[c(2, 4)],
    xlim = bbox[c(1, 3)],
  )

```


## 4.3 Save Points on Map

```{r save-points-map}
# set location and name to save plot
plt_save2 <- 
  here("data", "raw", "placencia",
  paste0("placencia_mangrove_map_w_points_update2_", Sys.Date(), ".jpeg")
  # paste0("placencia_mangrove_map_w_points_all_", Sys.Date(), ".jpeg")
  ) %T>% 
  print()

if (FALSE) {
  ggsave(
    filename = plt_save2,
    plot = last_plot(),
    device   = "jpeg",
    dpi      = 600,
    width    = 8.5,
    height   = 9,
    units    = "in",
    bg       = "white"
  )
} else {
  message("Not saving mangrove map with points!")
}
```


# 6.0 IDW Interpolation

```{r idw-interp}
idw_sf_utm <-
  gstat::gstat(
    # formula   = total_biomass_kg ~ 1,
    formula   = total_biomass_per_100_sq_m ~ 1,
    data      = as.data.frame(biom_df),
    locations = ~ x + y,
    set       = list(idp = idw_options$idp_pow),
    maxdist   = idw_options$maxdist_m,
    nmin      = idw_options$nmin_nb,
    nmax      = idw_options$nmax_nb
  )

# create a blank grid based on spatial extent of mangrove cropped area
# edit `res` to desired resolution in meters
# rast_test <- rast(ext(mang_crop), res = 1000, crs = paste0("EPSG:", belize_epsg))
rast_test <- rast(ext(mang_crop), res = 10, crs = paste0("EPSG:", belize_epsg))

idw_rast_smooth <- 
  # run model on grid
  terra::interpolate(rast_test, idw_sf_utm) %>%
  
  # gaussian smooth
  terra::focal(
    w         = gaussian_kernel(size = 10, sigma = 5),
    fun       = mean,
    na.policy = "omit",
    na.rm     = TRUE,
    fillvalue = NA
  ) %>%
  
  # mask to valid mangrove points
  terra::mask(
    terra::vect(
      as.data.frame(mang_crop, xy = TRUE), 
      crs = paste0("EPSG:", belize_epsg)
      )
    )
```


## 6.1 Plot IDW Interpolation onto Map

```{r plot-idw}
ggplot() +
  geom_sf(data = belize_sf, fill = "grey", alpha = 0.5) + # gray basemap
  # geom_spatraster_rgb(data = belize_tile) + # dark base map
  
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA) +
  
  # add interpolated map
  tidyterra::geom_spatraster(
    data = idw_rast_smooth,
    aes(fill = var1.pred),
    na.rm   = TRUE,
    maxcell = Inf
  ) +
  
  # add sample points
  geom_sf(
    data = biom_df,
    # aes(fill = total_biomass_kg, geometry = geometry),
    # aes(fill = total_biomass_per_100_sq_m, geometry = geometry),
    aes(fill = total_biomass_per_100_sq_m, geometry = geometry, color = upper_lower),
    alpha  = 0.75,
    size   = rel(1.5),
    stroke = rel(0.75),
    # color  = "white",
    shape  = 21,
    key_glyph = draw_key_rect
  ) +
  
  ggrepel::geom_text_repel(
    data = biom_df,
    aes(x = x, y = y, label = station_id, color = upper_lower),
    min.segment.length = 0,
    max.overlaps       = 100,
    # color              = "white",
    bg.color           = "black",
    segment.color      = "black",
    # bg.color           = "white",
    bg.r = 0.05,
    # segment.color      = "white",
    alpha              = 0.75,
    show.legend = FALSE
  ) +
  
  # region colors
  scale_color_manual(
    name   = "Region",
    values = region_col,
    breaks = names(region_col)
    ) +
  
  scale_fill_stepsn(
    # name     = "Total Biomass (kg)",
    name     = expression("Total Biomass (kg"~100~m^-2*")"),
    colours  = viridisLite::viridis(length((breaks_8))),
    na.value = NA,
    oob      = scales::squish,
    breaks   = ceiling(breaks_8),
    limits   = c(0, 11000),
    trans    = "sqrt",
    # breaks   = c(0, 25, 50, 75, 100, 200, 400),
    # breaks   = c(0, 500, 1000, 2000, 3000, 6000, 8000, 11000),
    # limits   = c(0, 400),
  ) +
  
  ggspatial::annotation_north_arrow(
    location = "br", 
    # pad_x = unit(0.4, "in"), 
    pad_y = unit(0.4, "in"),
    style  = ggspatial::north_arrow_orienteering(fill = c("white", "white"))
    ) + 
  ggspatial::annotation_scale(location = "br") +
  
  scale_y_continuous(breaks = seq(16.52, 16.68, 0.02), labels = plain_coord,
                     expand = expansion()) +
  scale_x_continuous(breaks = seq(-88.44, -88.32, 0.02), labels = plain_coord
                     ) +
  guides(
    color = guide_legend(order = 1),
    fill  = guide_legend(order = 2, reverse = TRUE)
  ) +
  labs(
    x = NULL,
    y = NULL
  ) +
  coord_sf(
    expand = FALSE,
    ylim = bbox[c(2, 4)],
    xlim = bbox[c(1, 3)],
    label_axes = "ENEN"
  ) +
  theme_bw() +
  theme(
    legend.position        = "inside",
    # legend.position.inside = c(0.15, 0.75),
    legend.position.inside = c(0.8, 0.4),
    legend.title           = element_text(size = rel(0.75)),
    legend.text            = element_text(size = rel(0.75)),
    axis.ticks             = element_blank()
  )


```

## 6.2 Save Interpolation Map
```{r save-interp}
# set location and name to save plot
plt_save3 <- 
  here("data", "raw", "placencia",
  paste0("placencia_mangrove_map_interp_l", Sys.Date(), ".jpeg")
  # paste0("placencia_mangrove_map_interp_black2_", Sys.Date(), ".jpeg")
  ) %T>% 
  print()

if (FALSE) {
  ggsave(
    filename = plt_save3,
    plot = last_plot(),
    device   = "jpeg",
    dpi      = 600,
    width    = 8.5,
    height   = 9,
    units    = "in",
    bg       = "white"
  )

  ggsave(
    filename = str_replace(plt_save3, "jpeg", "svg"),
    plot = last_plot(),
    device   = "svg",
    dpi      = 600,
    width    = 8.5,
    height   = 9,
    units    = "in",
    bg       = "white"
  )
} else {
  message("Not saving mangrove map with points!")
}
```


## 6.3 Load OpenStreetMap from Maptiles and Plot

```{r interpo-imagery}
# shelf(OpenStreetMap)
# 
# # bounds <- 
# #   st_transform(placencia_bbox, crs = wgs_crs) %>%
# #   st_bbox() %>%
# #   as.numeric()
# 
# bounds <- c(-88.5, -88.2, 16.4,   16.7)
# 
# 
# # extract openstreemap tiles
# osm_map <-
#   openmap(
#     bounds[c(4, 1)], # upperleft
#     bounds[c(3, 2)], # loweright
#     zoom = 14,
#     # type = "esri-physical", 
#     type = "esri-imagery", 
#     mergeTiles = TRUE
#   ) %>%
#   openproj(projection = paste("EPSG:", belize_epsg))

# plot openstree
# 
ggplot() + 
  geom_spatraster_rgb(data = belize_tile) +
  # OpenStreetMap::autoplot.OpenStreetMap(osm_map) +
  
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA) +
  
  # add interpolated map
  tidyterra::geom_spatraster(
    data = idw_rast_smooth,
    aes(fill = var1.pred),
    na.rm   = TRUE,
    maxcell = Inf
  ) +
  
  # add sample points
  geom_sf(
    data = biom_df,
    # aes(fill = total_biomass_kg, geometry = geometry),
    # aes(fill = total_biomass_per_100_sq_m, geometry = geometry),
    aes(fill = total_biomass_per_100_sq_m, geometry = geometry, color = upper_lower),
    alpha  = 0.75,
    size   = rel(1.5),
    stroke = rel(0.75),
    # color  = "white",
    shape  = 21,
    key_glyph = draw_key_rect
  ) +
  
  ggrepel::geom_text_repel(
    data = biom_df,
    aes(x = x, y = y, label = station_id, color = upper_lower),
    min.segment.length = 0,
    max.overlaps       = 100,
    # color              = "white",
    bg.color           = "black",
    segment.color      = "black",
    # bg.color           = "white",
    bg.r = 0.05,
    # segment.color      = "white",
    alpha              = 0.75,
    show.legend = FALSE
  ) +
  
  # region colors
  scale_color_manual(
    name   = "Region",
    values = region_col,
    breaks = names(region_col)
    ) +
  
  scale_fill_stepsn(
    # name     = "Total Biomass (kg)",
    name     = expression("Total Biomass (kg"~100~m^-2*")"),
    colours  = viridisLite::viridis(length((breaks_8))),
    na.value = NA,
    oob      = scales::squish,
    breaks   = ceiling(breaks_8),
    limits   = c(0, 11000),
    trans    = "sqrt",
    # breaks   = c(0, 500, 1000, 2000, 3000, 6000, 8000, 11000),
    # breaks   = c(0, 25, 50, 75, 100, 200, 400),
    # limits   = c(0, 400),
  ) +
  
  ggspatial::annotation_north_arrow(
    location = "br", 
    pad_y    = unit(0.4, "in"),
    style    = ggspatial::north_arrow_orienteering(fill = c("white", "white"))
    ) + 
  ggspatial::annotation_scale(location = "br", text_col = "white") + 
  
  scale_y_continuous(breaks = seq(16.52, 16.68, 0.02), labels = plain_coord) +
  scale_x_continuous(breaks = seq(-88.44, -88.32, 0.02), labels = plain_coord) +
  guides(
    color = guide_legend(order = 1),
    fill  = guide_legend(order = 2, reverse = TRUE)
  ) +
  labs(
    x = NULL,
    y = NULL
  ) +
  coord_sf(
    expand = TRUE,
    # ylim = c(bbox[c(2)] - 50, bbox[c(4)]),
    ylim = bbox[c(2, 4)],
    xlim = bbox[c(1, 3)],
    label_axes = "ENEN"
  ) +
  theme_bw() +
  theme(
    legend.position        = "inside",
    # legend.position.inside = c(0.15, 0.75),
    legend.position.inside = c(0.8, 0.4),
    legend.title           = element_text(size = rel(0.75)),
    legend.text            = element_text(size = rel(0.75)),
    axis.ticks             = element_blank()
  )
```

```{r save-interp}
# set location and name to save plot
plt_save4 <- 
  here("data", "raw", "placencia",
  paste0("placencia_mangrove_map_interp_imagery_", Sys.Date(), ".jpeg")
  ) %T>% 
  print()

if (FALSE) {
  ggsave(
    filename = plt_save4,
    plot = last_plot(),
    device   = "jpeg",
    dpi      = 600,
    width    = 8.5,
    height   = 9,
    units    = "in",
    bg       = "white"
  )
} else {
  message("Not saving mangrove map with points!")
}
```

# 7.0 Save Interpolation as `tiff` File

```{r save-tiff}
idw_rast_smooth
tiff_save <- 
  here("data", "raw", "placencia",
  paste0("placencia_mangrove_map_interp_", Sys.Date(), ".tiff")
  ) %T>% 
  print()
terra::writeRaster(idw_rast_smooth, tiff_save)

# load tiff
mang_bio <- 
  here("data", "raw", "placencia") %>%
  dir_ls(regexp = "placencia_mangrove_map_interp.*tiff") %>%
  sort(decreasing = TRUE) %>%
  .[1] %>%
  # rast(tiff_save) %T>% 
  rast() 
mang_bio %T>% 
  print() %>%
  plot()
```

```{r}
sector_shp <- 
  here("data", "raw", "placencia") %>%
  dir_ls(regexp = "merged") %>%
  dir_ls(regexp = "shp") %>%
  sf::st_read() %>%
  mutate(
    region2 = str_replace(region, "_", " "),
    region2 = str_to_title(region2),
    region2 = str_replace(region2, "Creek", "Creeks"),
  )

ggplot() +
  geom_sf(
    data = sector_shp, aes(fill = region2)
  ) 


  ggplot() +
  geom_sf(data = belize_sf, fill = "grey", alpha = 0.5) +
  geom_sf(
    data = sector_shp, aes(fill = region2)
  ) +
  ggspatial::annotation_north_arrow(
    location = "br", 
    # pad_x = unit(0.4, "in"), 
    pad_y = unit(0.4, "in"),
    style  = ggspatial::north_arrow_orienteering(fill = c("white", "white"))
    ) + 
  ggspatial::annotation_scale(location = "br") +
  
  ggrepel::geom_text_repel(
    data = biom_df,
    aes(x = x, y = y, label = station_id),#, color = upper_lower),
    min.segment.length = 0,
    max.overlaps       = 100,
    color              = "white",
    bg.color           = "black",
    segment.color      = "black",
    # alpha              = 0.75,
    segment.ncp = 10,
    show.legend = FALSE,
  ) +
    
  geom_point( 
     data = biom_df,
    aes(x = x, y = y),
    ) +  
  labs(
    x = NULL,
    y = NULL,
    fill = "Region"
  ) +
  
  scale_fill_manual(values = region_col) +
  # scale_color_manual(
  #   name = "Region",
  #   values = region_col,
  #   breaks = names(region_col)
  #   ) +
  scale_y_continuous(breaks = seq(16.52, 16.68, 0.02), labels = plain_coord,
                     expand = expansion()) +
  scale_x_continuous(breaks = seq(-88.44, -88.32, 0.02), labels = plain_coord
                     ) +
  
  coord_sf(
    expand = FALSE,
    ylim = bbox[c(2,4)],
    xlim = bbox[c(1,3)],
    label_axes = "ENEN"
  ) +
  theme_bw()

```


```{r}
plt_save5 <- 
  here("data", "raw", "placencia",
  paste0("placencia_mangrove_map_section", Sys.Date(), ".jpeg")
  # paste0("placencia_mangrove_map_interp_black2_", Sys.Date(), ".jpeg")
  ) %T>% 
  print()

if (FALSE) {
  ggsave(
    filename = plt_save5,
    plot = last_plot(),
    device   = "jpeg",
    dpi      = 600,
    width    = 8.5,
    height   = 9,
    units    = "in",
    bg       = "white"
  )

  ggsave(
    filename = str_replace(plt_save5, "jpeg", "svg"),
    plot = last_plot(),
    device   = "svg",
    dpi      = 600,
    width    = 8.5,
    height   = 9,
    units    = "in",
    bg       = "white"
  )
} else {
  message("Not saving mangrove map with points!")
}
```


```{r}
plot(st_geometry(sector_shp), main = "Total Area")
upper_region <- sector_shp[sector_shp$region2 == "Upper", ]
plot(st_geometry(upper_region), main = "Upper")

upper_bio <- crop(mang_bio, vect(upper_region))

lower_region <- sector_shp[sector_shp$region2 == "Lower", ]
plot(st_geometry(lower_region), main = "Lower")

lower_bio <- crop(mang_bio, vect(lower_region))

main_region <- sector_shp[sector_shp$region2 == "Main Creeks", ]
plot(st_geometry(main_region), main = "Main Creeks")

main_bio <- crop(mang_bio, vect(main_region))

plot(mang_bio)
plot(upper_bio)
plot(lower_bio)
plot(main_bio)
# total biomass per region
bind_rows(
  global(mang_bio, fun = "sum", na.rm = TRUE),
  global(upper_bio, fun = "sum", na.rm = TRUE),
  global(lower_bio, fun = "sum", na.rm = TRUE),
  global(main_bio, fun = "sum", na.rm = TRUE)
) %>%
  filter(!is.na(sum)) %>%
  bind_cols(
    region = c("Total", "Upper", "Lower", "Main Creeks")
  ) %>%
  mutate(
    percent = sum / max(sum) * 100,
    percent = round(percent, 2),
    biomass_kg = scales::label_comma()(sum)
  ) %>%
  bind_cols(
    ha_mult = c(
      1,
      25287.9 * 1000 * 2.0109,
      19868.98 * 1000 * 6.7776,
      23183.8 * 1000 * 4.0935
    )
  ) %>%
  mutate(
     ha_mult = scales::label_comma()(ha_mult)
  )

upper_region %>%
  mutate(
    area    = st_area(geometry),
    area_km = units::set_units(area, "km2"),
    area_km = units::drop_units(area_km)
    ) %>%
  summarise(sum = sum(area_km)) %>%
  as.data.frame()
lower_region %>%
  mutate(
    area    = st_area(geometry),
    area_km = units::set_units(area, "km2"),
    area_km = units::drop_units(area_km)
    ) %>%
  summarise(sum = sum(area_km)) %>%
  as.data.frame()
main_region %>%
  mutate(
    area    = st_area(geometry),
    area_km = units::set_units(area, "km2"),
    area_km = units::drop_units(area_km)
    ) %>%
  summarise(sum = sum(area_km)) %>%
  as.data.frame()
```

