---
title: "REACH MAp"
author: "Sebastian Di Geronimo"
date: 2025-10-13
format: html
editor: source
---

# 1.0 ---- Summary of Document ----



# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional
  readxl, janitor, hms,
  OpenStreetMap, leaflet,
  rerddap,
  sf, metR
  
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  

```


## 2.2 Load Dataset

```{r load-data}
# set location to save data and plots
dir_plt_save  <- here("data", "plots", "reach_map")

# select path to cruise directory
dir_path <- 
  here() %>%
  rstudioapi::selectDirectory()

# metadata logsheet
log_path <- 
  dir_path %>%
  dir_ls(regexp = "(?i)logsheet") %>%
  str_subset("(?i)~|aquatic", negate = TRUE) %T>% 
  print()

# sheet names
sheet_names_log <-
  log_path %>%
  readxl::excel_sheets() %T>% 
  print()

# load filter metadata
filter_meta <- 
  log_path %>%
  read_excel(
    sheet        = sheet_names_log[str_detect(sheet_names_log, "logsheet")], 
    .name_repair = janitor::make_clean_names
    ) %>%
  select(
   "location_id" = station, 
   "date" = date_mm_dd_yy, "time" = sample_collection_time_gmt,
   "long" = lon_actual, "lat" = lat_actual
  ) %>%
  filter(!is.na(location_id), !is.na(time)) %>%
  distinct() %>%
  arrange(date, time) %>%
  mutate(
    time    = hms::as_hms(time), 
    lon_lag = dplyr::lead(long, default = last(long)),
    lat_lag = dplyr::lead(lat, default = last(lat)),
    ) %T>% 
  print()

# zooplankton net metadata
zoo_label <-
  log_path %>%
  read_excel(
    sheet = sheet_names_log[str_detect(sheet_names_log, "zoo_log")],
    .name_repair = janitor::make_clean_names
  ) %>%
  filter(!is.na(mesh_size_um)) %>%
  distinct(cruise_id, location_id, net_type) %>%
  mutate(sampled = "x") %>%
  pivot_wider(
    id_cols      = c(cruise_id, location_id),
    names_from   = c(net_type),
    values_from  = c(sampled), 
    names_sep    = "_",
    names_repair = janitor::make_clean_names
    ) %>%
  mutate(
    .keep = "unused",
    label = case_when(
      (!is.na(ba) | !is.na(bb)) & !is.na(t) ~ "Bongo/Tucker",
      (!is.na(ba) | !is.na(bb)) ~ "Bongo",
      !is.na(t) ~ "Tucker",
    )
  ) %T>%
  print()

# merge filter and zooplankton metadata
metadata <-
  left_join(
    filter_meta,
    zoo_label,
    by = c("location_id")
  ) %>%
  mutate(
    label = case_when(
      is.na(label) ~ "CTD",
      .default = paste0("CTD/", label)
    )
  ) %T>% 
  print()

# expected locations from schedule
loc_path <- 
  dir_path %>%
  dir_ls(regexp = "(?i)schedule") %>%
  str_subset("~", negate = TRUE) %T>% 
  print()

sheet_names_expec <- 
  loc_path %>%
  excel_sheets() %T>% 
  print()

loc_dat <- 
  loc_path %>%
  read_xlsx(
    sheet = sheet_names_expec[str_detect(sheet_names_expec, "(?i)location")],
    .name_repair = janitor::make_clean_names, 
    ) %>%
  janitor::remove_empty("cols") %>%
  drop_na(latitude_decimal_degree, longitude_decimal_degree)  %>%
  filter(str_detect(description, "CTD")) %>%
  select(
    waypoint, description, depth_m,
    "lat" = latitude_decimal_degree, "long" = longitude_decimal_degree, 
    ) %>%
  mutate(
    description = if_else(str_detect(description, "0"), "CTD/Net Trawl", description),
    long = -long
    ) %T>% 
  print()


# stations not sampled
stn_miss <- 
  loc_dat %>%
  filter(!str_detect(waypoint, str_c(metadata$location_id, collapse = "|"))) %>%
  rename("location_id" = waypoint) %>%
  mutate(label = "Missed") %T>% 
  print()

plot_data <-
  bind_rows(
    metadata,
    stn_miss
  )
```

## 2.3 Load Rice's Whale Distrbution Shapefile
Shapefile of Rice's Whale Distrbution from:
<https://www.fisheries.noaa.gov/resource/map/rices-whale-core-distribution-area-map-gis-data>
```{r load-data}
# shapefile path
whale_shp_path <- 
  here("data", "raw") %>%
  dir_ls(regexp = "Rices_whale") %>%
  dir_ls(regexp = "shp$")

# load whale distrbution shapefile
whale_shp <- st_read(whale_shp_path)
```


# 3.0 Plot using `leaflet`

Overall points

```{r leaflet}
leaflet::leaflet() %>%
  addTiles() %>%
  addMarkers(loc_dat$long, loc_dat$lat)

```


## 3.1 Load Bathymetry

```{r load-bathymetry}
source(here("scripts", "map_files_dwnlod.R"))

# set bathymetric shapfile locations
bathy_path <- here("data", "map_shp", "reach")

# add boundaries for openstreemap
adj_deg <- 1.5 # degree widen
bounds <- 
  c(
    range(loc_dat$long) + c(-adj_deg, adj_deg),
    range(loc_dat$lat) + c(-adj_deg, adj_deg - 0.8)
  ) %T>%
  print()

# change default ERDDAP to NOAA Coast Watch
Sys.setenv(RERDDAP_DEFAULT_URL = "https://coastwatch.pfeg.noaa.gov/erddap/")

# download bathymetric data to path_topo location
world_download(
  path_land     = NULL,
  path_topo     = bathy_path,
  extent        = bounds,
  etopo_version = "ETOPO_2022_v1_15s"
)

# change to default ERDDAO
Sys.unsetenv("RERDDAP_DEFAULT_URL")

# read bathymetric data
bathy <- 
  bathy_path %>%
  dir_ls()  %>%
  terra::rast() %>%
  as.data.frame(xy = TRUE)

```

```{r}
# long_lat_lag <- 
#   loc_dat %>%
#   filter(!str_detect(waypoint, str_c(metadata$location_id, collapse = "|"))) %>%
#   mutate(
#     lon_lag = dplyr::lead(long, default = last(long)),
#     lat_lag = dplyr::lead(lat, default = last(lat)),
#     ) %>%
#   mutate(
#     .by = waypoint,
#     te = row_number(),
#     t2 = n(),
#     waypoint = if_else(row_number() > 1, "", waypoint)
#   ) %T>% 
#   print()

# # stations not sampled
# stn_miss <- 
#   loc_dat %>%
#   filter(!str_detect(waypoint, str_c(metadata$location_id, collapse = "|"))) %>%
#   rename("station" = waypoint) %>%
#   mutate(label = "Missed") %T>% 
#   print()


```

# 4.0 Plt using `OpenStreeMap`

Inspired from: <https://ajsmit.github.io/Intro_R_Official/mapping-google.html>

```{r mangrove-stations}
# add boundaries for openstreemap
# adj_deg <- c(0.3, 1.25) # degree widen
# bounds <- 
#   c(
#     range(loc_dat$long) + c(-adj_deg[1], adj_deg[2]),
#     range(loc_dat$lat) + c(-adj_deg[1], adj_deg[2] - 0.8)
#   ) %T>%
#   print()

# bounds[1] <- -88.5
# bounds[2] <- -82
# bounds[3] <- 26.5

bounds <- c(-88.5, -82, 26.5, 30.6)

# extract openstreemap tiles
osm_map <-
  openmap(
    bounds[c(4, 1)], # upperleft
    bounds[c(3, 2)], # loweright
    zoom = 8,
    type = "esri-physical", 
    mergeTiles = TRUE
  ) %>%
  openproj()

# OpenStreetMap::autoplot.OpenStreetMap(osm_map)
  
(loc_map <-
  # plot openstreemap
  OpenStreetMap::autoplot.OpenStreetMap(osm_map) +
    
  # contour lines
  metR::geom_contour2(
    data = bathy,
    aes(
      x = x,
      y = y,
      z = -z
    ),
    col = "grey90",
    breaks = c(3000, 2000, 1000, 500, 100, 50, 25)
  ) +
    
  # whale habitat polygon
  geom_sf(
    data        = whale_shp,
    color       = "grey40",
    fill        = prismatic::clr_lighten("grey40"),
    linewidth   = 0.75,
    alpha       = 0.5,
    linetype    = "dashed",
    inherit.aes = FALSE
  ))
```

TODO: need to add a way to change the description for when only sampled CTD, or 
bongo or tucker


```{r plt-stations}
# shp_label_dict <-
#   c(
#     # "Arrive"           = 21,
#     # "Waypoint"         = 21,
#     "CTD"              = 24,
#     "CTD/Bongo"        = 21,
#     "CTD/Bongo/Tucker" = 22,
#     "CTD/Tucker"       = 22,
#     "Bongo"            = 22,
#     "Bongo/Tucker"     = 22,
#     "Tucker"           = 22,
#     "Surface Buoy/CTD" = 25,
#     "Missed"           = 4,
#   )
# col_label_dict <-
#   c(
#     # "Arrive"           = 21,
#     # "Waypoint"         = 21,
#     "CTD"              = "lightblue",
#     "CTD/Bongo"        = "#546",
#     "CTD/Bongo/Tucker" = "#ffd",
#     "CTD/Tucker"       = "#778",
#     "Bongo"            = "#434",
#     "Bongo/Tucker"     = "yellow",
#     "Tucker"           = "green",
#     "Surface Buoy/CTD" = "blue",
#     "Missed"           = "red"
#   )

scale_key <- 
  tribble(
                ~label, ~shape,       ~fill,  ~color,
                 "CTD",     24, "lightblue", "black",
           "CTD/Bongo",     21,      "#546", "black",
    "CTD/Bongo/Tucker",     22,      "#ffd", "black",
          "CTD/Tucker",     22,      "#778", "black",
               "Bongo",     22,      "#434", "black",
        "Bongo/Tucker",     22,    "yellow", "black",
              "Tucker",     22,     "green", "black",
    "Surface Buoy/CTD",     25,      "blue", "black",
              "Missed",      4,       "red",   "red"
  )

loc_map_pts <-
  # ggplot() +
  loc_map +

  # add arrows for cruise track
  ggarchery::geom_arrowsegment(
    data = plot_data,
    aes(
      x     = long,
      xend  = lon_lag,
      y     = lat,
      yend  = lat_lag,
      group = fct_reorder(format(date, "%B %d"), date, .na_rm = TRUE)
    ),
    arrow_positions = 0.5,
    arrows          = arrow(length = unit(0.1, "inches")),
    na.rm           = TRUE
  ) +
  
  # add stations sampled and missed
  geom_point(
    data = plot_data,
    aes(
      x     = long, 
      y     = lat, 
      shape = label,
      color = label,
      fill  = label
      ),
    size = 3,
  ) +
  scale_color_manual(
    values = setNames(scale_key$color, scale_key$label),
    name   = NULL
    ) +
  ggnewscale::new_scale_color() +
  
  # add station labels
  ggrepel::geom_text_repel(
    data = plot_data,
    aes(
      x     = long, 
      y     = lat, 
      label = location_id, 
      color = label
      ),
    max.overlaps       = 15,
    min.segment.length = 0.0,
    size               = 3.5,
    fontface           = "bold",
    bg.color           = "black",
    bg.r               = 0.15,
    segment.color      = "black", 
    show.legend        =  FALSE
  ) +
  labs(
    title = "REACH-004 (March 1 - 7)",
    x     = NULL,
    y     = NULL,
    color = NULL,
    shape = NULL,
    fill  = NULL
  ) +
  
  # scales guides
  scale_shape_manual(values = setNames(scale_key$shape, scale_key$label)) +
  scale_fill_manual(values  = setNames(scale_key$fill, scale_key$label)) +
  scale_color_manual(values = setNames(scale_key$fill, scale_key$label)) +
  guides(
    color = guide_legend(position = "bottom", nrow = 1),
    shape = guide_legend(position = "bottom", nrow = 1),
    fill = guide_legend(position = "bottom", nrow = 1),
  ) +
    
  coord_sf(expand = FALSE, xlim = bounds[1:2], ylim = bounds[3:4]) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text     = element_text(size = 7)
  )
```


```{r show-stations}
loc_map_pts
```
```{r}
bind_rows(
  mutate(metadata),
  stn_miss
)
```

```{r}
loc_dat$description %>%
  unique()

```


# 5.0 Save Plots using `camcorder`

After running, will need to redisplay plots for it to

```{r camcorder}
shelf(camcorder)
camcorder::gg_record(
  dir    = here::here("data", "cam_test", "reach_map"),
  device = "png",
  dpi    = 600,
  width = 11,
  height = 8.5,
  units  = "in",
  bg     = "white"
)
camcorder::record_polaroid()
shell.exec(here::here("data", "cam_test", "reach_map"))

# uncomment if need to resize
# camcorder::gg_resize_film(width = 90, height = 90 * 0.5625,  dpi = 600, units = "mm")

```

## 5.1 When done saving plots

```{r stop-camcorder}
camcorder::gg_stop_recording()
```


```{r}
bathy2 <- 
  bathy_path %>%
  dir_ls()  %>%
  terra::rast()

# 1. Extract contours using base R
clines <- contourLines(
  x = sort(unique(bathy$x)),
  y = sort(unique(bathy$y)),
  z = rev(matrix(-bathy$z, nrow = length(unique(bathy$x)), byrow = FALSE)),
  levels = c(25, 50, 100, 500, 1000, 2000, 3000)
)
  rev(matrix(-bathy$z, nrow = length(unique(bathy$x)), byrow = FALSE)[1:10, 1:10])
clines %>%
  str()
# 2. Find the longest segment for each level
contour_df <- do.call(rbind, lapply(clines, function(cl) {
  data.frame(
    level = cl$level, 
    x = cl$x, 
    y = cl$y, 
    length = length(cl$x)
  )
}))
unique(contour_df$level)
longest <- 
  contour_df %>%
  # group_by(level) %>%
  summarise(
    .by = level,
    # Use the segment with the most points (or your preferred metric)
    idx = which.max(length)
  )
distinct(contour_df, length)
# 3. Label only these segments
# (Alternatively, you may need to join or filter your original 'bathy' or contour data for plotting)

# 4. In your ggplot: use metR geom_label_contour/data for those only (omitting many details since
# extra wrangling is needed).
label_points <- contour_df %>%
  group_by(level) %>%
  summarise(
    x_label = mean(x),  # or x[middle point]
    y_label = mean(y)
  )

ggplot() +
  geom_path(
    data = contour_df,  # all your contour segments for all levels
    aes(x = x, y = y, group = interaction(level, length), color = factor(level))
  ) +
  geom_text(
    data = label_points,
    aes(x = x_label, y = y_label, label = level),
    color = "black", fontface = "bold", size = 4
  )
```



