---
title: "A script for test"
author: "Sebastian Di Geronimo"
date: "2023-10-07"
format: html
---

# 1.0 ---- Summary of Document ----
- 3.0: Illustration of projecting data points


# 2.0 ---- Setup ----

## 2.1 Load Libraries
```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  # broom # optional
  
  # additional
  
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
```

# 3.0 ---- Illustration of projecting data points ----
From: 
\cite{shaliziAdvancedDataAnalysis2021}
“Figure 15.1” ([Shalizi, 2021, p. 321](https://www.stat.cmu.edu/~cshalizi/ADAfaEPoV/))

Illustration of projecting data points $ \vec x $  (black dots) on to an 
arbitrary line through the space (blue, dashed), represented by a unit 
vector $ \vec w $ along the line (also blue but solid). The blue dots are the 
projections on to the blue line,$ (\vec x~\dot~\vec w)\dot~\vec w $ ; the gray 
lines are the vector residuals,$ \vec x~-~(\vec x~\dot~\vec w)\dot~\vec w $. 
These are not the residuals from regressing one of the components of the data 
vector on the other.


```{r section-3.0}
demo.theta <- runif(10, min    = 0, max = pi/2) 
demo.x <- cbind(cos(demo.theta), sin(demo.theta)) 
demo.x <- scale(demo.x, center = TRUE, scale = FALSE) 

demo.w <- c(cos(-3 * pi/8), sin(-3 * pi/8)) 

# x * y * x
# [10 x 2] * [2 x 1] * [10 x 2]
projection <- demo.x %*% demo.w %*% demo.w 

{
plot(demo.x, 
     xlab = expression(x^1), 
     ylab = expression(x^2), 
     xlim = c(-1, 1), 
     ylim = c(-1, 1))
arrows(0, 0, demo.w[1], demo.w[2], col = "blue") 
text(demo.w[1], demo.w[2], pos = 4, labels = expression(w)) 
abline(0, b = demo.w[2]/demo.w[1], col = "blue", lty = "dashed") 

points(projections, pch = 16, col = "blue") 
segments(x0 = demo.x[, 1], y0 = demo.x[, 2], x1 = projections[, 1], y1 = projections[, 2], col = "grey")
}
```
# 4.0 ---- Trelliscope Test ----

from [trelliscope](https://github.com/trelliscope/trelliscope)
web: [trelliscope reference manual](https://trelliscope.org/trelliscope/)

Supersedes: [trelliscopejh](https://github.com/hafen/trelliscopejs)
  - although has a nice looking interface: 
    - [JS version](https://hafen.github.io/trelliscopejs/articles/trelliscopejs.html#-ggplot2-interface-facet-trelliscope)


Not to be confused with: [trelliscope](https://github.com/delta-rho/trelliscope)

```{r trelliscope}
shelf(trelliscope)

szn <- c("Winter", "Spring", "Summer", "Autumn")

# rstudioapi::selectDirectory()
file_dir <- here("..", "new_MBON_phytoplankton_pigments", "data", "plots", "map", "ratio")

test <-
  file_dir %>%
  dir_ls() %>%
  tibble(img_src = .) %>%
  # slice(1:2) %>%
  mutate(
    .before = 1,
    id      = row_number(), 
    base    = basename(img_src),
    base    = path_ext_remove(base),
    season  = str_extract(base, ".*_ratio_(.*)_20", group = 1),
    season  = str_to_title(season),
    season  = fct(season, szn),
    pigment = str_extract(base, "mbon_pigm_(.*)_ratio", group = 1),
    ) %>% 
  as_trelliscope_df(
    name = "Display Pigments", 
    key_sig = "img_src",
    key_cols = c("season", "pigment"),
    tags = "base_name"
    ) %>%
    set_default_layout(ncol = 2, page = 2) %>%
    add_inputs(
      input_checkbox(
        name = "Checkbox Input",
        label = "A space to add custom button inputs",
        options = c("yes", "no")
      )
    )

show_info(test)
view_trelliscope(test) 


# here("C:\\Users", "spd19", "AppData", "Local", "Temp", "Rtmpsf3mnt", "file65f4368151a3", "index.html") |> shell.exec()
```
```{r}
test2 <- 
  mars_rover %>%
  slice_sample(prop = 0.01) %>%
  rename("rando" = img_src) %>%
  as_trelliscope_df(name = "test2")

show_info(test2)
```


```{r}
# Use `as_trelliscope_df()` to convert panel metadata to a special
# trelliscope data frame
library(ggplot2)
library(dplyr)

panel_dat <-
  (
  ggplot(gap, aes(year, life_exp)) +
    geom_point() +
    facet_panels(vars(country, continent))
  ) |>
    as_panels_df()

meta_dat <- 
  summarise(
    gap,
    .by = c(country, continent),
    mean_life_exp = mean(life_exp),
    min_life_exp = min(life_exp),
    max_life_exp = max(life_exp),
    mean_gdp = mean(gdp_percap)
  )

joined_dat <- 
  left_join(panel_dat, meta_dat) |>
  as_trelliscope_df(name = "life_expectancy", path = tempfile())

if (FALSE) {
view_trelliscope(joined_dat)
}


trell <- 
  joined_dat |>
# as_trelliscope_df(name = "Select Images") |>
  add_inputs(
    # input_radio(
    #   name = "Radio Input",
    #   label = "A space to add custom ranking for sorting",
    #   options = c("yes", "no")
    # ),
    input_select(
      name = "Select Input",
      label = "A space to add custom dropdown inputs",
      options = c("yes", "no"),
      active = TRUE
    ),
    # input_checkbox(
    #   name = "Checkbox Input",
    #   label = "A space to add custom button inputs",
    #   options = c("yes", "no")
    # ),
    vars = c("continent", "country")
  )

view_trelliscope(trell)


here("data") %>%
  dir_ls(regexp = "_2trel") %>%
  read_csv() 
```


```{r}
shelf(trelliscope)
# rstudioapi::selectDirectory()
file_dir <- 
  here("C:\\Users", "spd19", "Box", "Davidson_Fellowship_2022", "SAMPLING", 
       "FEBRUARY_2023", "Mairim", "photo") 

save_csv_name <- 
  str_extract(file_dir, ".*((?i)sampling)/(.*)", group = 2) %>%
  str_replace_all("/", "_") %>%
  str_to_lower() %T>% print()  
  
nat_files <-
  file_dir %>%
  dir_ls(regexp = "\\.JPG") %>%
  tibble(src_img = .) %>%
  # slice(1:15) %>%
  mutate(
    .before = 1,
    id      = str_extract(src_img, "GOPR(\\d+)\\.JPG", group = 1)
    ) %>%
  mutate(
    dir_og = dirname(src_img),
    loc  = str_extract(src_img, "SAMPLING/(.*20\\d{2})/(.*)/photo", group = 2),
    time = str_extract(src_img, "SAMPLING/(.*20\\d{2})", group = 1),
    time = str_to_title(time),
    time = str_replace(time, "_", " ")
  )  %T>% print() %>%
  as_trelliscope_df(name = str_c(save_csv_name, "_to_remove")) %>%
  set_default_layout(ncol = 4, page = 2) %>%
  add_inputs(
    input_radio(
      name = "remove_bad_images",
      label = "Yes images will be removed, no imagaes will be saved",
      options = c("yes", "no"),
      active = TRUE
    ),
    vars = c("dir_og", "src_img", "loc", "time")
  )

view_trelliscope(nat_files)

```

```{r}
files_mv <- 
  here("data") %>%
  dir_ls(regexp = "(?i)photo_to_remove") %>%
  str_subset("_1") %>%
  read_csv(show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  mutate(
    # .keep = "used",
    reg = basename(src_img),
    # og_loc2 = dirname(dir_og),
    file_rm = map2_chr( 
      .x = dir_og,
      .y = reg,
      ~ dir_ls(.x, regexp = .y)),
    # new_loc = str_c(dirname(dir_og), "bad", sep = "/"),
    new_loc = str_c(here("data"), "bad", sep = "/")
  ) %T>% print()


dir_create(unique(files_mv$new_loc))

files_mv  %>%
  filter(str_detect(remove_bad_images, "(?i)yes"))  %>%
  nest(.by = c(dir_og, new_loc)) %$%
  walk2(
    .x = new_loc,
    .y = data,
    \(.x, .y) {
     cli::cli_alert_info("Moving files to {.path {.x}}") 
     cli::cat_bullet(basename(.y$file_rm))
     
     if (menu(c("Yes, move files", "No, don't move files")) == 1) {
       cli::cli_alert_info("Moving files!")
       dir_create(unique(.x))
      file_copy(.y$file_rm, .x)
      # file_move(.y$file_rm, .y$new_loc)
       
     }
     
    }
  )


```


```{r}
dir_create(here("data", "bad2"))

here("data", "bad") %>%
  dir_ls() %T>% print() %>% 
  file_move(., here("data", "bad2"))
  
```

# 5.0 ---- Test Formula Notation ----

From [datacamp](https://www.datacamp.com/tutorial/r-formula-tutorial)

Section: `How to Concatenate Formulae`

```{r formula-notation}
i <- "y ~ x"
j <- "y ~ x + x1"
k <- "y ~ x + x1 + x2"

l <- c(i, j, k)

lapply(l, as.formula)
l

class(j)
class(as.formula(j))
class(y ~ x)

identical(as.formula(i), y ~ x)

```

# 6.0 ---- Test model.frame function ----

from [datacamp](https://www.datacamp.com/tutorial/r-formula-tutorial)

Section: `Modeling Functions`

Returns a dataframe that match the formula columns
- this may do some preprocessing of the data like log transformations

Subset can be added to filter the data further

```{r model-frame}
# use model.frame with the iris data set 

iris
model.frame(Sepal.Length ~ Sepal.Width, data = iris)
model.frame(
  Sepal.Length ~ Sepal.Width * Petal.Length, 
  data = iris)
model.frame(
  Sepal.Length ~ log(Sepal.Width) * Petal.Length, 
  data = iris)

model.frame(
  Sepal.Length ~ Sepal.Width * Petal.Length, 
  data = iris,
  subset = Species == "setosa" # extra filtering
  )
```


# 7.0 ---- Test `adespatial` ----

```{r test-adespatial}
vignette("tutorial", package = "adespatial")
```




# 8.0 ---- Test Kernal Density Estimation ----

```{r test-kde}
# equation for kernel density estimation

kernel_test <- function(radius, .dist) {
  
   max(0, radius^2 - .dist^2)
  
    # ---- end of function kernel_test
}

tibble(x = 1:10) %>%
  mutate(
    .dist = 5,
    radius = 2,
    kernel = map2_dbl(radius, .dist, kernel_test)
  ) 

```


# 9.0 ---- Test `openxlsx2` Formula Builder ----

Using `openxslx2::add_formula`
- basically you create an array formula where you use first col first row to
  first column last row to create an array formula
  i.e. "< column X > < row Y > : < column X > < row Z >" 

- then add the dimensions to put the results of the formula
  - same rows as formula but different columns
    i.e. dim = "< column XX > < row Y > : < column XX > < row Z >" 
  - NOTE: its recommended to have the same length (i.e. row 1 - 10 for both 
    formula and result)
    - they don't have to be the same though, but at least the same number of 
      rows (i.e. formula rows A1:A3 with results A4:A6)
    - if not same number of rows, the formula will stop wherever the row number
      stops


```{r}
shelf(openxlsx2)

set.seed(123)
# 3 x 5 matrix of random numbers from 0 to 1
dat <- matrix(runif(15), nrow = 3, ncol = 5) %T>% print()

# Create artificial xlsx file
wb <- 
  wb_workbook()$
  add_worksheet()$
  add_data(x = dat, col_names = FALSE)
  
# see startind wb
wb_to_df(wb, col_names = FALSE, , show_formula = TRUE)

# add formulas to individual rows
wb <- 
  wb$
  add_formula(x = "SUM(A2, B2)", dims = "F2")$
  add_formula(x = "A1 + B1", dims = "F1")$
  add_formula(x = "A3 * B3", dims = "F3")
  
# these formulas won't be evaluated until saved and opened in excel
wb_to_df(wb, col_names = FALSE, show_formula = TRUE)
wb$open()

# add array formulas
# run the same operations to many rows
# the formula will only show at the first instance of the formula but will be
# evaluated for all rows once saved
wb <- 
  wb_workbook()$
  add_worksheet()$
  add_data(x = dat, col_names = FALSE)$
  add_formula(x = "A1:A3 * B1:B3", 
              dims = "F1:F3",
              # dims = "A4:A5", # can be evaluated anywhere 
              array = TRUE) 
  
wb_to_df(wb, col_names = FALSE, show_formula = TRUE)

wb$open()

unshelf(openxlsx2)
```

