---
title: "Surface_maps"
author: "Mostafa Soliman"
date: "2026-03-06"
output: html_document
---


ONE cruise: 4-panel Surface Maps (0–5 m mean)
Variables: Salinity, Temperature, Oxygen, Chlorophyll (CHL_0_5m + chl_source)
Method: IDW (UTM grid) + LIGHT DISPLAY SMOOTHING

Edge: TIGHT HARD SUPPORT MASK + SOFT FEATHERED DISPLAY EDGE
Style: subtle contour lines + north arrow + scale bar

Title:  "WS0603 — Surface (0–5 m mean)"
Subtitle: "YYYY-MM-DD to YYYY-MM-DD
CHL panel title: "Chlorophyll fluorescence" OR "Chlorophyll concentration"
Units ONLY on colorbar title


ALL cruises: 4-panel Surface Maps (0–5 m mean) — LOOP VERSION
Variables: Salinity, Temperature, Oxygen, Chlorophyll (CHL_0_5m + chl_source)
Method: IDW (UTM grid) + LIGHT DISPLAY SMOOTHING

Style: subtle contour lines + north arrow + scale bar
Labels: station names on ALL panels with SHORT STRAIGHT ARROW leader lines
FOCUS RANGES (pattern emphasis):
  Salinity 28–35 PSU | Oxygen 1–5 mL/L | Temp 18–30 °C | Chl 0–5 mg m^-3
Output: PNG


# Load Libraries

```{r setup}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional packages, if from GitHub, include username
  sf, ggspatial, terra, geodata, gstat, viridis, ggrepel, grid, 
  patchwork # combine panels
)

conflicts_prefer(
  dplyr::filter,
  dplyr::select
  )
```

# File Paths

```{r paths}
base_dir <- here("data", "ctd_maps")
dir_create(base_dir)

ts_dir <- here(base_dir, "Time_series_new")
dir_create(ts_dir)

# file with all ctd casts
in_file <- 
  ts_dir  %>%
  # dir_ls(regexp = "CTD_surface_0_5m_ALL")
  dir_ls(regexp = "five_regions") %T>% 
  print()

stopifnot(length(in_file) != 0)

map_dir <- here(ts_dir, "Figures_freshening", "SurfaceMaps_0_5m_4vars_NOHALO")
dir_create(map_dir)
```

# Base Map of Florida

Obtained from `geodata` package

Map is projected into EPSG:32617 or EPSG:2237

```{r base-map}
# projection
utm_epsg <- 32617 # between 84W and 78W
# utm_epsg <- 2237 # for Florida zone 17N

# florida map object from geodata then projected
florida_sf <- 
  geodata::gadm(country = "USA", level = 1, path = tempdir()) %>%
  st_as_sf() %>%
  filter(NAME_1 == "Florida") %>%
  st_transform(4326) %>% 
  st_crop(xmin = -85, xmax = -79, ymin = 24, ymax = 29) %>%
  st_transform(utm_epsg) 

# set overall bounding box for plotting
bb_utm <-
  c(
    xmin = -84.5,
    xmax = -79.5,
    ymin = 24.2,
    ymax = 28.6
  ) %>%
  st_bbox(crs = 4326) %>%
  st_as_sfc() %>%
  
  # converts dd to meters
  st_transform(utm_epsg) %>%
  st_bbox()
```

# Smoothing Parameters

This does NOT change source data, only the plotted raster

```{r smoother-options}
# local IDW = tighter, less spread, more visually refined
idw_options <-
  list(
    "idp_pow"   = 2.0,
    "maxdist_m" = 120000,
    "nmin_nb"   = 3,
    "nmax_nb"   = 10
  )

# Light 3x3 weighted kernel (gentle smoothing only)
smooth_kernel <-
  matrix(
    c(
      1, 2, 1,
      2, 4, 2,
      1, 2, 1
    ),
    nrow = 3, byrow = TRUE
  )
smooth_kernel <- smooth_kernel / sum(smooth_kernel)
```

# Read Data 

```{r read-all-data}
df0 <- read_csv(in_file, show_col_types = FALSE) %>%
  rename(
    "latitude" = latitude_mean,
    "longitude" = longitude_mean,
    # "date_raw" = date
  ) %T>% 
  print()

# check missing columns in data file
# need_base <- c("cruise_file", "station", "latitude", "longitude")
need_base <- c("cruise_id", "station", "latitude", "longitude")
miss_base <- setdiff(need_base, names(df0))
if (length(miss_base) > 0) stop("Missing base columns: ", paste(miss_base, collapse = ", "))


# of the list of columns, pick the first one
pick_first_col <- function(candidates, col_names) {
  hit <- candidates[candidates %in% col_names]
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

var_names <- list(
  col_sal     = c("salinity_analysis", "S_0_5m", "sal_0_5m", "mean_sal_psu", "salinity"),
  col_temp    = c("temperature_analysis", "T_0_5m", "temp_0_5m", "mean_temp_C", "temperature", "temp_C"),
  col_o2      = c("dissolved_oxygen_analysis", "O2_0_5m", "oxy_0_5m", "mean_oxy", "oxygen", "DO"),
  col_date    = c("date", "Date", "cast_date", "datetime", "time", "timestamp"),
  # confirmed by you:
  col_chl     = c("chlorophyll_analysis", "chlorophyll_mean", "CHL_0_5m"),
  col_chl_src = c("chlorophyll_source_column", "chl_source")
) %>%
  map(
    y = names(df0),
    ~ pick_first_col(.x, .y)
  ) %>%
  unlist()

if (any(is.na(var_names))) {
  stop(
    paste0("Could not find column(s):\n- ", 
           paste(names(var_names[is.na(var_names)]), collapse = "\n- "))
       )
}

for (i in seq(var_names)) do.call("<-", list(names(var_names)[i], var_names[i]))

# rename columns
df0 <- 
  df0 %>%
  mutate(
    # cruise_file = as.character(cruise_file),
    cruise_file = as.character(cruise_id),
    station     = str_squish(as.character(station)),
    latitude    = suppressWarnings(as.numeric(latitude)),
    longitude   = suppressWarnings(as.numeric(longitude)),
    sal         = suppressWarnings(as.numeric(.data[[col_sal]])),
    temp        = suppressWarnings(as.numeric(.data[[col_temp]])),
    oxy         = suppressWarnings(as.numeric(.data[[col_o2]])),
    chl         = suppressWarnings(as.numeric(.data[[col_chl]])),
    chl_source  = as.character(.data[[col_chl_src]]),
    date_raw    = if (!is.na(col_date)) .data[[col_date]] else NA,
    # date_raw    = mdy(date_raw)
    date_raw    = ymd(date_raw)
  ) %>%
  arrange(date) %T>% 
  print()

# extract all crusie IDs
cruise_ids <- 
  df0 %>%
  pull(cruise_file) %>%
  unique() %>%
  tools::file_path_sans_ext() %T>% 
  print()

cat("Total cruises found:", length(cruise_ids), "\n")

# remove variables
rm(list = names(var_names), var_names)


max_date <- max(df0$date)
```


```{r salinity-ecdf}
df0 <- 
  df0 %>%
  filter(str_detect(region, "(?i)keys|thousand|everglades")) %>%
  mutate(
        region = factor(
      region,
      levels = c(
        "Ten_Thousand_Islands",
        "Everglades",
        "Lower_Keys",
        "Middle_Keys",
        "Upper_Keys"
      )
    )
  ) %>%
  filter(sal < 50)

# class int
brk_num <- 3 # edit number of breaks
brk_int <- 
  classInt::classIntervals(df0$sal, brk_num, style = "pretty") %T>% 
  print()

break_df <- data.frame(
  xmin = head(brk_int$brks, -1),
  xmax = tail(brk_int$brks, -1),
  y = 0,
  cls = factor(seq_len(length(brk_int$brks) - 1))
)

# ---- all regions together ---- #
ecdf_df <- df0 %>%
  select("x" = sal) %>%
  arrange(x) %>%
  mutate(
    .keep = "used",
    ecdf = seq_along(x) / length(x),
    range = between(x, 32, 34)
    )

ggplot() +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA) +
  # annotate("text", x = c(32, 34) - 0.35, y = -1, label = c("32", "34"), color = "grey50") +
  geom_rect(
    # data = break_df,
    data = tribble(
      ~xmin, ~xmax, ~cls,
      10, 32, 1,
      32, 34, 2,
      34, ceiling(max(df0$sal)), 3
      ) %>% mutate(cls = as.factor(cls)),
    aes(xmin = xmin, xmax = xmax, ymin = -0.15, ymax = -0.02, fill = cls),
    color = "grey40"
  ) +
  geom_point(data = ecdf_df, aes(x = x, y = ecdf), color = "black", alpha = 0.7, size = rel(1)) +
  geom_step(data = ecdf_df, aes(x = x, y = ecdf), color = "blue", alpha = 1) +
  # geom_vline(xintercept = brk_int$brks, color = "grey", linewidth = 0.4,
  geom_vline(xintercept = c(32, 34), color = "grey", linewidth = 1,
             linetype = "dashed") +
  annotate(
    "text",
    x = c(
    (10 + 32) / 2, 
    33, 
    (max(ecdf_df$x[ecdf_df$x > 34]) + 34) / 2
    ),
    y     = -0.15 / 2,
    size  = rel(3.9),
    label = c("Low Salinity (<32)", "Moderate\n(32-<34)", "Marine/Background (≥34)"),
    color = c("white", "white", "black"),
    lineheight = 0.8,
    fontface = "bold"
    ) +
  labs(
    # title = paste("ECDF with", brk_num, "Breaks"),
    x = "Salinity",
    y = "Cumulative Proportion"
  ) +
  # facet_wrap(~region) +
  # scale_fill_manual(values = viridisLite::viridis(nrow(break_df)), guide = "none") +
  scale_fill_manual(values = c("#457db4", "#39a585", "#cfcfcf"), guide = "none") +
  scale_x_continuous(breaks = scales::breaks_width(5), expand = expansion()) +
  # scale_x_continuous(breaks = c(seq(10,30,5), 32, 34, 35, 40, 45), expand = expansion()) +
  scale_y_continuous(expand = expansion(add = c(0, .1))) +
  theme_classic() +
  theme(
    # axis.title.x       = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.title = element_text(size = rel(1.5)),
    axis.title.y       = element_text(angle = 90),
    axis.text.y        = element_text(angle = 90, hjust = 0.5),
    axis.text          = element_text(size = rel(1.5)),
    axis.line          = element_line(linewidth = rel(1.5))
  )

if (FALSE) {
  ggsave(
    here(base_dir, glue("salinity_edcf_all_regions_{Sys.Date()}.jpeg")), 
    width  = 10, 
    height = 5, 
    dpi    = 600)
  shell.exec(base_dir)
}
```


```{r salinity-ecdf}
# ---- ecdf by each region ---- #
ecdf_df <- df0 %>%
  select("x" = sal, region) %>%
  arrange(x) %>%
  mutate(
    .by = region,
    .keep = "used",
    ecdf = seq_along(x) / length(x),
    region = factor(
      region,
      levels = c(
        "Ten_Thousand_Islands", "Everglades",
        "Lower_Keys", "Middle_Keys", "Upper_Keys"
      )
    )
  )

ggplot() +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA) +
  geom_rect(
    # data = break_df,
    data = tribble(
      ~xmin, ~xmax, ~cls,
      10, 32, 1,
      32, 34, 2,
      34, ceiling(max(df0$sal)), 3
      ) %>% mutate(cls = as.factor(cls)),
    aes(xmin = xmin, xmax = xmax, ymin = -0.06, ymax = -0.02, fill = cls),
    color = "grey40"
  ) +
  geom_step(data = ecdf_df, aes(x = x, y = ecdf), color = "black") +
  geom_point(data = ecdf_df, aes(x = x, y = ecdf), color = "black", alpha = 0.2) +
  # geom_vline(xintercept = brk_int$brks, color = "grey", linewidth = 0.4,
  geom_vline(xintercept = c(32, 34), color = "grey", linewidth = 0.4,
             linetype = "dashed") +
  labs(
    x = "Salinity",
    y = "Cumulative Proportion"
  ) +
  facet_wrap(~region) +
  # scale_fill_manual(values = viridisLite::viridis(nrow(break_df)), guide = "none") +
  scale_fill_manual(values = c("#457db4", "#39a585", "#cfcfcf"), guide = "none") +
  scale_x_continuous(breaks = scales::breaks_width(5), expand = expansion()) +
  scale_y_continuous(expand = expansion()) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.title.y       = element_text(angle = 90),
    axis.text.y        = element_text(angle = 90),
    strip.background = element_rect(fill = "white")
  )

if (FALSE) {
  ggsave(here(base_dir, "salinity_edcf_each_regions.jpeg"), width = 10, height = 5)
  shell.exec(base_dir)
}

# ---- ecdf by colored by month each region ---- #
palette_12 <- RColorBrewer::brewer.pal(12, "Paired") 

ecdf_df <-
  df0 %>%
  select("x" = sal, region, date) %>%
  arrange(x) %>%
  mutate(
    .by = region,
    .keep = "used",
    ecdf = seq_along(x) / length(x),
    region = factor(
      region,
      levels = c(
        "Ten_Thousand_Islands", "Everglades",
        "Lower_Keys", "Middle_Keys", "Upper_Keys"
      )
    ),
    # date = mdy(date),
    date = ymd(date),
    month = format(date, "%B"),
    month = factor(month, levels = month.name)
  )
test_plt <- 
ggplot() +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = NA) +
  geom_rect(
    # data = break_df,
    data = tribble(
      ~xmin, ~xmax, ~cls,
      10, 32, 1,
      32, 34, 2,
      34, ceiling(max(df0$sal)), 3
      ) %>% mutate(cls = as.factor(cls)),
    aes(xmin = xmin, xmax = xmax, ymin = -0.06, ymax = -0.02, fill = cls),
    color = "grey40"
  ) +
  geom_step(data = ecdf_df, aes(x = x, y = ecdf), color = "black") +
  # geom_point(data = ecdf_df, aes(x = x, y = ecdf), color = "black", alpha = 0.2) +
  geom_point(data = ecdf_df, aes(x = x, y = ecdf, color = month), alpha = 0.5) +
  # geom_vline(xintercept = brk_int$brks, color = "grey", linewidth = 0.4,
  geom_vline(xintercept = c(32, 34), color = "grey", linewidth = 0.4,
             linetype = "dashed") +
  labs(
    x = "Salinity",
    y = "Cumulative Proportion"
  ) +
  facet_wrap(~region) +
  scale_color_manual(values = palette_12) +
  # scale_fill_manual(values = viridisLite::viridis(nrow(break_df)), guide = "none") +
  scale_fill_manual(values = c("#457db4", "#39a585", "#cfcfcf"), guide = "none") +
  scale_x_continuous(breaks = scales::breaks_width(5), expand = expansion()) +
  scale_y_continuous(expand = expansion()) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.title.y       = element_text(angle = 90),
    axis.text.y        = element_text(angle = 90),
    strip.background = element_rect(fill = "white")
  )

plotly::ggplotly(test_plt)



```

```{r}
sfer_data <- 
  here("data", "raw") %>%
  dir_ls(regexp = "SFER") %>%
  read_csv(show_col_types = FALSE) %>%
  filter(datetime <= max_date) %T>% 
  print()


regions <- 
  sfer_data %>% 
  distinct(station, lat_dec, lon_dec) %>%
  summarise(
    .by = station,
    lat = mean(lat_dec),
    lon = mean(lon_dec)
  ) %>%
  filter(lon > -82.4 & lat < 26) 

regions %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = station))

station <-
  df0 %>%
  select(station, region, latitude, longitude) %>%
  summarise(
    .by = c(station, region),
    lat1 = mean(latitude),
    lon1 = mean(longitude)
  ) %>%
  distinct() %>%
  full_join(
    regions
  ) %>%
  filter(!is.na(lat)) %>%
  select(station, region)
  
shelf(patchwork)

(left_join(
  sfer_data,
  station
  ) %>%
  filter(!is.na(region))  %>%
  filter(!(chl <= 0 )) %>%
  mutate(
    sal_class =
      case_when(
        sal < 32 ~ "Low",
        between(sal, 32, 34) ~ "Moderate",
        sal >= 34 ~ "Marine"
      ),
    sal_class = factor(sal_class, levels = c("Marine", "Moderate", "Low")),
    sal       = if_else(sal < 0, 0.0001, sal),
    chl_log   = log10(chl),
    
     region = factor(
      region,
      levels = c(
        "Ten_Thousand_Islands", "Everglades",
        "Lower_Keys", "Middle_Keys", "Upper_Keys"
      )
    )
  ) %>%
  filter(!is.na(sal_class))  %T>% 
  print() %>%
    
  ggplot(aes(y = chl, x = region, color = sal_class)) +
  geom_boxplot(aes(fill = after_scale(prismatic::clr_lighten(color))), 
               box.color = "black", position = position_dodge2(preserve = "single")
               ) +
  geom_jitter(aes(fill = after_scale(prismatic::clr_darken(color))), 
              alpha = 0.5, position = position_jitterdodge(), show.legend = FALSE,) +
  scale_y_log10(
    breaks = 10^(-2:1),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_color_manual(values = rev(c("#457db4", "#39a585", "#cfcfcf"))) +
  annotation_logticks(side = "l") +
  coord_cartesian(ylim = c(NA, max(c(df0$chl, 20), na.rm = TRUE))) +
  theme_bw()) /

(df0 %>%
  filter(!(chl <= 0 )) %>%
  mutate(
    sal_class =
      case_when(
        sal < 32 ~ "Low",
        between(sal, 32, 34) ~ "Moderate",
        sal >= 34 ~ "Marine"
      ),
    sal_class = factor(sal_class, levels = c("Marine", "Moderate", "Low")),
    sal       = if_else(sal < 0, 0.0001, sal),
    chl_log   = log10(chl),
  ) %>%
  ggplot(aes(y = chl, x = region, color = sal_class)) +
  geom_boxplot(aes(fill = after_scale(prismatic::clr_lighten(color))), box.color = "black") +
  geom_jitter(aes(fill = after_scale(prismatic::clr_darken(color))), 
              alpha = 0.5, position = position_jitterdodge(), show.legend = FALSE,) +
  scale_y_log10(
    breaks = 10^(-2:1),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_color_manual(values = rev(c("#457db4", "#39a585", "#cfcfcf"))) +
  annotation_logticks(side = "l") +
  coord_cartesian(ylim = c(NA, max(c(df0$chl, 20)))) +
  theme_bw())
```
```{r}
sfer_summ_reg <- 
  left_join(
  sfer_data,
  station
  ) %>%
  filter(!is.na(region))  %>%
  filter(!(chl <= 0 )) %>%
  mutate(
    sal_class =
      case_when(
        sal < 32 ~ "Low",
        between(sal, 32, 34) ~ "Moderate",
        sal >= 34 ~ "Marine"
      ),
    sal_class = factor(sal_class, levels = c("Marine", "Moderate", "Low")),
    # sal_class = fct_na_level_to_value()
    sal       = if_else(sal < 0, 0.0001, sal),
    chl_log   = log10(chl),
    
     region = factor(
      region,
      levels = c(
        "Ten_Thousand_Islands", "Everglades",
        "Lower_Keys", "Middle_Keys", "Upper_Keys"
      )
    )
  ) %>%
  filter(!is.na(sal_class)) %>%
   
  summarise(
    .by = c(cruise_id, region, sal_class),
    chl = mean(chl, na.rm = TRUE)
  ) %T>% 
  print() %T>% 
  {
    count(., region, sal_class) %>%
    print()
  }

(sfer_summ_reg %>%
    
  ggplot(aes(y = chl, x = region, color = sal_class)) +
  geom_boxplot(aes(fill = after_scale(prismatic::clr_lighten(color))), 
               box.color = "black", position = position_dodge(1, preserve = "single")
               ) +
  geom_jitter(aes(fill = after_scale(prismatic::clr_darken(color))), 
              alpha = 0.5, position = position_jitterdodge(1), show.legend = FALSE,) +
  
  geom_point(
    data = tibble(x = c(1, 3), y = c(1,1)),
    aes(x = x, y = y), color = "red", inherit.aes = FALSE
    ) +
  geom_point(
    data = tibble(x = c(1, 3) + 0.33, y = c(1,1)),
    aes(x = x, y = y), color = "red", inherit.aes = FALSE
    ) +
  
  scale_y_log10(
    breaks = 10^(-2:1),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_color_manual(values = rev(c("#457db4", "#39a585", "#cfcfcf"))) +
  scale_x_discrete(drop = FALSE) +
  annotation_logticks(side = "l") +
  coord_cartesian(ylim = c(NA, max(c(df0$chl, 20)))) +
  theme_bw()
  
  ) /

(df0 %>%
  filter(!(chl <= 0 )) %>%
  mutate(
    sal_class =
      case_when(
        sal < 32 ~ "Low",
        between(sal, 32, 34) ~ "Moderate",
        sal >= 34 ~ "Marine"
      ),
    sal_class = factor(sal_class, levels = c("Marine", "Moderate", "Low")),
    sal       = if_else(sal < 0, 0.0001, sal),
    chl_log   = log10(chl),
  ) %>%
  summarise(
    .by = c(cruise_id, region, sal_class),
    chl = mean(chl, na.rm = TRUE)
  ) %>%
  ggplot(aes(y = chl, x = region, color = sal_class)) +
  geom_boxplot(aes(fill = after_scale(prismatic::clr_lighten(color))), box.color = "black",
              position = position_dodge(1, "single")) +
  geom_jitter(aes(fill = after_scale(prismatic::clr_darken(color))), 
              alpha = 0.5, position = position_jitterdodge(), show.legend = FALSE,) +
  scale_y_log10(
    breaks = 10^(-2:1),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_color_manual(values = rev(c("#457db4", "#39a585", "#cfcfcf"))) +
  annotation_logticks(side = "l") +
  coord_cartesian(ylim = c(NA, max(c(df0$chl, 20)))) +
  theme_bw())
```

```{r}
# df1 %>%
df0 %>%
   mutate(
    region = factor(
      region_label,
      levels = c(
        "Ten Thousand Islands", "Everglades",
        "Lower Keys", "Middle Keys", "Upper Keys"
      )
    ),
    sal_class =
      case_when(
        sal < 32 ~ "Low",
        between(sal, 32, 34) ~ "Moderate",
        sal >= 34 ~ "Marine"
      ),
     sal_class = factor(sal_class, levels = c("Marine", "Moderate", "Low"))
    ) %T>%
  {
    count(., cruise_id, region, sal_class) %>%
      mutate(
        sor_rec = if_else(str_detect(region, "(?i)keys"), "Recieve", "Source")
      ) %>%
      filter(!str_detect(sal_class, "Marine")) %>%
      summarise(
        .by = c(cruise_id, sor_rec, sal_class),
        n = sum(n, na.rm = TRUE)
      ) %>%
      pivot_wider(
        data         = .,
        id_cols      = c(cruise_id), # *optional* vector of unaffected columns,
        names_from   = c(sor_rec, sal_class), # category column(s) to pivot wide
        values_from  = c(n), # value column(s) that hold data for each category column
        names_sep    = "_",
        names_repair = janitor::make_clean_names
        ) %>%
      select(cruise_id, source_moderate, source_low, recieve_moderate, recieve_low) %>%
      mutate(
        source_moderate = source_moderate + source_low,
        recieve_moderate = recieve_moderate + recieve_low
      ) %>%
      filter(
        str_detect(
          cruise_id,
          "WS0603|WS1018|WS14335|WS16004|WS17282|WS18008|WS20342|WB22215|WS22281|WS23011")
      ) %>%
      print()
  } %>%
  count(region, sal_class) %>%
  mutate(
    .by    = region,
    tot    = sum(n),
    prop   = (n / tot),
    pct    = scales::label_percent(0.01)(prop),
    region = factor(
      region,
      levels = c(
        "Ten Thousand Islands", "Everglades",
        "Lower Keys", "Middle Keys", "Upper Keys"
      )
    ),
  ) %T>% 
  print() %T>% 
  {
      text_annotate <<- filter(., !str_detect(sal_class, "Marine")) %>%
      summarise(
        .by = region, 
        prop_sum = sum(prop),
        pct = scales::label_percent(0.01)(prop_sum)
        ) %T>%
      print()
  } %>%
  ggplot(aes(x = region, y = prop, fill = sal_class)) +
  geom_col() +
  geom_text(
    data = text_annotate,
    aes(x = region, y = prop_sum + 0.025, label = pct),
    inherit.aes = FALSE,
    vjust = 0.5
  ) +
  labs(
    x = NULL,
    y = "Percent (%)",
    fill = "Salinity Class"
  ) +
  scale_fill_manual(drop = FALSE, values = rev(c("#457db4", "#39a585", "#cfcfcf"))) +
  scale_y_continuous(expand = expansion(), labels = scales::label_percent(1)) +
  theme_bw() +
  theme(
    legend.position = "top"
  )
  
```


```{r}
sfer_summ_reg %>%
  ggplot(aes(y = chl, x = region, color = sal_class)) +
  geom_boxplot(aes(fill = after_scale(prismatic::clr_lighten(color)),
                   group = interaction(region, sal_class)),
    box.color = "black", position = position_dodge(1, preserve = "single")
  ) +
  geom_point(
    aes(fill = after_scale(prismatic::clr_darken(color)),
        group = interaction(region, sal_class)),
    alpha = 0.5,
    position = position_jitter(
      width = 0.03,
      height = 0
    ), 
    show.legend = FALSE,
  ) +
  scale_y_log10(
    breaks = 10^(-2:1),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_color_manual(drop = FALSE, values = rev(c("#457db4", "#39a585", "#cfcfcf"))) +
  scale_x_discrete(drop = FALSE) +
  scale_fill_discrete(drop = FALSE) +
  annotation_logticks(side = "l") +
  coord_cartesian(ylim = c(NA, max(c(df0$chl, 20)))) +
  theme_bw()

library(dplyr)
library(ggplot2)

region_lvls <- levels(factor(sfer_summ_reg$region))
sal_lvls <- levels(factor(sfer_summ_reg$sal_class))

df1 <- 
  sfer_summ_reg %>%
  mutate(
    region     = factor(region, levels = region_lvls),
    sal_class  = factor(sal_class, levels = sal_lvls),
    region_num = as.numeric(region),
    sal_num    = as.numeric(sal_class),
    x          = region_num + (sal_num - 2) * 0.33,
    # region_num = as.character(region),
  )

ggplot(df1, aes(y = chl)) +
  geom_boxplot(
    mapping = aes(x = region_num, group = interaction(region, sal_class), color = sal_class),
    position = position_dodge(width = 1, preserve = "single"),
    # aes(fill = after_scale(prismatic::clr_lighten(color))),
    box.color = "black",
    outlier.shape = NA
  ) +
  geom_point(
    mapping = aes(x = x, color = sal_class, fill = after_scale(prismatic::clr_darken(color))),
    alpha = 0.5,
    position = position_jitter(width = 0.03, height = 0),
    show.legend = FALSE
  ) +
  scale_x_continuous(breaks = seq_along(region_lvls), labels = region_lvls) +
  scale_color_manual(drop = FALSE, values = rev(c("#457db4", "#39a585", "#cfcfcf"))) +
  scale_y_log10(
    breaks = 10^(-2:1),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  annotation_logticks(side = "l") +
  coord_cartesian(ylim = c(NA, max(c(df0$chl, 20)))) +
  theme_bw()
```



# ============================================================================ #
# Functions:


## Extract Cruise and Convex Hull Information

Outputs:
  pts_utm          - points of data as sf object
  support_poly_utm - convex hull around the data points
  main_title       - title of plot in format: 
                       `<cruise id> - Surface (0-5 m mean)`
                       or `No rows matched cruise_id`
  sub_title        - subtitle of plot in format: 
                      `<date start> to <date end> | IDW + (UTM grid)`
  chl_panel_title  - One of three:
                       `Chlorophyll concentration`
                       `Chlorophyll fluorescence`
                       `Chlorophyll`


```{r cruise-data}
cruise_extract <- function(cruise_id, dat) {
  
  # ---- main title text ---- #
  main_title <- paste0(cruise_id, " — Surface (0–5 m mean)")

  df <-
    dat %>%
    filter(
      str_detect(cruise_file, cruise_id), # select cruise
      !is.na(station),
      station != "",
      !is.na(latitude),
      !is.na(longitude)
    )

  if (nrow(df) == 0) stop("No rows matched cruise_id: ", cruise_id)

  # ---- subtitle text ---- #
  date_vec <- df$date_raw
  date_min <- min(date_vec, na.rm = TRUE)
  date_max <- max(date_vec, na.rm = TRUE)
  sub_title <-
    if (all(!is.na(c(date_min, date_max)))) {
      paste(format(date_min, "%Y-%m-%d"), "to", format(date_max, "%Y-%m-%d"))
    } else {
      "Date range: NA"
    }

  sub_title <- paste0(sub_title, " | IDW + (UTM grid)")

  # ---- chlorophyll source MODE for THIS cruise (panel title only) ---- #
  chl_src_tbl <- df %>%
    filter(!is.na(chl_source), chl_source != "") %>%
    count(chl_source, name = "n") %>%
    arrange(desc(n))

  chl_src_mode <- if (nrow(chl_src_tbl) == 0) {
    "unknown"
  } else {
    chl_src_tbl$chl_source[1]
  }
  
  chl_panel_title <- 
    if (identical(chl_src_mode, "chlorophyll_concentration")) {
      "Chlorophyll concentration"
      } else if (identical(chl_src_mode, "chlorophyll_fluorescence")) {
        "Chlorophyll fluorescence"
        } else {
          "Chlorophyll"
          }
  

  # ---- station means (ONLY aggregation; no smoothing of source data) ---- #
  df_st <-
    df %>%
    summarise(
      .by = station,
      across(
        c(latitude, longitude, sal, temp, oxy, chl),
        \(.x) mean(.x, na.rm = TRUE)
      )
    ) %>%
    filter(!is.na(latitude), !is.na(longitude))
  
  # convert dataframe of stations to sf object and project
  pts_utm <- 
    st_as_sf(df_st, coords = c("longitude","latitude"), crs = 4326, remove = FALSE)  %>%
    st_transform(utm_epsg) %>%
    mutate(
      longitude = st_coordinates(.)[,1],
      latitude = st_coordinates(.)[,2]
      ) 
  
  # ---- TIGHT HARD SUPPORT MASK — computed once ---- #  
  # create a smaller, irregular bounding box based on convex hull or buffer (<3)
  if (nrow(pts_utm) >= 3) {
    # convex hull, needs at least 3 points
    support_poly_utm <- 
      pts_utm %>%
      st_union() %>%
      st_convex_hull() %>%
      st_buffer(8000)   # 8 km buffer
  } else {
    # buffer around each point
    support_poly_utm <- 
      pts_utm %>%
      st_buffer(30000) %>%
      st_union()
  }

  return(
    list(
      "pts_utm"          = pts_utm,
      "support_poly_utm" = support_poly_utm,
      "main_title"       = main_title,
      "sub_title"        = sub_title,
      "chl_panel_title"  = chl_panel_title
      )
  )
  # ---- end of function cruise_extract ---- #
}
```

## Create Blank Raster Grid to Add IDW values

Output:
  grid_rast - raster version
  grid_sf   - sf object version

```{r blank-grid}
grid_blank <- function(.support_poly, utm_epsg, grid_res_m = 1000, buffer_m = 10000) {
  # create blank raster grid with equal grid numbers
  grid_rast <- 
    # add 10 km boundary around convex hull
    st_buffer(.support_poly, dist = buffer_m) %>%
    st_bbox() %>%
    {
      terra::rast(
        xmin       = .[["xmin"]],
        xmax       = .[["xmax"]],
        ymin       = .[["ymin"]],
        ymax       = .[["ymax"]],
        resolution = grid_res_m,
        crs        = paste0("EPSG:", utm_epsg)
      )
    } 
  
  # convert projected raster to sf object with selected resolution
  # this will be the sf object to map predicted values from interpolation (IDW)  
  grid_sf <- 
    as.data.frame(grid_rast, xy = TRUE, cells = TRUE, na.rm = FALSE) %>%
    st_as_sf(coords = c("x","y"), crs = utm_epsg)
  
  return(
    list(
      "grid_rast" = grid_rast,
      "grid_sf"   = grid_sf
    )
  )
  # ---- end of function grid_blank ---- #
}
```




## Edge of Convex Hull Settings (Opacity of Points)

This only softens the IDW convex hull near the edge of the mask (boundary). The
edge of the boundary is set with a default of 5000 m from the edge to taper off

alpha refers to alpha of ggplot meaning opacity

alpha_utm - raster image with alpha levels (0 - 1) to show opacity

```{r edge}
edge_opacity <- function(
    .grid_rast,
    .support_poly_utm,
    .edge_feather_m = 5000 # try 4000–7000 depending on taste
    ) {
  
  # convert convex hull to linestring
  support_boundary_line <- st_boundary(.support_poly_utm) %>%
    terra::vect()
  
  # calculate distance from convex hull (as linestring) then mask to within convex hull
  dist_to_edge_utm <-
    terra::distance(.grid_rast, support_boundary_line, names = "alpha") %>%
    terra::mask(terra::vect(.support_poly_utm))

  # divides distances by edge buffer then sets upper bounds to 1 by converting all values
  # above 1 to 1 and leaves all values below as is
  alpha_utm <-
    terra::clamp(dist_to_edge_utm / .edge_feather_m, lower = 0, upper = 1, values = TRUE)

  # ---- end of function edge_opacity ---- #
}
```


## IDW Calculation and Figure Template


```{r idw-panel}
# make_panel <- function(
#   col_name, 
#   var_title, 
#   legend_units, 
#   .legend_limits    = NULL,
#   .df_st            = pts_utm,  
#   .alpha_utm        = alpha_utm,
#   .smooth_kernel    = smooth_kernel,
#   .grid_sf          = grid_sf,
#   .grid_rast        = grid_rast,
#   .utm_epsg         = utm_epsg,
#   .support_poly_utm = support_poly_utm,
#   .florida_sf       = florida_sf,
#   .bb_utm           = bb_utm,
#   idw_opts          = idw_options
#   ) {
# 
#   # keep columns needed for IDW interpolation and plotting, converts selected
#   # parameter name to val, removes NA 
#   dd <- 
#     .df_st %>%
#     mutate(
#       .keep = "none",
#       station,
#       latitude,
#       longitude,
#       val = .data[[col_name]],
#       geometry
#     ) %>%
#     filter(!is.na(val))
#   
#   # ---- skip plotting if less than 5 points ---- #
#   if (nrow(dd) < 5) {
#     return(
#       ggplot() +
#         theme_void() +
#         labs(title = paste0(var_title, " (too few stations)")) +
#         theme(plot.title = element_text(face = "bold", size = 12))
#     )
#   }
# 
#   # ---- IDW interpolation as sf object ---- #
#   # uses options set before
#   idw_sf_utm <- 
#     gstat::idw(
#       formula   = val ~ 1,
#       locations = dd,
#       newdata   = .grid_sf,
#       idp       = idw_opts$idp_pow,
#       maxdist   = idw_opts$maxdist_m,
#       nmin      = idw_opts$nmin_nb,
#       nmax      = idw_opts$nmax_nb
#     )
# 
#   # replace empty grid with IDW values
#   terra::values(.grid_rast) <- as.numeric(idw_sf_utm$var1.pred)
# 
# 
#   # ---- LIGHT DISPLAY SMOOTHING ONLY ---- #
#   r_pred_water <-
#     # mask data outside convex hull
#     terra::mask(.grid_rast, terra::vect(.support_poly_utm)) %>%
#     
#     # smooths over values with weighting of 3x3 kernel
#     terra::focal(
#       w         = .smooth_kernel,
#       fun       = mean,
#       na.policy = "omit",
#       na.rm     = TRUE,
#       fillvalue = NA
#     ) %>%
#     
#     # combine IDW and alpha (opacity)
#     c(., alpha_utm) %>%
#     
#     # mask again so smoothing does not bleed outside supported footprint
#     # of convex hull
#     terra::mask(terra::vect(.support_poly_utm)) %>% # <-- may not be necessary
#     
#     # mask land mass
#     terra::mask(terra::vect(.florida_sf), inverse = TRUE) 
#     
#   # convert into data frame
#   plot_df <- 
#     r_pred_water %>%
#     as.data.frame(xy = TRUE, na.rm = FALSE) %>%
#     rename("pred" = focal_mean)  %>%
#     mutate(alpha = ifelse(is.na(alpha), 0, alpha)) %>%
#     filter(!is.na(pred), alpha > 0)
#   
#   
#   # ---- contours ---- #
#   contour_sf <- NULL
# 
#   rng <- range(plot_df$pred, na.rm = TRUE)
# 
#   if (all(is.finite(rng)) && diff(rng) > 0) {
#     brks <- pretty(rng, n = 6)
#     brks <- brks[brks > rng[1] & brks < rng[2]]
#     
#     if (length(brks) > 0) {
#       contour_sf <- try(
#         terra::as.contour(r_pred_water, levels = brks),
#         silent = TRUE
#       )
#       # if error, convert to sf
#       if (!inherits(contour_sf, "try-error")) {
#         contour_sf <- try(sf::st_as_sf(contour_sf), silent = TRUE)
#       }
#       
#       # if error on conversion, set to NULL
#       if (inherits(contour_sf, "try-error")) contour_sf <- NULL
#     }
#   }
# 
#   if (is.null(.legend_limits)) .legend_limits <- c(NA, NA)
#   
#   # ---- plot ---- #
#   p <- 
#     ggplot() +
#     annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "#cfe8f3") +
#     geom_raster(data = plot_df, aes(x = x, y = y, fill = pred, alpha = alpha)) +
#     scale_fill_viridis_c(
#       name = legend_units, option = "C", direction = 1, na.value = NA, 
#       limits = .legend_limits, oob = scales::squish) +
#     scale_alpha_identity() +
#     geom_sf(data = .florida_sf, fill = "#e8e6df", color = "grey35", linewidth = 0.55) +
# 
#   # add contours if possible
#   {
#     if (!is.null(contour_sf) && nrow(contour_sf) > 0) {
#       geom_sf(
#         data      = contour_sf,
#         color     = "white",
#         linewidth = 0.22,
#         alpha     = 0.45
#       )
#       }
#     } +
#     geom_sf(data = dd, color = "black", size = 0.55, alpha = 0.95) +
#     ggrepel::geom_text_repel(
#       data = dd,
#       aes(x = longitude, y = latitude, label = station),
#       seed               = 1,
#       size               = 2.0,
#       color              = "black",
#       box.padding        = 0.18,
#       point.padding      = 0.10,
#       min.segment.length = 0,
#       segment.color      = "grey35",
#       segment.size       = 0.30,
#       segment.curvature  = 0.0,
#       arrow              = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
#       max.overlaps       = Inf
#     ) +
#     annotation_north_arrow(
#       location    = "tr",
#       which_north = "true",
#       pad_x       = unit(0.10, "cm"),
#       pad_y       = unit(0.10, "cm"),
#       height      = unit(0.85, "cm"),
#       width       = unit(0.85, "cm"),
#       style       = north_arrow_orienteering(
#         line_col = "grey20",
#         fill     = c("grey20", "white"),
#         text_col = "grey20"
#       )
#     ) +
#     annotation_scale(
#       location   = "bl",
#       width_hint = 0.22,
#       pad_x      = unit(0.12, "cm"),
#       pad_y      = unit(0.12, "cm"),
#       text_cex   = 0.55,
#       line_width = 0.6
#     ) +
#     coord_sf(xlim = .bb_utm[c(1,3)], ylim = .bb_utm[c(2,4)], expand = FALSE) +
#     labs(title = var_title, x = NULL, y = NULL) +
#     theme_bw() +
#     theme(
#       panel.background = element_rect(fill = "#cfe8f3"),
#       panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
#       panel.grid.minor = element_blank(),
#       legend.position  = "right",
#       plot.title       = element_text(face = "bold", size = 12)
#     )
# 
#   p
#   
#   # ---- end of function make_panel ---- #
# }
# 
# old way of idw and plots
# now separates into two distinct functions
# make panels
# p_sal <- make_panel("sal", "Salinity", "PSU", legends_lim[[1]])
# p_tmp <- make_panel("temp", "Temperature", expression(paste(degree, "C")), legends_lim[[2]])
# p_o2  <- make_panel("oxy", "Oxygen", expression(paste("mL ", L^{-1})), legends_lim[[3]])
# p_chl <- make_panel("chl", chl_panel_title, expression(paste("mg ", m^{-3})), legends_lim[[4]])
```


### IDW Calculation

```{r idw-func}
# ---- inverse distance weighting function ---- #
idw_func <- function(
  col_name,
  .df_st            = pts_utm,  
  .alpha_utm        = alpha_utm,
  .smooth_kernel    = smooth_kernel,
  .grid_sf          = grid_sf,
  .grid_rast        = grid_rast,
  .support_poly_utm = support_poly_utm,
  .florida_sf       = florida_sf,
  idw_opts          = idw_options
  ) {

  # keep columns needed for IDW interpolation and plotting, converts selected
  # parameter name to val, removes NA 
  dd <- 
    .df_st %>%
    mutate(
      .keep = "none",
      station,
      latitude,
      longitude,
      val = .data[[col_name]],
      geometry
    ) %>%
    filter(!is.na(val))
  
  # ---- skip plotting if less than 5 points ---- #
  if (nrow(dd) < 5) {
    return(
      list(
        "dd"           = dd,
        "r_pred_water" = NULL,
        "plot_df"      = NULL
      )
    )
  }

  # ---- IDW interpolation as sf object ---- #
  # uses options set before
  idw_sf_utm <- 
    gstat::idw(
      formula   = val ~ 1,
      locations = dd,
      newdata   = .grid_sf,
      idp       = idw_opts$idp_pow,
      maxdist   = idw_opts$maxdist_m,
      nmin      = idw_opts$nmin_nb,
      nmax      = idw_opts$nmax_nb
    )

  # replace empty grid with IDW values
  terra::values(.grid_rast) <- as.numeric(idw_sf_utm$var1.pred)

  # ---- LIGHT DISPLAY SMOOTHING ONLY ---- #
  r_pred_water <-
    # mask data outside convex hull
    terra::mask(.grid_rast, terra::vect(.support_poly_utm)) %>%
    
    # smooths over values with weighting of 3x3 kernel
    terra::focal(
      w         = .smooth_kernel,
      fun       = mean,
      na.policy = "omit",
      na.rm     = TRUE,
      fillvalue = NA
    ) %>%
    
    # combine IDW and alpha (opacity)
    c(., .alpha_utm) %>%
    
    # mask again so smoothing does not bleed outside supported footprint
    # of convex hull
    terra::mask(terra::vect(.support_poly_utm)) %>% # <-- may not be necessary
    
    # mask land mass
    terra::mask(terra::vect(.florida_sf), inverse = TRUE) 
  
  # convert into data frame
  plot_df <- 
    r_pred_water %>%
    as.data.frame(xy = TRUE, na.rm = FALSE) %>%
    rename("pred" = focal_mean)  %>%
    mutate(alpha = ifelse(is.na(alpha), 0, alpha)) %>%
    filter(!is.na(pred), alpha > 0)
  
  
  return(
    list(
      "dd"           = dd,
      "r_pred_water" = r_pred_water,
      "plot_df"      = plot_df
    )
  )
  # ---- end of function idw_func ---- #
}
```


### Figure Template

```{r plot-func}
# ---- plot function ---- #
make_panel <- function(
    var_title, 
    .dd,
    .r_pred_water  = r_pred_water,
    .plot_df       = plot_df,
    legend_units, 
    .legend_limits = NULL, 
    .florida_sf, 
    .bb_utm
    ) {
  # ---- skip plotting if less than 5 points ---- #
  if (nrow(.dd) < 5) {
    return(
      ggplot() +
        theme_void() +
        labs(title = paste0(var_title, " (too few stations)")) +
        theme(plot.title = element_text(face = "bold", size = 12))
    )
  }

  # ---- contours ---- #
  contour_sf <- NULL

  rng <- range(.plot_df$pred, na.rm = TRUE)

  if (all(is.finite(rng)) && diff(rng) > 0) {
    brks <- pretty(rng, n = 6)
    brks <- brks[brks > rng[1] & brks < rng[2]]
    
    if (length(brks) > 0) {
      contour_sf <- try(
        terra::as.contour(.r_pred_water, levels = brks),
        silent = TRUE
      )
      # if error, convert to sf
      if (!inherits(contour_sf, "try-error")) {
        contour_sf <- try(sf::st_as_sf(contour_sf), silent = TRUE)
      }
      
      # if error on conversion, set to NULL
      if (inherits(contour_sf, "try-error")) contour_sf <- NULL
    }
  }

  if (is.null(.legend_limits)) .legend_limits <- c(NA, NA)
  
  # ---- plot ---- #
  p <- 
    ggplot() +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "#cfe8f3") +
    geom_raster(data = .plot_df, aes(x = x, y = y, fill = pred, alpha = alpha)) +
    geom_sf(data = .florida_sf, fill = "#e8e6df", color = "grey35", linewidth = 0.55) +
    scale_fill_viridis_c(
      name = legend_units, option = "C", direction = 1, na.value = NA, 
      limits = .legend_limits, oob = scales::squish
      ) +
    scale_alpha_identity() +

  # add contours if possible
  {
    if (!is.null(contour_sf) && nrow(contour_sf) > 0) {
      geom_sf(
        data      = contour_sf,
        color     = "white",
        linewidth = 0.22,
        alpha     = 0.45
      )
      }
    } +
    
    # plot data points and labels
    geom_sf(data = .dd, color = "black", size = 0.55, alpha = 0.95) +
    ggrepel::geom_text_repel(
      data = .dd,
      aes(x = longitude, y = latitude, label = station),
      seed               = 1,
      size               = 2.0,
      color              = "black",
      box.padding        = 0.18,
      point.padding      = 0.10,
      min.segment.length = 0,
      segment.color      = "grey35",
      segment.size       = 0.30,
      segment.curvature  = 0.0,
      arrow              = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
      max.overlaps       = Inf
    ) +
    
    # add arrow and scale bar
    annotation_north_arrow(
      location    = "tr",
      which_north = "true",
      pad_x       = unit(0.10, "cm"),
      pad_y       = unit(0.10, "cm"),
      height      = unit(0.85, "cm"),
      width       = unit(0.85, "cm"),
      style       = north_arrow_orienteering(
        line_col = "grey20",
        fill     = c("grey20", "white"),
        text_col = "grey20"
      )
    ) +
    annotation_scale(
      location   = "bl",
      width_hint = 0.22,
      pad_x      = unit(0.12, "cm"),
      pad_y      = unit(0.12, "cm"),
      text_cex   = 0.55,
      line_width = 0.6
    ) +
    
    # limits
    coord_sf(xlim = .bb_utm[c(1,3)], ylim = .bb_utm[c(2,4)], expand = FALSE) +
    
    # labels
    labs(title = var_title, x = NULL, y = NULL) +
    theme_bw() +
    theme(
      panel.background = element_rect(fill = "#cfe8f3"),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      legend.position  = "right",
      plot.title       = element_text(face = "bold", size = 12)
    )

  p
  
  # ---- end of function make_panel ---- #
}


# plot the data points, raster grid and convex hull
plot_convex <- function(
  .grid_sf,
  .florida_sf,
  .support_poly_utm,
  .pts_utm
  ) {
  ggplot() +
  geom_sf(data = grid_sf, color = "#500", alpha = 0.5) + # base grid for interpolation
  geom_sf(data = florida_sf) + # florida map
  geom_sf(
    data = .support_poly_utm,
    color = "blue", fill = "lightblue", alpha = 0.5
  ) + # convex hull
  geom_sf(data = pts_utm) + # data points used
  coord_sf(expand = FALSE) +
  labs(x = NULL, y = NULL) +
  ggrepel::geom_text_repel(
    data = pts_utm,
    aes(x = longitude, y = latitude, label = station),
    seed               = 1,
    size               = 2.0,
    color              = "black",
    box.padding        = 0.18,
    point.padding      = 0.10,
    min.segment.length = 0,
    segment.color      = "grey35",
    segment.size       = 0.30,
    segment.curvature  = 0.0,
    arrow              = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
    max.overlaps       = Inf
  )
  
  # ---- end of function plot_convex ---- #
}
```



## Save Plots

```{r save-func}
save_maps <- function(plt, .map_dir, .cruise_id, verbose = TRUE) {
  
  png_out <- here(.map_dir, "png")
  svg_out <- here(.map_dir, "svg")
  
  dir_create(png_out)
  dir_create(svg_out)
  
  # file names
  png_out  <- here(png_out, paste0(.cruise_id, "_0_5m_4vars_softedge_nav.png"))
  tiff_out <- here(svg_out, paste0(.cruise_id, "_0_5m_4vars_softedge_nav.tiff"))

  # save
  ggsave(png_out, plot = plt, width = 12.5, height = 9.5, dpi = 600)
  ggsave(tiff_out, plot = plt, width = 12.5, height = 9.5, dpi = 600, compression = "lzw")

  if (!verbose) return(invisible())
  cat("\nOutput folder:\n", .map_dir, "\n\n")
  cat("\nSaved PNG:\n", basename(png_out), "\nexists =", file_exists(png_out), "\n")
  cat("Saved TIFF:\n", basename(tiff_out), "\nexists =", file_exists(tiff_out), "\n")
  print(list.files(.map_dir))
  print(basename(dir_ls(.map_dir)))
  # ---- end of function save_maps ---- #
}


```

# ============================================================================ #
# Cruise Processing

```{r process}
# legend limits set for each variable
legends_lim <- list(
  sal_limits  = c(29, 37),
  temp_limits = c(18, 32),
  oxy_limits  = c(1, 5),
  chl_limits  = c(0, 5)
)
cruise_id <- cruise_ids[40]
# for (cruise_id in cruise_ids) {
for (cruise_id in cruise_ids[-c(1:48)]) {
  
  # ========================================================================== #
  # ---- Select Cruise Data ---- #
  # ========================================================================== #
  cat(paste("Cruise ID:", cruise_id))

  # extract cruise data
  cruise_temp      <- cruise_extract(cruise_id, dat = df0)
  pts_utm          <- cruise_temp$pts_utm
  support_poly_utm <- cruise_temp$support_poly_utm
  main_title       <- cruise_temp$main_title
  sub_title        <- cruise_temp$sub_title
  chl_panel_title  <- cruise_temp$chl_panel_title

  # create blank grid
  grid_temp <- grid_blank(support_poly_utm, utm_epsg)
  grid_rast <- grid_temp$grid_rast
  grid_sf   <- grid_temp$grid_sf

  # show locations, convex hull in relation to Florida
  # print(plot_convex(grid_sf, florida_sf, support_poly_utm, pts_utm))

  # create edge opacity
  alpha_utm <- edge_opacity(grid_rast, support_poly_utm)

  # print(
  #   ggplot() +
  #   tidyterra::geom_spatraster(data = alpha_utm) +
  #   scale_fill_viridis()
  #   )

  
  # ========================================================================== #
  # ---- IDW and Plotting ---- #
  # ========================================================================== #
  
  # ---- salinity ---- #
  idw_sal <- 
    idw_func(
      col_name          = "sal",
      .df_st            = pts_utm,  
      .alpha_utm        = alpha_utm,
      .smooth_kernel    = smooth_kernel,
      .grid_sf          = grid_sf,
      .grid_rast        = grid_rast,
      .support_poly_utm = support_poly_utm,
      .florida_sf       = florida_sf,
      idw_opts          = idw_options
    )
  
  p_sal <- 
    make_panel(
      var_title      = "Salinity", 
      legend_units   = "PSU", 
      .legend_limits = legends_lim[[1]], 
      .dd            = idw_sal$dd,
      .r_pred_water  = idw_sal$r_pred_water,
      .plot_df       = idw_sal$plot_df,
      .florida_sf    = florida_sf, 
      .bb_utm        = bb_utm
    )
  
  # ---- temperature ---- #
  idw_temp <- 
    idw_func(
      col_name          = "temp",
      .df_st            = pts_utm,  
      .alpha_utm        = alpha_utm,
      .smooth_kernel    = smooth_kernel,
      .grid_sf          = grid_sf,
      .grid_rast        = grid_rast,
      .support_poly_utm = support_poly_utm,
      .florida_sf       = florida_sf,
      idw_opts          = idw_options
    )
  
  p_tmp <- 
    make_panel(
      var_title      = "Temperature", 
      legend_units   = expression(paste(degree, "C")),
      .legend_limits = legends_lim[[2]], 
      .dd            = idw_temp$dd,
      .r_pred_water  = idw_temp$r_pred_water,
      .plot_df       = idw_temp$plot_df,
      .florida_sf    = florida_sf, 
      .bb_utm        = bb_utm
    )
  
  # ---- oxygen ---- #
  idw_oxy <- 
    idw_func(
      col_name = "oxy",
      .df_st            = pts_utm,  
      .alpha_utm        = alpha_utm,
      .smooth_kernel    = smooth_kernel,
      .grid_sf          = grid_sf,
      .grid_rast        = grid_rast,
      .support_poly_utm = support_poly_utm,
      .florida_sf       = florida_sf,
      idw_opts          = idw_options
    )
  
  p_o2 <- 
    make_panel(
      var_title      = "Oxygen", 
      legend_units   = expression(paste("mL ", L^{-1})),
      .legend_limits = legends_lim[[3]], 
      .dd            = idw_oxy$dd,
      .r_pred_water  = idw_oxy$r_pred_water,
      .plot_df       = idw_oxy$plot_df,
      .florida_sf    = florida_sf, 
      .bb_utm        = bb_utm
    )
  
  
  # ---- chlorophyll-a ---- #
  idw_chl <- 
    idw_func(
      col_name          = "chl",
      .df_st            = pts_utm,  
      .alpha_utm        = alpha_utm,
      .smooth_kernel    = smooth_kernel,
      .grid_sf          = grid_sf,
      .grid_rast        = grid_rast,
      .support_poly_utm = support_poly_utm,
      .florida_sf       = florida_sf,
      idw_opts          = idw_options
    )
  
  p_chl <- 
    make_panel(
      var_title      = chl_panel_title, 
      legend_units   = expression(paste("mg ", m^{-3})),
      .legend_limits = legends_lim[[4]], 
      .dd            = idw_chl$dd,
      .r_pred_water  = idw_chl$r_pred_water,
      .plot_df       = idw_chl$plot_df,
      .florida_sf    = florida_sf, 
      .bb_utm        = bb_utm
    )
  
  # ---- combine panels ---- #
  final_plot <-
    (p_sal + p_tmp) / (p_o2 + p_chl) +
    plot_annotation(
      title = main_title,
      subtitle = sub_title
    ) +
    plot_annotation(
      theme = theme(
        plot.title    = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 11)
      )
    )

  print(final_plot)
  
  # ---- save plot ---- #
  save_maps(final_plot, map_dir, cruise_id)
}
```




# FOCUS RANGES (pattern emphasis)
```{r}
sal_limits  <- c(28, 35)
oxy_limits  <- c(1,  5)
temp_limits <- c(18, 30)
chl_limits  <- c(0,  5)

cat("\nFOCUS color limits:\n")
cat("Salinity:    ", sal_limits[1],  " to ", sal_limits[2],  "\n", sep = "")
cat("Oxygen:      ", oxy_limits[1],  " to ", oxy_limits[2],  "\n", sep = "")
cat("Temperature: ", temp_limits[1], " to ", temp_limits[2], "\n", sep = "")
cat("Chlorophyll: ", chl_limits[1],  " to ", chl_limits[2],  "\n\n", sep = "")
```

