---
title: "Surface_maps"
author: "Mostafa Soliman"
date: "2026-03-06"
output: html_document
---

```{r}
# =========================================================================== #
# ONE cruise: 4-panel Surface Maps (0–5 m mean)
# Variables: Salinity, Temperature, Oxygen, Chlorophyll (CHL_0_5m + chl_source)
# Method: IDW (UTM grid) + LIGHT DISPLAY SMOOTHING

# Edge: TIGHT HARD SUPPORT MASK + SOFT FEATHERED DISPLAY EDGE
# Style: subtle contour lines + north arrow + scale bar

# Title:  "WS0603 — Surface (0–5 m mean)"
# Subtitle: "YYYY-MM-DD to YYYY-MM-DD
# CHL panel title: "Chlorophyll fluorescence" OR "Chlorophyll concentration"
# Units ONLY on colorbar title

# =========================================================================== #

if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional packages, if from GitHub, include username
  sf, ggspatial, terra, geodata, gstat, viridis, ggrepel, grid, 
  patchwork # combine panels
)

conflicts_prefer(
  dplyr::filter,
  dplyr::select
  )

# library(dplyr)
# library(readr)
# library(sf)
# library(ggplot2)
# library(ggspatial)
# library(stringr)
# library(terra)
# library(geodata)
# library(gstat)
# library(viridis)
# library(ggrepel)
# library(grid)      
# library(patchwork) # combine panels
# library(lubridate)
```

# 0) PATHS + CRUISE PICK

```{r}
base_dir <- here("data", "ctd_maps")
dir_create(base_dir)

ts_dir <- here(base_dir, "Time_series_new")
dir_create(ts_dir)

# file with all ctd casts
in_file <- 
  ts_dir  %>%
  dir_ls(regexp = "CTD_surface_0_5m_ALL")

stopifnot(length(in_file) != 0)

map_dir <- here(ts_dir, "Figures_freshening", "SurfaceMaps_0_5m_4vars_NOHALO")
dir_create(map_dir)

# base_dir <- "C:/Users/mostafaawada/OneDrive - University of South Florida/Desktop/CTD_data_codes/data/cleaned_ctd_data_v02/cleaned_ctd_data_v02/merged_by_cruise"
# ts_dir   <- file.path(base_dir, "Time_series_new")

# in_file  <- file.path(ts_dir, "CTD_surface_0_5m_ALL_casts_withRegionShelf_ALLPARAMS.csv")
# stopifnot(file.exists(in_file))


# map_dir <- file.path(ts_dir, "Figures_freshening", "SurfaceMaps_0_5m_4vars_NOHALO")
# dir.create(map_dir, showWarnings = FALSE, recursive = TRUE)

# name of cruise in cruise file column
cruise_pick <- "WS0603.csv"   # <--- CHANGE THIS (must match EXACT cruise_file values)
cruise_id <-
  str_remove(cruise_pick, "\\.csv$") %>%
  str_replace_all("[^A-Za-z0-9_\\-]+", "_")

cruise_id   <- "SV18067"   # <--- CHANGE THIS (must match EXACT cruise_file values)
```

# 1) READ + COLUMN DETECTION

```{r}
df0 <- read_csv(in_file, show_col_types = FALSE) %T>% 
  print()

# check missing columns in data file
need_base <- c("cruise_file", "station", "latitude", "longitude")
miss_base <- setdiff(need_base, names(df0))
if (length(miss_base) > 0) stop("Missing base columns: ", paste(miss_base, collapse = ", "))


# of the list of columns, pick the first one
pick_first_existing <- function(candidates, col_names) {
  hit <- candidates[candidates %in% col_names]
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

var_names <- list(
  "sal"  = c("S_0_5m","sal_0_5m","mean_sal_psu","salinity"),
  "temp" = c("T_0_5m","temp_0_5m","mean_temp_C","temperature","temp_C"),
  "o2"   = c("O2_0_5m","oxy_0_5m","mean_oxy","oxygen","DO"),
  "date" = c("date","Date","cast_date","datetime","time","timestamp")
)

col_sal  <- pick_first_existing(var_names$sal, names(df0))
col_tmp  <- pick_first_existing(var_names$temp, names(df0))
col_o2   <- pick_first_existing(var_names$o2, names(df0))
col_date <- pick_first_existing(var_names$date, names(df0))

# col_sal <- pick_first_existing(, names(df0))
# col_tmp <- pick_first_existing(c("T_0_5m","temp_0_5m","mean_temp_C","temperature","temp_C"), names(df0))
# col_o2  <- pick_first_existing(c("O2_0_5m","oxy_0_5m","mean_oxy","oxygen","DO"), names(df0))

# confirmed by you:
col_chl     <- "CHL_0_5m"
col_chl_src <- "chl_source"

if (is.na(col_sal)) stop("Could not find salinity column.")
if (is.na(col_tmp)) stop("Could not find temperature column.")
if (is.na(col_o2))  stop("Could not find oxygen column.")
stopifnot(col_chl %in% names(df0))
stopifnot(col_chl_src %in% names(df0))

# col_date <- pick_first_existing(c("date","Date","cast_date","datetime","time","timestamp"), names(df0))

df0 <- 
  df0 %>%
  mutate(
    cruise_file = as.character(cruise_file),
    station     = str_squish(as.character(station)),
    latitude    = suppressWarnings(as.numeric(latitude)),
    longitude   = suppressWarnings(as.numeric(longitude)),
    sal         = suppressWarnings(as.numeric(.data[[col_sal]])),
    temp        = suppressWarnings(as.numeric(.data[[col_tmp]])),
    oxy         = suppressWarnings(as.numeric(.data[[col_o2]])),
    chl         = suppressWarnings(as.numeric(.data[[col_chl]])),
    chl_source  = as.character(.data[[col_chl_src]]),
    date_raw    = if (!is.na(col_date)) .data[[col_date]] else NA,
    date_raw    = mdy(date_raw)
  ) %T>% 
  print()
```


```{r}
df <-
  df0 %>%
  filter(
    # str_detect(cruise_file, cruise_pick), # select cruise
    str_detect(cruise_file, cruise_id), # select cruise
    !is.na(station), 
    station != "",
    !is.na(latitude), 
    !is.na(longitude)
  ) %T>% 
  print()

if (nrow(df) == 0) stop("No rows matched cruise_pick = ", cruise_pick)

# ---- date range text ---- #
date_vec <- df$date_raw
# date_vec <- NULL
# if (!is.na(col_date)) {
#   
#   date_vec <- suppressWarnings(as.Date(df$date_raw))
#   if (all(is.na(date_vec))) {
#     date_vec <- suppressWarnings(as.Date(parse_date_time(
#       df$date_raw,
#       orders = c("ymd","mdy","dmy","ymd HMS","mdy HMS","dmy HMS","Ymd HMS","mdY HMS","dmY HMS")
#     )))
#   }
# }
# date_min <- if (!is.null(date_vec) && any(!is.na(date_vec))) min(date_vec, na.rm = TRUE) else NA
# date_max <- if (!is.null(date_vec) && any(!is.na(date_vec))) max(date_vec, na.rm = TRUE) else NA
date_min <- min(date_vec, na.rm = TRUE)
date_max <- max(date_vec, na.rm = TRUE)
date_txt <-
  # if (!is.na(date_min) && !is.na(date_max)) {
  if (all(!is.na(c(date_min, date_max)))) {
    paste(format(date_min, "%Y-%m-%d"), "to", format(date_max, "%Y-%m-%d"))
  } else {
    "Date range: NA"
  }

# ---- chlorophyll source MODE for THIS cruise (panel title only) ---- #
chl_src_tbl <- df %>%
  filter(!is.na(chl_source), chl_source != "") %>%
  count(chl_source, name = "n") %>%
  arrange(desc(n))

chl_src_mode <- if (nrow(chl_src_tbl) == 0) {
  "unknown"
} else {
  chl_src_tbl$chl_source[1]
}

## ---- station means (ONLY aggregation; no smoothing of source data)
# df_st1 <- df %>%
#   # group_by(station) %>%
#   summarise(
#     latitude  = mean(latitude,  na.rm = TRUE),
#     longitude = mean(longitude, na.rm = TRUE),
#     sal  = mean(sal,  na.rm = TRUE),
#     temp = mean(temp, na.rm = TRUE),
#     oxy  = mean(oxy,  na.rm = TRUE),
#     chl  = mean(chl,  na.rm = TRUE),
#     # .groups = "drop"
#   ) %>%
#   filter(!is.na(latitude), !is.na(longitude))

df_st <- 
  df %>%
  summarise(
    .by = station,
    across(
      c(latitude, longitude, sal, temp, oxy, chl),
      \(.x) mean(.x, na.rm = TRUE)
    )
  ) %>%
  filter(!is.na(latitude), !is.na(longitude)) %T>% 
  print()

# pts_ll <- st_as_sf(df_st, coords = c("longitude","latitude"), crs = 4326)
```


## 2) BASEMAP: FLORIDA (SAME EXTENT)

```{r}
# bound
xlim <- c(-84.5, -79.5)
ylim <- c(24.2, 28.6)

# projection
utm_epsg <- 32617 # between 84W and 78W
# utm_epsg <- 2237 # for Florida zone 17N

# usa_states_v <- geodata::gadm(country = "USA", level = 1, path = tempdir())
# florida_sf <- 
#   st_as_sf(usa_states_v) %>%
#   filter(NAME_1 == "Florida") %>%
#   st_transform(4326) %>%
#   st_crop(xmin = -85, xmax = -79, ymin = 24, ymax = 29)

# florida map object from geodata then projected
florida_sf <- 
  geodata::gadm(country = "USA", level = 1, path = tempdir()) %>%
  st_as_sf() %>%
  filter(NAME_1 == "Florida") %>%
  st_transform(4326) %>% 
  st_crop(xmin = -85, xmax = -79, ymin = 24, ymax = 29) %>%
  st_transform(utm_epsg) 


```

# 3) FIXED TEMPLATE GRID (UTM) + IDW SETTINGS

```{r}
# corners_ll <- st_as_sf(
#   data.frame(
#     lon = c(xlim[1], xlim[2], xlim[2], xlim[1]),
#     lat = c(ylim[1], ylim[1], ylim[2], ylim[2])
#   ),
#   coords = c("lon","lat"),
#   crs = 4326
# )

# create sf object bounds bounding box and projected
# bb_utm <- st_bbox(st_transform(corners_ll, utm_epsg))
bb_utm <-
  data.frame(
    lon = c(xlim[1], xlim[2], xlim[2], xlim[1]),
    lat = c(ylim[1], ylim[1], ylim[2], ylim[2])
  ) %>%
  st_as_sf(
    coords = c("lon", "lat"),
    crs = 4326
  ) %>%
  # converts dd to meters
  st_transform(utm_epsg) %>%
  st_bbox()


# finer grid = smoother-looking surface without changing source data
grid_res_m <- 1000

# create raster grid with equal grid numbers in x and y
r_tmpl_utm <- terra::rast(
  xmin       = bb_utm["xmin"], 
  xmax       = bb_utm["xmax"],
  ymin       = bb_utm["ymin"], 
  ymax       = bb_utm["ymax"],
  resolution = grid_res_m,
  crs        = paste0("EPSG:", utm_epsg)
)

# convert projected raster to sf object with selected resolution
# this will be the sf object to map predicted values from interpolation (IDW)  
grid_sf <- 
  as.data.frame(r_tmpl_utm, xy = TRUE, cells = TRUE, na.rm = FALSE) %>%
  st_as_sf(coords = c("x","y"), crs = utm_epsg)

# pts_utm <- st_transform(pts_ll, utm_epsg)

# local IDW = tighter, less spread, more visually refined
# idp_pow   <- 2.0
# maxdist_m <- 120000
# nmin_nb   <- 3
# nmax_nb   <- 10
```


# 4) TIGHT HARD SUPPORT MASK — computed once

```{r}
# convert dataframe of stations to sf object and project
pts_utm <- 
  st_as_sf(df_st, coords = c("longitude","latitude"), crs = 4326, remove = FALSE)  %>%
  st_transform(utm_epsg) %>%
  mutate(
    longitude = st_coordinates(.)[,1],
    latitude = st_coordinates(.)[,2]
    ) 


# create a smaller, irregular bounding box based on convex hull or buffer (<3)
if (nrow(df_st) >= 3) {
  # convex hull, needs at least 3 points
  support_poly_utm <- 
    pts_utm %>%
    st_union() %>%
    st_convex_hull() %>%
    st_buffer(8000)   # 8 km buffer
} else {
  # buffer around each point
  support_poly_utm <- 
    pts_utm %>%
    st_buffer(30000) %>%
    st_union()
}

# transfomrm to geographic (decimal degrees)
# support_poly_ll <- st_transform(support_poly_utm, 4326) # <-- not used

# convert convex hull to linestring
support_boundary_utm <- st_boundary(support_poly_utm)

# show locations, convex hull in relation to Florida
ggplot() +
  geom_sf(data = grid_sf, color = "#500", alpha = 0.5) + # base grid for interpolation
  geom_sf(data = florida_sf) + # florida map
  geom_sf(
    data = support_poly_utm,
    color = "blue", fill = "lightblue", alpha = 0.5
  ) + # convex hull
  geom_sf(data = pts_utm) + # data points used
  coord_sf(expand = FALSE) +
  ggrepel::geom_text_repel(
    data = pts_utm,
    aes(x = longitude, y = latitude, label = station),
    seed = 1,
    size = 2.0,
    color = "black",
    box.padding = 0.18,
    point.padding = 0.10,
    min.segment.length = 0,
    segment.color = "grey35",
    segment.size = 0.30,
    segment.curvature = 0.0,
    arrow = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
    max.overlaps = Inf
  )
```

# 5) DISPLAY SMOOTHING SETTINGS

This does NOT change source data, only the plotted raster

```{r}
# Light 3x3 weighted kernel (gentle smoothing only)
smooth_kernel <-
  matrix(
    c(
      1, 2, 1,
      2, 4, 2,
      1, 2, 1
    ),
    nrow = 3, byrow = TRUE
  )
smooth_kernel <- smooth_kernel / sum(smooth_kernel)
```


# 6) SOFT EDGE SETTINGS (opacity of points)
This only softens the DISPLAY edge near the mask boundary

alpha refers to alpha of ggplot meaning opacity

```{r}
edge_feather_m <- 5000  # try 4000–7000 depending on taste

# support_boundary_utm <- st_boundary(support_poly_utm)

# calculate distance from convex hull (as linestring) then mask to within convex hull
# dist_to_edge_utm <- terra::distance(r_tmpl_utm, terra::vect(support_boundary_utm))
# dist_to_edge_utm <- terra::mask(dist_to_edge_utm, terra::vect(support_poly_utm))
dist_to_edge_utm <- 
  terra::distance(r_tmpl_utm, terra::vect(support_boundary_utm), names = "alpha") %>%
  terra::mask(terra::vect(support_poly_utm))

# divides distances by edge buffer then sets upper bounds to 1 by converting all values
# above 1 to 1 and leaves all values below as is
# alpha_utm <- dist_to_edge_utm / edge_feather_m
# alpha_utm <- terra::clamp(alpha_utm, lower = 0, upper = 1, values = TRUE)
alpha_utm <- 
  terra::clamp(dist_to_edge_utm / edge_feather_m, lower = 0, upper = 1, values = TRUE) 


ggplot() +
  tidyterra::geom_spatraster(data = alpha_utm) +
  scale_fill_viridis()
```


# 7) PANEL MAKER

```{r}
val_vec = df_st$sal
col_name = "sal"
var_title = "Salinity"
legend_units = "PSU"
.df_st            = df_st
.alpha_utm        = alpha_utm
.smooth_kernel    = smooth_kernel
.grid_sf          = grid_sf
.r_tmpl_utm       = r_tmpl_utm
.utm_epsg         = utm_epsg
.support_poly_utm = support_poly_utm
.florida_sf       = florida_sf
idw_opts          = idw_options

col_name <- "sal"

tet <- 
 pts_utm %>%
    # transmute(
    mutate(
      .keep = "none",
      station,
      y = latitude,
      x = longitude,
      # val = suppressWarnings(as.numeric(val_vec))
      val = .data[[col_name]],
      geometry
    ) %>%
    filter(!is.na(val)) %T>% 
  print()
 
pts_ll2

 # terra::values(.r_tmpl_utm)[.grid_sf$cell] <- as.numeric(idw_sf_utm$var1.pred)
ggplot() +
 tidyterra::geom_spatraster(
   data = r_pred_utm_sm,
   # aes(lyr = 1)
 )
ggplot() +
 tidyterra::geom_spatraster(
   data = r_tmpl_utm
 ) +
  scale_fill_viridis(na.value = NA)
ggplot() +
 tidyterra::geom_spatraster(
   data = r_pred_utm
 ) +
  scale_fill_viridis(na.value = NA)

ggplot() +
 # tidyterra::geom_spatvector(data = terra::vect(.florida_sf))
 tidyterra::geom_spatraster(
   data = r_pred_water 
   # data = terra::mask(r_pred_utm_sm,  terra::vect(.florida_sf), inverse = TRUE)
# data = terra::mask(r_pred_ll, terra::vect(.florida_sf), inverse = TRUE)
 ) +
  scale_fill_viridis(na.value = NA)
ggplot() +
geom_sf(data = .florida_sf)
terra::vect(.florida_sf)




r_pred_utm_sm <- 
    # st_intersection(idw_sf_utm, .support_poly_utm) %>%
    
    terra::focal(
      
      w         = .smooth_kernel,
      fun       = mean,
      na.policy = "omit",
      na.rm     = TRUE,
      fillvalue = NA
    )
    
ggplot() +
 tidyterra::geom_spatraster(
   data =  r_pred_water
 ) 
  # scale_fill_viridis(na.value = NA)
ggplot() +
 tidyterra::geom_spatraster(
   data =  c(r_pred_water, alpha_utm),
   aes(fill = focal_mean)
 ) +
  scale_fill_viridis(na.value = NA) 
ggplot() +
 tidyterra::geom_spatraster(
   data =  r_pred_water1,
   aes(fill = dist)
 ) +
  scale_fill_viridis(na.value = NA) 
ggplot() +
 tidyterra::geom_spatvector(
   data =  contour_sf
 ) +
  scale_fill_viridis(na.value = NA) 

ggplot() +
  geom_sf(data = sf::st_as_sf(contour_sf)) +
 tidyterra::geom_spatvector(
   data =  contour_sf,
   color = "green",
   linetype = "dashed"
 ) 


ggplot() +
  geom_raster(data = plot_df, aes(x = x, y = y, fill = pred, alpha = alpha)) + 
  theme(legend.title = element_blank())

ggplot() +
  geom_raster(
    data = as.data.frame(r_pred_water, xy = TRUE), aes(x = x, y = y, fill = focal_mean, alpha = alpha)
  ) + 
  theme(legend.title = element_blank())


pts_utm$geometry
```


```{r}
# local IDW = tighter, less spread, more visually refined
idw_options <-
  list(
    "idp_pow"   = 2.0,
    "maxdist_m" = 120000,
    "nmin_nb"   = 3,
    "nmax_nb"   = 10
  )

make_panel <- function(
  # val_vec, 
  col_name, 
  var_title, 
  legend_units, 
  # .df_st            = df_st,  
  .df_st            = pts_utm,  
  .alpha_utm        = alpha_utm,
  .smooth_kernel    = smooth_kernel,
  .grid_sf          = grid_sf,
  .r_tmpl_utm       = r_tmpl_utm,
  .utm_epsg         = utm_epsg,
  .support_poly_utm = support_poly_utm,
  .florida_sf       = florida_sf,
  .bb_utm = bb_utm,
  idw_opts          = idw_options
  ) {

  # dd <- .df_st %>%
  #   # transmute(
  #   mutate(
  #     .keep = "none",
  #     station,
  #     latitude,
  #     longitude,
  #     # val = suppressWarnings(as.numeric(val_vec))
  #     val = .data[[col_name]]
  #   ) %>%
  #   filter(!is.na(latitude), !is.na(longitude), !is.na(val))

  # keep columns needed for IDW interpolation and plotting, converts selected
  # parameter name to val, removes NA 
  dd <- 
    pts_utm %>%
    mutate(
      .keep = "none",
      station,
      latitude,
      longitude,
      val = .data[[col_name]],
      geometry
    ) %>%
    filter(!is.na(val))
  
  # skips if less than 5 points
  if (nrow(dd) < 5) {
    return(
      ggplot() +
        theme_void() +
        labs(title = paste0(var_title, " (too few stations)")) +
        theme(plot.title = element_text(face = "bold", size = 12))
    )
  }

  # adds geometry, probably not needed
  # pts_ll2  <- st_as_sf(dd, coords = c("longitude","latitude"), crs = 4326)
  # pts_utm2 <- st_transform(pts_ll2, .utm_epsg)

  # IDW interpolation as sf object
  # uses options set before
  # idw_res <- 
  idw_sf_utm <- 
    gstat::idw(
      formula   = val ~ 1,
      # locations = pts_utm2,
      # locations = pts_ll2,
      locations = dd,
      newdata   = .grid_sf,
      idp       = idw_opts$idp_pow,
      maxdist   = idw_opts$maxdist_m,
      nmin      = idw_opts$nmin_nb,
      nmax      = idw_opts$nmax_nb
    )
  

  # idw_sf_utm <- st_as_sf(idw_res)
  # st_crs(idw_sf_utm) <- .utm_epsg

  # replace empty grid with IDW values
  # r_pred_utm <- .r_tmpl_utm
  # terra::values(r_pred_utm)[.grid_sf$cell] <- as.numeric(idw_sf_utm$var1.pred)
  # terra::values(.r_tmpl_utm)[.grid_sf$cell] <- as.numeric(idw_sf_utm$var1.pred)
  terra::values(.r_tmpl_utm) <- as.numeric(idw_sf_utm$var1.pred)

  # hard mask first
  # from convex hull of data points
  # r_pred_utm <- terra::mask(r_pred_utm, terra::vect(.support_poly_utm))
  r_pred_utm <- terra::mask(.r_tmpl_utm, terra::vect(.support_poly_utm))

  # LIGHT DISPLAY SMOOTHING ONLY
  # smooths over values with weighting of 3x3 kernel
  # r_pred_utm_sm <- 
  r_pred_water <-
     terra::mask(.r_tmpl_utm, terra::vect(.support_poly_utm)) %>%
    terra::focal(
      # r_pred_utm,
      w         = .smooth_kernel,
      fun       = mean,
      na.policy = "omit",
      na.rm     = TRUE,
      fillvalue = NA
    ) %>%
    
    # combine IDW and alpha
    c(., alpha_utm) %>%
    
    # mask again so smoothing does not bleed outside supported footprint
    # of convex hull
    terra::mask(terra::vect(.support_poly_utm)) %>% # <-- may not be necessary
    
    # mask land mass
    terra::mask(terra::vect(.florida_sf), inverse = TRUE) 
    
  plot_df <- 
    r_pred_water %>%
    # convert predicted values and alpha into data frames and combine
    as.data.frame(xy = TRUE, na.rm = FALSE) %>%
    rename("pred" = focal_mean)  %>%
    mutate(alpha = ifelse(is.na(alpha), 0, alpha)) %>%
    filter(!is.na(pred), alpha > 0)
  
  
  
  # mask again so smoothing does not bleed outside supported footprint
  # r_pred_utm_sm <- terra::mask(r_pred_utm_sm, terra::vect(.support_poly_utm))

  # alpha for soft display edge
  # within set distance from edge, should already be clipped
  # alpha_pred_utm <- terra::mask(.alpha_utm, terra::vect(.support_poly_utm))
  # alpha_pred_utm <- .alpha_utm

  # project for plotting
  # r_pred_ll <- terra::project(r_pred_utm_sm, "EPSG:4326", method = "bilinear")
  # alpha_ll  <- terra::project(alpha_pred_utm, "EPSG:4326", method = "bilinear")

  # remove/mask land mass
  # r_pred_water <- terra::mask(r_pred_ll, terra::vect(.florida_sf), inverse = TRUE)
  # alpha_water  <- terra::mask(alpha_ll,  terra::vect(.florida_sf), inverse = TRUE)
  # r_pred_water <- terra::mask(r_pred_utm_sm, terra::vect(.florida_sf), inverse = TRUE)
  # alpha_water  <- terra::mask(.alpha_utm, terra::vect(.florida_sf), inverse = TRUE)

  # # convert predicted values and alpha into data frames and combine
  # pred_df <- 
  #   as.data.frame(r_pred_water, xy = TRUE, na.rm = FALSE) %>%
  #   select("x" = 1, "y" = 2, "pred" = 3)
  # # names(pred_df) <- c("x","y","pred")
  # 
  # alpha_df <- as.data.frame(alpha_water, xy = TRUE, na.rm = FALSE) %>%
  #   select("x" = 1, "y" = 2, "alpha" = 3)
  # # names(alpha_df) <- c("x","y","alpha")

  # plot_df <- 
  #   pred_df %>%
  #   # left_join(alpha_df, by = c("x","y")) %>%
  #   mutate(alpha = ifelse(is.na(alpha), 0, alpha)) %>%
  #   filter(!is.na(pred), alpha > 0)

  
  # subtle contours for a polished look
  rng <- range(plot_df$pred, na.rm = TRUE)
  contour_sf <- NULL

  if (all(is.finite(rng)) && diff(rng) > 0) {
    brks <- pretty(rng, n = 6)
    # brks <- brks[brks > rng[1] & brks < rng[2]]
    brks <- brks[between(brks, rng[1], rng[2])]

    if (length(brks) > 0) {
      # contour_obj <- try(
      contour_sf <- try(
        terra::as.contour(r_pred_water, levels = brks),
        silent = TRUE
      )
      # if error, convert to sf
      # if (!inherits(contour_obj, "try-error")) {
      if (!inherits(contour_sf, "try-error")) {
        # contour_sf <- try(sf::st_as_sf(contour_obj), silent = TRUE)
        contour_sf <- try(sf::st_as_sf(contour_sf), silent = TRUE)
      }
      
      # if error on conversion, set to NULL
      if (inherits(contour_sf, "try-error")) contour_sf <- NULL
    }
  }

  p <- 
    ggplot() +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "#cfe8f3") +
    geom_raster(data = plot_df, aes(x = x, y = y, fill = pred, alpha = alpha)) +
    scale_fill_viridis_c(name = legend_units, option = "C", direction = 1, na.value = NA) +
    scale_alpha_identity() +
    geom_sf(data = .florida_sf, fill = "#e8e6df", color = "grey35", linewidth = 0.55)

  # add contours if possible
  if (!is.null(contour_sf) && nrow(contour_sf) > 0) {
    p <- p + 
      geom_sf(
        data      = contour_sf,
        color     = "white",
        linewidth = 0.22,
        alpha     = 0.45
    )
  }

  p <- p +
    # geom_sf(data = pts_ll2, color = "black", size = 0.55, alpha = 0.95) +
    geom_sf(data = dd, color = "black", size = 0.55, alpha = 0.95) +
    ggrepel::geom_text_repel(
      data = dd,
      aes(x = longitude, y = latitude, label = station),
      seed               = 1,
      size               = 2.0,
      color              = "black",
      box.padding        = 0.18,
      point.padding      = 0.10,
      min.segment.length = 0,
      segment.color      = "grey35",
      segment.size       = 0.30,
      segment.curvature  = 0.0,
      arrow              = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
      max.overlaps       = Inf
    ) +
    annotation_north_arrow(
      location    = "tr",
      which_north = "true",
      pad_x       = unit(0.10, "cm"),
      pad_y       = unit(0.10, "cm"),
      height      = unit(0.85, "cm"),
      width       = unit(0.85, "cm"),
      style       = north_arrow_orienteering(
        line_col = "grey20",
        fill     = c("grey20", "white"),
        text_col = "grey20"
      )
    ) +
    annotation_scale(
      location   = "bl",
      width_hint = 0.22,
      pad_x      = unit(0.12, "cm"),
      pad_y      = unit(0.12, "cm"),
      text_cex   = 0.55,
      line_width = 0.6
    ) +
    # coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    coord_sf(xlim = .bb_utm[c(1,3)], ylim = .bb_utm[c(2,4)], expand = FALSE) +
    labs(title = var_title, x = NULL, y = NULL) +
    theme_bw() +
    theme(
      panel.background = element_rect(fill = "#cfe8f3"),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      legend.position  = "right",
      # legend.title     = element_text(face = "bold"),
      plot.title       = element_text(face = "bold", size = 12)
    )

  p
}
```


# 8) CHL PANEL TITLE (no parentheses) + UNITS ON COLORBAR ONLY
```{r}
chl_panel_title <- if (identical(chl_src_mode, "chlorophyll_concentration")) {
  "Chlorophyll concentration"
} else if (identical(chl_src_mode, "chlorophyll_fluorescence")) {
  "Chlorophyll fluorescence"
} else {
  "Chlorophyll"
}

chl_units <- expression(paste("mg ", m^{-3}))
```


# 9) BUILD 4 PANELS
```{r}
# p_sal <- make_panel(df_st$sal,  "Salinity", "PSU")
# p_tmp <- make_panel(df_st$temp, "Temperature", expression(paste(degree, "C")))
# p_o2  <- make_panel(df_st$oxy,  "Oxygen", "mL/L")
# p_chl <- make_panel(df_st$chl,  chl_panel_title, chl_units)
p_sal <- make_panel("sal",  "Salinity", "PSU")
p_tmp <- make_panel("temp", "Temperature", expression(paste(degree, "C")))
p_o2  <- make_panel("oxy",  "Oxygen", "mL/L")
p_chl <- make_panel("chl",  chl_panel_title, chl_units)
```


# 10) COMBINE + TITLE/SUBTITLE (NO caption)

```{r}
main_title <- paste0(cruise_id, " — Surface (0–5 m mean)")
sub_title  <- paste0(date_txt, " | IDW + (UTM grid)")

final_plot <- (p_sal + p_tmp) / (p_o2 + p_chl) +
  plot_annotation(
    title = main_title,
    subtitle = sub_title
  )

final_plot <- final_plot + plot_annotation(
  theme = theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 11)
  )
)

print(final_plot)
```


# 11) SAVE (SHORT FILENAMES to avoid OneDrive truncation)
```{r}
png_out  <- here(map_dir, paste0(cruise_id, "_0_5m_4vars_softedge_nav.png"))
tiff_out <- here(map_dir, paste0(cruise_id, "_0_5m_4vars_softedge_nav.tiff"))

ggsave(png_out,  plot = final_plot, width = 12.5, height = 9.5, dpi = 600)
ggsave(tiff_out, plot = final_plot, width = 12.5, height = 9.5, dpi = 600, compression = "lzw")

cat("\nOutput folder:\n", map_dir, "\n\n")
cat("\nSaved PNG:\n", basename(png_out),  "\nexists =", file_exists(png_out),  "\n")
cat("Saved TIFF:\n", basename(tiff_out), "\nexists =", file_exists(tiff_out), "\n")
print(list.files(map_dir))
print(basename(dir_ls(map_dir)))
```



# =========================================================================== #

```{r}
# =========================================================================== #
# ALL cruises: 4-panel Surface Maps (0–5 m mean) — LOOP VERSION
# Variables: Salinity, Temperature, Oxygen, Chlorophyll (CHL_0_5m + chl_source)
# Method: IDW (UTM grid) + LIGHT DISPLAY SMOOTHING

# Style: subtle contour lines + north arrow + scale bar
# Labels: station names on ALL panels with SHORT STRAIGHT ARROW leader lines
# FOCUS RANGES (pattern emphasis):
#   Salinity 28–35 PSU | Oxygen 1–5 mL/L | Temp 18–30 °C | Chl 0–5 mg m^-3
# Output: PNG

# =========================================================================== #

library(dplyr)
library(readr)
library(sf)
library(ggplot2)
library(ggspatial)
library(stringr)
library(terra)
library(geodata)
library(gstat)
library(viridis)
library(ggrepel)
library(grid)     
library(patchwork) # combine panels
library(lubridate)
library(scales)
```


# 0) PATHS

```{r}
base_dir <- "C:/Users/mostafaawada/OneDrive - University of South Florida/Desktop/CTD_data_codes/data/cleaned_ctd_data_v02/cleaned_ctd_data_v02/merged_by_cruise"
ts_dir   <- file.path(base_dir, "Time_series_new")

in_file  <- file.path(ts_dir, "CTD_surface_0_5m_ALL_casts_withRegionShelf_ALLPARAMS.csv")
stopifnot(file.exists(in_file))

# NEW folder for focus-range maps
map_dir <- file.path(ts_dir, "Figures_freshening", "Surf4_focusRanges_png")
dir.create(map_dir, showWarnings = FALSE, recursive = TRUE)
```


# 1) READ + COLUMN DETECTION

```{r}
df0 <- read_csv(in_file, show_col_types = FALSE)

need_base <- c("cruise_file","station","latitude","longitude")
miss_base <- setdiff(need_base, names(df0))
if (length(miss_base) > 0) stop("Missing base columns: ", paste(miss_base, collapse = ", "))

pick_first_existing <- function(candidates, col_names) {
  hit <- candidates[candidates %in% col_names]
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

col_sal <- pick_first_existing(c("S_0_5m","sal_0_5m","mean_sal_psu","salinity"), names(df0))
col_tmp <- pick_first_existing(c("T_0_5m","temp_0_5m","mean_temp_C","temperature","temp_C"), names(df0))
col_o2  <- pick_first_existing(c("O2_0_5m","oxy_0_5m","mean_oxy","oxygen","DO"), names(df0))

# confirmed by you:
col_chl     <- "CHL_0_5m"
col_chl_src <- "chl_source"

if (is.na(col_sal)) stop("Could not find salinity column.")
if (is.na(col_tmp)) stop("Could not find temperature column.")
if (is.na(col_o2))  stop("Could not find oxygen column.")
stopifnot(col_chl %in% names(df0))
stopifnot(col_chl_src %in% names(df0))

col_date <- pick_first_existing(c("date","Date","cast_date","datetime","time","timestamp"), names(df0))
```


# 2) FOCUS RANGES (pattern emphasis)
```{r}
sal_limits  <- c(28, 35)
oxy_limits  <- c(1,  5)
temp_limits <- c(18, 30)
chl_limits  <- c(0,  5)

cat("\nFOCUS color limits:\n")
cat("Salinity:    ", sal_limits[1],  " to ", sal_limits[2],  "\n", sep = "")
cat("Oxygen:      ", oxy_limits[1],  " to ", oxy_limits[2],  "\n", sep = "")
cat("Temperature: ", temp_limits[1], " to ", temp_limits[2], "\n", sep = "")
cat("Chlorophyll: ", chl_limits[1],  " to ", chl_limits[2],  "\n\n", sep = "")
```


# 3) CLEAN ALL DATA ONCE
```{r}
df_all <- df0 %>%
  mutate(
    cruise_file = as.character(cruise_file),
    station     = str_squish(as.character(station)),
    latitude    = suppressWarnings(as.numeric(latitude)),
    longitude   = suppressWarnings(as.numeric(longitude)),
    sal         = suppressWarnings(as.numeric(.data[[col_sal]])),
    temp        = suppressWarnings(as.numeric(.data[[col_tmp]])),
    oxy         = suppressWarnings(as.numeric(.data[[col_o2]])),
    chl         = suppressWarnings(as.numeric(.data[[col_chl]])),
    chl_source  = as.character(.data[[col_chl_src]]),
    date_raw    = if (!is.na(col_date)) .data[[col_date]] else NA
  ) %>%
  filter(
    !is.na(cruise_file), cruise_file != "",
    !is.na(station), station != "",
    !is.na(latitude), !is.na(longitude)
  )

if (nrow(df_all) == 0) stop("No valid rows found after filtering.")
```


# 4) BASEMAP: FLORIDA (SAME EXTENT) — load once
```{r}
xlim <- c(-84.5, -79.5)
ylim <- c(24.2, 28.6)

usa_states_v <- geodata::gadm(country = "USA", level = 1, path = tempdir())
florida_sf <- st_as_sf(usa_states_v) %>%
  filter(NAME_1 == "Florida") %>%
  st_transform(4326) %>%
  st_crop(xmin = -85, xmax = -79, ymin = 24, ymax = 29)
```


# 5) FIXED TEMPLATE GRID (UTM) + IDW SETTINGS — build once
```{r}

utm_epsg <- 32617

corners_ll <- st_as_sf(
  data.frame(
    lon = c(xlim[1], xlim[2], xlim[2], xlim[1]),
    lat = c(ylim[1], ylim[1], ylim[2], ylim[2])
  ),
  coords = c("lon","lat"),
  crs = 4326
)
bb_utm <- st_bbox(st_transform(corners_ll, utm_epsg))

grid_res_m <- 1000

r_tmpl_utm <- terra::rast(
  xmin = bb_utm["xmin"], xmax = bb_utm["xmax"],
  ymin = bb_utm["ymin"], ymax = bb_utm["ymax"],
  resolution = grid_res_m,
  crs = paste0("EPSG:", utm_epsg)
)

grid_df <- as.data.frame(r_tmpl_utm, xy = TRUE, cells = TRUE, na.rm = FALSE)
grid_sf <- st_as_sf(grid_df, coords = c("x","y"), crs = utm_epsg)

# local IDW = tighter, less spread
idp_pow   <- 2.0
maxdist_m <- 120000
nmin_nb   <- 3
nmax_nb   <- 10
```


# 6) DISPLAY SMOOTHING KERNEL — fixed
```{r}
smooth_kernel <- matrix(
  c(1, 2, 1,
    2, 4, 2,
    1, 2, 1),
  nrow = 3, byrow = TRUE
)
smooth_kernel <- smooth_kernel / sum(smooth_kernel)
```


# 7) SOFT EDGE SETTINGS — fixed
```{r}
edge_feather_m <- 5000
```


# 8) PANEL MAKER
```{r}

make_panel <- function(df_st, support_poly_utm, alpha_utm,
                       val_vec, var_title, legend_units, legend_limits) {

  dd <- df_st %>%
    transmute(
      station,
      latitude,
      longitude,
      val = suppressWarnings(as.numeric(val_vec))
    ) %>%
    filter(!is.na(latitude), !is.na(longitude), !is.na(val))

  if (nrow(dd) < 5) {
    return(
      ggplot() +
        theme_void() +
        labs(title = paste0(var_title, " (too few stations)")) +
        theme(plot.title = element_text(face = "bold", size = 12))
    )
  }

  pts_ll2  <- st_as_sf(dd, coords = c("longitude","latitude"), crs = 4326)
  pts_utm2 <- st_transform(pts_ll2, utm_epsg)

  idw_res <- gstat::idw(
    formula   = val ~ 1,
    locations = pts_utm2,
    newdata   = grid_sf,
    idp       = idp_pow,
    maxdist   = maxdist_m,
    nmin      = nmin_nb,
    nmax      = nmax_nb
  )

  idw_sf_utm <- st_as_sf(idw_res)
  st_crs(idw_sf_utm) <- utm_epsg

  r_pred_utm <- r_tmpl_utm
  terra::values(r_pred_utm)[grid_df$cell] <- as.numeric(idw_sf_utm$var1.pred)

  # hard mask first
  r_pred_utm <- terra::mask(r_pred_utm, terra::vect(support_poly_utm))

  # LIGHT DISPLAY SMOOTHING ONLY
  r_pred_utm_sm <- terra::focal(
    r_pred_utm,
    w = smooth_kernel,
    fun = mean,
    na.policy = "omit",
    na.rm = TRUE,
    fillvalue = NA
  )

  # mask again so smoothing does not bleed outside supported footprint
  r_pred_utm_sm <- terra::mask(r_pred_utm_sm, terra::vect(support_poly_utm))

  # alpha for soft display edge
  alpha_pred_utm <- terra::mask(alpha_utm, terra::vect(support_poly_utm))

  # project for plotting
  r_pred_ll <- terra::project(r_pred_utm_sm, "EPSG:4326", method = "bilinear")
  alpha_ll  <- terra::project(alpha_pred_utm, "EPSG:4326", method = "bilinear")

  # remove land
  r_pred_water <- terra::mask(r_pred_ll, terra::vect(florida_sf), inverse = TRUE)
  alpha_water  <- terra::mask(alpha_ll,  terra::vect(florida_sf), inverse = TRUE)

  pred_df <- as.data.frame(r_pred_water, xy = TRUE, na.rm = FALSE)
  names(pred_df) <- c("x","y","pred")

  alpha_df <- as.data.frame(alpha_water, xy = TRUE, na.rm = FALSE)
  names(alpha_df) <- c("x","y","alpha")

  plot_df <- pred_df %>%
    left_join(alpha_df, by = c("x","y")) %>%
    mutate(alpha = ifelse(is.na(alpha), 0, alpha)) %>%
    filter(!is.na(pred), alpha > 0) %>%
    select(x, y, pred, alpha)

  # subtle contours for a polished look
  rng <- range(plot_df$pred, na.rm = TRUE)
  contour_sf <- NULL

  if (is.finite(rng[1]) && is.finite(rng[2]) && diff(rng) > 0) {
    brks <- pretty(rng, n = 6)
    brks <- brks[brks > rng[1] & brks < rng[2]]

    if (length(brks) > 0) {
      contour_obj <- try(
        terra::as.contour(r_pred_water, levels = brks),
        silent = TRUE
      )

      if (!inherits(contour_obj, "try-error")) {
        contour_sf <- try(sf::st_as_sf(contour_obj), silent = TRUE)
        if (inherits(contour_sf, "try-error")) contour_sf <- NULL
      }
    }
  }

  p <- ggplot() +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "#cfe8f3") +
    geom_raster(data = plot_df, aes(x = x, y = y, fill = pred, alpha = alpha)) +
    scale_fill_viridis_c(
      name = legend_units,
      option = "C",
      direction = 1,
      limits = legend_limits,
      oob = scales::squish,
      na.value = NA
    ) +
    scale_alpha_identity() +
    geom_sf(data = florida_sf, fill = "#e8e6df", color = "grey35", linewidth = 0.55)

  if (!is.null(contour_sf) && nrow(contour_sf) > 0) {
    p <- p + geom_sf(
      data = contour_sf,
      color = "white",
      linewidth = 0.22,
      alpha = 0.45
    )
  }

  p <- p +
    geom_sf(data = pts_ll2, color = "black", size = 0.55, alpha = 0.95) +
    ggrepel::geom_text_repel(
      data = dd,
      aes(x = longitude, y = latitude, label = station),
      seed = 1,
      size = 2.0,
      color = "black",
      box.padding = 0.18,
      point.padding = 0.10,
      min.segment.length = 0,
      segment.color = "grey35",
      segment.size = 0.30,
      segment.curvature = 0.0,
      arrow = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
      max.overlaps = Inf
    ) +
    annotation_north_arrow(
      location = "tr",
      which_north = "true",
      pad_x = unit(0.10, "cm"),
      pad_y = unit(0.10, "cm"),
      height = unit(0.85, "cm"),
      width = unit(0.85, "cm"),
      style = north_arrow_orienteering(
        line_col = "grey20",
        fill = c("grey20", "white"),
        text_col = "grey20"
      )
    ) +
    annotation_scale(
      location = "bl",
      width_hint = 0.22,
      pad_x = unit(0.12, "cm"),
      pad_y = unit(0.12, "cm"),
      text_cex = 0.55,
      line_width = 0.6
    ) +
    coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
    labs(title = var_title, x = NULL, y = NULL) +
    theme_bw() +
    theme(
      panel.background = element_rect(fill = "#cfe8f3"),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      legend.position  = "right",
      legend.title     = element_text(face = "bold"),
      plot.title       = element_text(face = "bold", size = 12)
    )

  p
}
```


# 9) CRUISE LIST

```{r}
cruise_list <- df_all %>%
  distinct(cruise_file) %>%
  arrange(cruise_file) %>%
  pull(cruise_file)

cat("Total cruises found: ", length(cruise_list), "\n")
```


# 10) LOOP OVER ALL CRUISES
```{r}
for (cruise_pick in cruise_list) {

  cat("\n----------------------------------\n")
  cat("Processing: ", cruise_pick, "\n", sep = "")

  cruise_id <- str_replace(cruise_pick, "\\.csv$", "")
  cruise_id <- str_replace_all(cruise_id, "[^A-Za-z0-9_\\-]+", "_")

  df <- df_all %>%
    filter(cruise_file == cruise_pick)

  if (nrow(df) == 0) {
    cat("Skipped: no rows.\n")
    next
  }

  ## ---- date range text
  date_vec <- NULL
  if (!is.na(col_date)) {
    date_vec <- suppressWarnings(as.Date(df$date_raw))
    if (all(is.na(date_vec))) {
      date_vec <- suppressWarnings(as.Date(parse_date_time(
        df$date_raw,
        orders = c("ymd","mdy","dmy","ymd HMS","mdy HMS","dmy HMS","Ymd HMS","mdY HMS","dmY HMS")
      )))
    }
  }

  date_min <- if (!is.null(date_vec) && any(!is.na(date_vec))) min(date_vec, na.rm = TRUE) else NA
  date_max <- if (!is.null(date_vec) && any(!is.na(date_vec))) max(date_vec, na.rm = TRUE) else NA
  date_txt <- if (!is.na(date_min) && !is.na(date_max)) {
    paste0(format(date_min, "%Y-%m-%d"), " to ", format(date_max, "%Y-%m-%d"))
  } else {
    "Date range: NA"
  }

  ## ---- chlorophyll source MODE
  chl_src_tbl <- df %>%
    filter(!is.na(chl_source), chl_source != "") %>%
    count(chl_source, name = "n") %>%
    arrange(desc(n))

  chl_src_mode <- if (nrow(chl_src_tbl) == 0) "unknown" else chl_src_tbl$chl_source[1]

  ## ---- station means
  df_st <- df %>%
    group_by(station) %>%
    summarise(
      latitude  = mean(latitude,  na.rm = TRUE),
      longitude = mean(longitude, na.rm = TRUE),
      sal  = mean(sal,  na.rm = TRUE),
      temp = mean(temp, na.rm = TRUE),
      oxy  = mean(oxy,  na.rm = TRUE),
      chl  = mean(chl,  na.rm = TRUE),
      .groups = "drop"
    ) %>%
    filter(!is.na(latitude), !is.na(longitude))

  if (nrow(df_st) < 2) {
    cat("Skipped: too few stations.\n")
    next
  }

  pts_ll  <- st_as_sf(df_st, coords = c("longitude","latitude"), crs = 4326)
  pts_utm <- st_transform(pts_ll, utm_epsg)

  ## ---- support mask
  if (nrow(df_st) >= 3) {
    support_poly_utm <- pts_utm %>%
      st_union() %>%
      st_convex_hull() %>%
      st_buffer(8000)
  } else {
    support_poly_utm <- pts_utm %>%
      st_buffer(30000) %>%
      st_union()
  }

  ## ---- soft edge alpha
  support_boundary_utm <- st_boundary(support_poly_utm)

  dist_to_edge_utm <- terra::distance(r_tmpl_utm, terra::vect(support_boundary_utm))
  dist_to_edge_utm <- terra::mask(dist_to_edge_utm, terra::vect(support_poly_utm))

  alpha_utm <- dist_to_edge_utm / edge_feather_m
  alpha_utm <- terra::clamp(alpha_utm, lower = 0, upper = 1, values = TRUE)

  ## ---- chl title
  chl_panel_title <- if (identical(chl_src_mode, "chlorophyll_concentration")) {
    "Chlorophyll concentration"
  } else if (identical(chl_src_mode, "chlorophyll_fluorescence")) {
    "Chlorophyll fluorescence"
  } else {
    "Chlorophyll"
  }

  chl_units <- expression(paste("mg ", m^{-3}))

  ## ---- build panels with focus limits
  p_sal <- make_panel(df_st, support_poly_utm, alpha_utm,
                      df_st$sal,  "Salinity",    "PSU",                    sal_limits)

  p_tmp <- make_panel(df_st, support_poly_utm, alpha_utm,
                      df_st$temp, "Temperature", expression(paste(degree, "C")), temp_limits)

  p_o2  <- make_panel(df_st, support_poly_utm, alpha_utm,
                      df_st$oxy,  "Oxygen",      "mL/L",                  oxy_limits)

  p_chl <- make_panel(df_st, support_poly_utm, alpha_utm,
                      df_st$chl,  chl_panel_title, chl_units,             chl_limits)

  ## ---- combine
  main_title <- paste0(cruise_id, " — Surface (0–5 m mean)")
  sub_title  <- paste0(date_txt, " | IDW + (UTM grid)")

  final_plot <- (p_sal + p_tmp) / (p_o2 + p_chl) +
    plot_annotation(title = main_title, subtitle = sub_title) +
    plot_annotation(
      theme = theme(
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 11)
      )
    )

  ## ---- save PNG only (short name)
  png_out <- file.path(map_dir, paste0(cruise_id, "_s4_focus.png"))
  ggsave(png_out, plot = final_plot, width = 12.5, height = 9.5, dpi = 300)

  cat("Saved: ", png_out, "\n", sep = "")
}

cat("\nFinished all cruises.\n")
cat("Output folder:\n", map_dir, "\n\n")
print(list.files(map_dir))
```



