---
title: "Surface_maps"
author: "Mostafa Soliman"
date: "2026-03-06"
output: html_document
---


ONE cruise: 4-panel Surface Maps (0–5 m mean)
Variables: Salinity, Temperature, Oxygen, Chlorophyll (CHL_0_5m + chl_source)
Method: IDW (UTM grid) + LIGHT DISPLAY SMOOTHING

Edge: TIGHT HARD SUPPORT MASK + SOFT FEATHERED DISPLAY EDGE
Style: subtle contour lines + north arrow + scale bar

Title:  "WS0603 — Surface (0–5 m mean)"
Subtitle: "YYYY-MM-DD to YYYY-MM-DD
CHL panel title: "Chlorophyll fluorescence" OR "Chlorophyll concentration"
Units ONLY on colorbar title


ALL cruises: 4-panel Surface Maps (0–5 m mean) — LOOP VERSION
Variables: Salinity, Temperature, Oxygen, Chlorophyll (CHL_0_5m + chl_source)
Method: IDW (UTM grid) + LIGHT DISPLAY SMOOTHING

Style: subtle contour lines + north arrow + scale bar
Labels: station names on ALL panels with SHORT STRAIGHT ARROW leader lines
FOCUS RANGES (pattern emphasis):
  Salinity 28–35 PSU | Oxygen 1–5 mL/L | Temp 18–30 °C | Chl 0–5 mg m^-3
Output: PNG

```{r setup}

if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
  
  # additional packages, if from GitHub, include username
  sf, ggspatial, terra, geodata, gstat, viridis, ggrepel, grid, 
  patchwork # combine panels
)

conflicts_prefer(
  dplyr::filter,
  dplyr::select
  )
```

# 0) PATHS

```{r paths}
base_dir <- here("data", "ctd_maps")
dir_create(base_dir)

ts_dir <- here(base_dir, "Time_series_new")
dir_create(ts_dir)

# file with all ctd casts
in_file <- 
  ts_dir  %>%
  dir_ls(regexp = "CTD_surface_0_5m_ALL")

stopifnot(length(in_file) != 0)

map_dir <- here(ts_dir, "Figures_freshening", "SurfaceMaps_0_5m_4vars_NOHALO")
dir_create(map_dir)

```

# 1) BASEMAP: FLORIDA (SAME EXTENT)

```{r base-map}
# projection
utm_epsg <- 32617 # between 84W and 78W
# utm_epsg <- 2237 # for Florida zone 17N

# florida map object from geodata then projected
florida_sf <- 
  geodata::gadm(country = "USA", level = 1, path = tempdir()) %>%
  st_as_sf() %>%
  filter(NAME_1 == "Florida") %>%
  st_transform(4326) %>% 
  st_crop(xmin = -85, xmax = -79, ymin = 24, ymax = 29) %>%
  st_transform(utm_epsg) 

# set overall bounding box for plotting
bb_utm <-
  c(
    xmin = -84.5,
    xmax = -79.5,
    ymin = 24.2,
    ymax = 28.6
  ) %>%
  st_bbox(crs = 4326) %>%
  st_as_sfc() %>%
  # converts dd to meters
  st_transform(utm_epsg) %>%
  st_bbox()
```

# 2) DISPLAY SMOOTHING SETTINGS

This does NOT change source data, only the plotted raster

```{r smoother-options}
# local IDW = tighter, less spread, more visually refined
idw_options <-
  list(
    "idp_pow"   = 2.0,
    "maxdist_m" = 120000,
    "nmin_nb"   = 3,
    "nmax_nb"   = 10
  )

# Light 3x3 weighted kernel (gentle smoothing only)
smooth_kernel <-
  matrix(
    c(
      1, 2, 1,
      2, 4, 2,
      1, 2, 1
    ),
    nrow = 3, byrow = TRUE
  )
smooth_kernel <- smooth_kernel / sum(smooth_kernel)
```

# 3) READ + COLUMN DETECTION

```{r read-all-data}
df0 <- read_csv(in_file, show_col_types = FALSE) %T>% 
  print()

# check missing columns in data file
need_base <- c("cruise_file", "station", "latitude", "longitude")
miss_base <- setdiff(need_base, names(df0))
if (length(miss_base) > 0) stop("Missing base columns: ", paste(miss_base, collapse = ", "))


# of the list of columns, pick the first one
pick_first_existing <- function(candidates, col_names) {
  hit <- candidates[candidates %in% col_names]
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

var_names <- list(
  col_sal     = c("S_0_5m", "sal_0_5m", "mean_sal_psu", "salinity"),
  col_temp    = c("T_0_5m", "temp_0_5m", "mean_temp_C", "temperature", "temp_C"),
  col_o2      = c("O2_0_5m", "oxy_0_5m", "mean_oxy", "oxygen", "DO"),
  col_date    = c("date", "Date", "cast_date", "datetime", "time", "timestamp"),
  # confirmed by you:
  col_chl     = c("CHL_0_5m"),
  col_chl_src = c("chl_source")
) %>%
  map(
    y = names(df0),
    ~ pick_first_existing(.x, .y)
  ) %>%
  unlist()

if (any(is.na(var_names))) {
  stop(
    paste0("Could not find column(s):\n- ", 
           paste(names(var_names[is.na(var_names)]), collapse = "\n- "))
       )
}

for (i in seq(var_names)) do.call("<-", list(names(var_names)[i], var_names[i]))

# rename columns
df0 <- 
  df0 %>%
  mutate(
    cruise_file = as.character(cruise_file),
    station     = str_squish(as.character(station)),
    latitude    = suppressWarnings(as.numeric(latitude)),
    longitude   = suppressWarnings(as.numeric(longitude)),
    sal         = suppressWarnings(as.numeric(.data[[col_sal]])),
    temp        = suppressWarnings(as.numeric(.data[[col_temp]])),
    oxy         = suppressWarnings(as.numeric(.data[[col_o2]])),
    chl         = suppressWarnings(as.numeric(.data[[col_chl]])),
    chl_source  = as.character(.data[[col_chl_src]]),
    date_raw    = if (!is.na(col_date)) .data[[col_date]] else NA,
    date_raw    = mdy(date_raw)
  ) %T>% 
  print()

# extract all crusie IDs
cruise_ids <- 
  df0 %>%
  pull(cruise_file) %>%
  unique() %>%
  tools::file_path_sans_ext() %T>% 
  print()

cat("Total cruises found: ", length(cruise_ids), "\n")
```

# ============================================================================ #
# 4) Functions:

```{r cruise-data}
cruise_extract <- function(cruise_id, dat) {
  
  # ---- main title text ---- #
  main_title <- paste0(cruise_id, " — Surface (0–5 m mean)")

  df <-
    dat %>%
    filter(
      str_detect(cruise_file, cruise_id), # select cruise
      !is.na(station),
      station != "",
      !is.na(latitude),
      !is.na(longitude)
    )

  if (nrow(df) == 0) stop("No rows matched cruise_id: ", cruise_id)

  # ---- subtitle text ---- #
  date_vec <- df$date_raw
  date_min <- min(date_vec, na.rm = TRUE)
  date_max <- max(date_vec, na.rm = TRUE)
  sub_title <-
    if (all(!is.na(c(date_min, date_max)))) {
      paste(format(date_min, "%Y-%m-%d"), "to", format(date_max, "%Y-%m-%d"))
    } else {
      "Date range: NA"
    }

  sub_title <- paste0(sub_title, " | IDW + (UTM grid)")

  # ---- chlorophyll source MODE for THIS cruise (panel title only) ---- #
  chl_src_tbl <- df %>%
    filter(!is.na(chl_source), chl_source != "") %>%
    count(chl_source, name = "n") %>%
    arrange(desc(n))

  chl_src_mode <- if (nrow(chl_src_tbl) == 0) {
    "unknown"
  } else {
    chl_src_tbl$chl_source[1]
  }
  
  chl_panel_title <- 
    if (identical(chl_src_mode, "chlorophyll_concentration")) {
      "Chlorophyll concentration"
      } else if (identical(chl_src_mode, "chlorophyll_fluorescence")) {
        "Chlorophyll fluorescence"
        } else {
          "Chlorophyll"
          }
  

  ## ---- station means (ONLY aggregation; no smoothing of source data)
  df_st <-
    df %>%
    summarise(
      .by = station,
      across(
        c(latitude, longitude, sal, temp, oxy, chl),
        \(.x) mean(.x, na.rm = TRUE)
      )
    ) %>%
    filter(!is.na(latitude), !is.na(longitude))
  
  # convert dataframe of stations to sf object and project
  pts_utm <- 
    st_as_sf(df_st, coords = c("longitude","latitude"), crs = 4326, remove = FALSE)  %>%
    st_transform(utm_epsg) %>%
    mutate(
      longitude = st_coordinates(.)[,1],
      latitude = st_coordinates(.)[,2]
      ) 
  
  # ---- TIGHT HARD SUPPORT MASK — computed once ---- #  
  # create a smaller, irregular bounding box based on convex hull or buffer (<3)
  if (nrow(pts_utm) >= 3) {
    # convex hull, needs at least 3 points
    support_poly_utm <- 
      pts_utm %>%
      st_union() %>%
      st_convex_hull() %>%
      st_buffer(8000)   # 8 km buffer
  } else {
    # buffer around each point
    support_poly_utm <- 
      pts_utm %>%
      st_buffer(30000) %>%
      st_union()
  }

  return(
    list(
      "pts_utm"          = pts_utm,
      "support_poly_utm" = support_poly_utm,
      "main_title"       = main_title,
      "sub_title"        = sub_title,
      "chl_panel_title"  = chl_panel_title
      )
  )
  # ---- end of function cruise_extract ---- #
}
```

# 5) FIXED TEMPLATE GRID (UTM) + IDW SETTINGS

```{r blank-grid}
grid_blank <- function(.support_poly, utm_epsg, grid_res_m = 1000) {
  # create blank raster grid with equal grid numbers
  grid_rast <- 
    # add 10 km boundary around convex hull
    st_buffer(.support_poly, dist = 10000) %>%
    st_bbox() %>%
    {
      terra::rast(
        xmin       = .[["xmin"]],
        xmax       = .[["xmax"]],
        ymin       = .[["ymin"]],
        ymax       = .[["ymax"]],
        resolution = grid_res_m,
        crs        = paste0("EPSG:", utm_epsg)
      )
    } 
  
  # convert projected raster to sf object with selected resolution
  # this will be the sf object to map predicted values from interpolation (IDW)  
  grid_sf <- 
    as.data.frame(grid_rast, xy = TRUE, cells = TRUE, na.rm = FALSE) %>%
    st_as_sf(coords = c("x","y"), crs = utm_epsg)
  
  return(
    list(
      "grid_rast" = grid_rast,
      "grid_sf"   = grid_sf
    )
  )
  # ---- end of function grid_blank ---- #
}
```




# 6) SOFT EDGE SETTINGS (opacity of points)
This only softens the DISPLAY edge near the mask boundary

alpha refers to alpha of ggplot meaning opacity

```{r edge}
edge_opacity <- function(
    .grid_rast,
    .support_poly_utm,
    .edge_feather_m = 5000 # try 4000–7000 depending on taste
    ) {
  
  # convert convex hull to linestring
  support_boundary_line <- st_boundary(support_poly_utm) %>%
    terra::vect()
  
  # calculate distance from convex hull (as linestring) then mask to within convex hull
  dist_to_edge_utm <-
    terra::distance(.grid_rast, support_boundary_line, names = "alpha") %>%
    terra::mask(terra::vect(.support_poly_utm))

  # divides distances by edge buffer then sets upper bounds to 1 by converting all values
  # above 1 to 1 and leaves all values below as is
  alpha_utm <-
    terra::clamp(dist_to_edge_utm / .edge_feather_m, lower = 0, upper = 1, values = TRUE)

  # ---- end of function edge_opacity ---- #
}
```


# 7) PANEL MAKER


```{r idw-panel}
make_panel <- function(
  col_name, 
  var_title, 
  legend_units, 
  .legend_limits    = NULL,
  .df_st            = pts_utm,  
  .alpha_utm        = alpha_utm,
  .smooth_kernel    = smooth_kernel,
  .grid_sf          = grid_sf,
  .grid_rast        = grid_rast,
  .utm_epsg         = utm_epsg,
  .support_poly_utm = support_poly_utm,
  .florida_sf       = florida_sf,
  .bb_utm           = bb_utm,
  idw_opts          = idw_options
  ) {

  # keep columns needed for IDW interpolation and plotting, converts selected
  # parameter name to val, removes NA 
  dd <- 
    .df_st %>%
    mutate(
      .keep = "none",
      station,
      latitude,
      longitude,
      val = .data[[col_name]],
      geometry
    ) %>%
    filter(!is.na(val))
  
  # ---- skip plotting if less than 5 points ---- #
  if (nrow(dd) < 5) {
    return(
      ggplot() +
        theme_void() +
        labs(title = paste0(var_title, " (too few stations)")) +
        theme(plot.title = element_text(face = "bold", size = 12))
    )
  }

  # ---- IDW interpolation as sf object ---- #
  # uses options set before
  idw_sf_utm <- 
    gstat::idw(
      formula   = val ~ 1,
      locations = dd,
      newdata   = .grid_sf,
      idp       = idw_opts$idp_pow,
      maxdist   = idw_opts$maxdist_m,
      nmin      = idw_opts$nmin_nb,
      nmax      = idw_opts$nmax_nb
    )

  # replace empty grid with IDW values
  terra::values(.grid_rast) <- as.numeric(idw_sf_utm$var1.pred)


  # ---- LIGHT DISPLAY SMOOTHING ONLY ---- #
  r_pred_water <-
    # mask data outside convex hull
    terra::mask(.grid_rast, terra::vect(.support_poly_utm)) %>%
    
    # smooths over values with weighting of 3x3 kernel
    terra::focal(
      w         = .smooth_kernel,
      fun       = mean,
      na.policy = "omit",
      na.rm     = TRUE,
      fillvalue = NA
    ) %>%
    
    # combine IDW and alpha (opacity)
    c(., alpha_utm) %>%
    
    # mask again so smoothing does not bleed outside supported footprint
    # of convex hull
    terra::mask(terra::vect(.support_poly_utm)) %>% # <-- may not be necessary
    
    # mask land mass
    terra::mask(terra::vect(.florida_sf), inverse = TRUE) 
    
  # convert into data frame
  plot_df <- 
    r_pred_water %>%
    as.data.frame(xy = TRUE, na.rm = FALSE) %>%
    rename("pred" = focal_mean)  %>%
    mutate(alpha = ifelse(is.na(alpha), 0, alpha)) %>%
    filter(!is.na(pred), alpha > 0)
  
  
  # ---- contours ---- #
  contour_sf <- NULL

  rng <- range(plot_df$pred, na.rm = TRUE)

  if (all(is.finite(rng)) && diff(rng) > 0) {
    brks <- pretty(rng, n = 6)
    brks <- brks[brks > rng[1] & brks < rng[2]]
    
    if (length(brks) > 0) {
      contour_sf <- try(
        terra::as.contour(r_pred_water, levels = brks),
        silent = TRUE
      )
      # if error, convert to sf
      if (!inherits(contour_sf, "try-error")) {
        contour_sf <- try(sf::st_as_sf(contour_sf), silent = TRUE)
      }
      
      # if error on conversion, set to NULL
      if (inherits(contour_sf, "try-error")) contour_sf <- NULL
    }
  }

  if (is.null(.legend_limits)) .legend_limits <- c(NA, NA)
  
  # ---- plot ---- #
  p <- 
    ggplot() +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "#cfe8f3") +
    geom_raster(data = plot_df, aes(x = x, y = y, fill = pred, alpha = alpha)) +
    scale_fill_viridis_c(
      name = legend_units, option = "C", direction = 1, na.value = NA, 
      limits = .legend_limits, oob = scales::squish) +
    scale_alpha_identity() +
    geom_sf(data = .florida_sf, fill = "#e8e6df", color = "grey35", linewidth = 0.55) +

  # add contours if possible
  {
    if (!is.null(contour_sf) && nrow(contour_sf) > 0) {
      geom_sf(
        data      = contour_sf,
        color     = "white",
        linewidth = 0.22,
        alpha     = 0.45
      )
      }
    } +
    geom_sf(data = dd, color = "black", size = 0.55, alpha = 0.95) +
    ggrepel::geom_text_repel(
      data = dd,
      aes(x = longitude, y = latitude, label = station),
      seed               = 1,
      size               = 2.0,
      color              = "black",
      box.padding        = 0.18,
      point.padding      = 0.10,
      min.segment.length = 0,
      segment.color      = "grey35",
      segment.size       = 0.30,
      segment.curvature  = 0.0,
      arrow              = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
      max.overlaps       = Inf
    ) +
    annotation_north_arrow(
      location    = "tr",
      which_north = "true",
      pad_x       = unit(0.10, "cm"),
      pad_y       = unit(0.10, "cm"),
      height      = unit(0.85, "cm"),
      width       = unit(0.85, "cm"),
      style       = north_arrow_orienteering(
        line_col = "grey20",
        fill     = c("grey20", "white"),
        text_col = "grey20"
      )
    ) +
    annotation_scale(
      location   = "bl",
      width_hint = 0.22,
      pad_x      = unit(0.12, "cm"),
      pad_y      = unit(0.12, "cm"),
      text_cex   = 0.55,
      line_width = 0.6
    ) +
    coord_sf(xlim = .bb_utm[c(1,3)], ylim = .bb_utm[c(2,4)], expand = FALSE) +
    labs(title = var_title, x = NULL, y = NULL) +
    theme_bw() +
    theme(
      panel.background = element_rect(fill = "#cfe8f3"),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      legend.position  = "right",
      plot.title       = element_text(face = "bold", size = 12)
    )

  p
}
```

# 8) SAVE (SHORT FILENAMES to avoid OneDrive truncation)
```{r save-func}
save_maps <- function(plt, .map_dir, .cruise_id, verbose = TRUE) {
  
  png_out <- here(.map_dir, "png")
  svg_out <- here(.map_dir, "svg")
  
  dir_create(png_out)
  dir_create(svg_out)
  
  # file names
  png_out  <- here(png_out, paste0(.cruise_id, "_0_5m_4vars_softedge_nav.png"))
  tiff_out <- here(svg_out, paste0(.cruise_id, "_0_5m_4vars_softedge_nav.tiff"))

  # save
  ggsave(png_out, plot = plt, width = 12.5, height = 9.5, dpi = 600)
  ggsave(tiff_out, plot = plt, width = 12.5, height = 9.5, dpi = 600, compression = "lzw")

  if (!verbose) return(invisible())
  cat("\nOutput folder:\n", .map_dir, "\n\n")
  cat("\nSaved PNG:\n", basename(png_out), "\nexists =", file_exists(png_out), "\n")
  cat("Saved TIFF:\n", basename(tiff_out), "\nexists =", file_exists(tiff_out), "\n")
  print(list.files(.map_dir))
  print(basename(dir_ls(.map_dir)))
  # ---- end of function save_maps ---- #
}


```


# 9) Cruise Processing

```{r process}
legends_lim <- list(
  sal_limits  = c(29, 37),
  temp_limits = c(18, 32),
  oxy_limits  = c(1, 5),
  chl_limits  = c(0, 5)
)

# for (cruise_id in cruise_ids) {
for (cruise_id in cruise_ids[-c(1:48)]) {
  cat(paste("Cruise ID:", cruise_id))

  # extract cruise data
  cruise_temp      <- cruise_extract(cruise_id, dat = df0)
  pts_utm          <- cruise_temp$pts_utm
  support_poly_utm <- cruise_temp$support_poly_utm
  main_title       <- cruise_temp$main_title
  sub_title        <- cruise_temp$sub_title
  chl_panel_title  <- cruise_temp$chl_panel_title

  # create blank grid
  grid_temp <- grid_blank(support_poly_utm, utm_epsg)
  grid_rast <- grid_temp$grid_rast
  grid_sf   <- grid_temp$grid_sf

  # # show locations, convex hull in relation to Florida
  # print(ggplot() +
  #   geom_sf(data = grid_sf, color = "#500", alpha = 0.5) + # base grid for interpolation
  #   geom_sf(data = florida_sf) + # florida map
  #   geom_sf(
  #     data = support_poly_utm,
  #     color = "blue", fill = "lightblue", alpha = 0.5
  #   ) + # convex hull
  #   geom_sf(data = pts_utm) + # data points used
  #   coord_sf(expand = FALSE) +
  #   labs(x = NULL, y = NULL) +
  #   ggrepel::geom_text_repel(
  #     data = pts_utm,
  #     aes(x = longitude, y = latitude, label = station),
  #     seed               = 1,
  #     size               = 2.0,
  #     color              = "black",
  #     box.padding        = 0.18,
  #     point.padding      = 0.10,
  #     min.segment.length = 0,
  #     segment.color      = "grey35",
  #     segment.size       = 0.30,
  #     segment.curvature  = 0.0,
  #     arrow              = grid::arrow(length = unit(0.006, "npc"), type = "closed"),
  #     max.overlaps       = Inf
  #   ))

  # create edge opacity
  alpha_utm <- edge_opacity(grid_rast, support_poly_utm)

 # print(ggplot() +
 #    tidyterra::geom_spatraster(data = alpha_utm) +
 #    scale_fill_viridis())
  
  # make panels
  p_sal <- make_panel("sal", "Salinity", "PSU", legends_lim[[1]])
  p_tmp <- make_panel("temp", "Temperature", expression(paste(degree, "C")), legends_lim[[2]])
  p_o2  <- make_panel("oxy", "Oxygen", expression(paste("mL ", L^{-1})), legends_lim[[3]])
  p_chl <- make_panel("chl", chl_panel_title, expression(paste("mg ", m^{-3})), legends_lim[[4]])

  # combine panels
  final_plot <-
    (p_sal + p_tmp) / (p_o2 + p_chl) +
    plot_annotation(
      title = main_title,
      subtitle = sub_title
    ) +
    plot_annotation(
      theme = theme(
        plot.title    = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 11)
      )
    )

  print(final_plot)

  save_maps(final_plot, map_dir, cruise_id)
}
```




# FOCUS RANGES (pattern emphasis)
```{r}
sal_limits  <- c(28, 35)
oxy_limits  <- c(1,  5)
temp_limits <- c(18, 30)
chl_limits  <- c(0,  5)

cat("\nFOCUS color limits:\n")
cat("Salinity:    ", sal_limits[1],  " to ", sal_limits[2],  "\n", sep = "")
cat("Oxygen:      ", oxy_limits[1],  " to ", oxy_limits[2],  "\n", sep = "")
cat("Temperature: ", temp_limits[1], " to ", temp_limits[2], "\n", sep = "")
cat("Chlorophyll: ", chl_limits[1],  " to ", chl_limits[2],  "\n\n", sep = "")
```

