---
title: "Unstructured Grid Example"
author: "Sebastian Di Geronimo"
date: 2026-09-01
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

This download, loads, crops and plots an unstructured grid data example from 
`psyplot` package in python, converted for R. 

# 2.0 ---- Setup ----


# 3.0 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
  install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, purrr, dplyr, glue, fs, magrittr, here,
  
  # additional
  ncdf4,
  sf,
  downloader
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
```

# 4.0 Load Dataset

```{r load-data}
# path to ugrid netcdf file
nc_file <- here::here("data", "raw","ugrid_demo1.nc")

# download unstructured grid file example
if (!fs::file_exists(nc_file)) {
  downloader::download(
    url = "https://github.com/psyplot/examples/blob/main/maps/ugrid_demo.nc",
    destfile = here::here("data", "raw", "ugrid_demo1.nc")
  )
}
  
# open netcdf 
nc <- nc_open(nc_file)
```

## 4.1 Insepct Variables

node names: 
* Mesh2_node_x:     Longitude of 2D mesh nodes in degrees_east
* Mesh2_node_y:     Latitude of 2D mesh nodes in degrees_north
* Mesh2_face_nodes: Maps every triangular face to its three corner nodes
* Mesh2_edge_nodes: Maps every edge to the two nodes it connects

vars:
* Mesh2_height: fluid height in m
* Mesh2_bathy:  bathymetry in m
* Mesh2_m_x:    momentum in x-direction in m^2/s
* Mesh2_m_y:    momentum in y-direction in m^2/s
* Mesh2_u_x:    velocity in x-direction in m/s
* Mesh2_u_y:    velocity in y-direction in m/s
* Mesh2_level:  grid level in None

```{r}
print(nc$var)  # See available variables
print(nc$dim)  # See dimensions
tidync::tidync(nc_file)
```

## 4.2 Extract Variables

need coordinates, faces and parameter

```{r}
# coordinates
node_lon  <- ncvar_get(nc, "Mesh2_node_x")
node_lat <- ncvar_get(nc, "Mesh2_node_y")

# face nodes
face_nodes <- ncvar_get(nc, "Mesh2_face_nodes")

# check face_nodes has one row per face and three columns
if (nrow(face_nodes) == 3 && ncol(face_nodes) != 3) {
  face_nodes <- t(face_nodes)
}

# UGRID file uses start_index = 0
face_nodes <- face_nodes + 1

# variable values per face
# height     <- ncvar_get(nc, "Mesh2_height")
height     <- ncvar_get(nc, "Mesh2_bathy")

# extract parameter as a vector
height_face <-
  if (length(dim(height)) == 2) {
    as.vector(height[1, ])
  } else {
    as.vector(height)
  }
```

# 5.0 Spatial Filter 

longitude: 0 - 360 -> need to subtract 360 from a negative longitude
latidude: -90 - 90

```{r}
# Gulf of Mexico filter
nodes_in_box <- 
  cbind(
    between(node_lon, 360 - 98, 360 - 80),
    between(node_lat, 18, 31)
  )

node_ids <- which(rowSums(nodes_in_box) == 2)

face_keep <- apply(
  face_nodes,
  1,
  function(ids) any(ids %in% node_ids, na.rm = TRUE)
)
face_ids <- which(face_keep)

# Remove invalid or missing connectivity values
face_nodes[face_nodes < 1] <- NA

# filter face node and parameter within spatial filter
height_sub <- height_face[face_ids]

face_nodes_sub <- face_nodes[face_ids, , drop = FALSE]
```


# 6.0 Convert Each Node Face to sf Object

```{r}
mesh_sf <- map_dfr(
  seq_along(face_ids),
  function(i) {
    # subset to first row
    ids <- face_nodes_sub[i, ]

    # remove NAs
    if (anyNA(ids)) {
      return(NULL)
    }

    # select 3 coordinate nodes
    coords <- cbind(
      node_lon[ids],
      node_lat[ids]
    )

    # close the triangle by adding first row to end
    coords <- rbind(coords, coords[1, ])

    # convert to sf polygon object
    st_sf(
      face = i,
      height = height_sub[i],
      geometry = st_sfc(
        st_polygon(list(coords)),
        crs = 4326
      )
    )
  }
)

# shift coordinates to -180 to 180 
mesh_sf <- 
  mesh_sf %>%
  st_transform(4326) %>%
  st_wrap_dateline(options = c("WRAPDATELINE=YES"))
```

# 7.0 Plot Data

## 7.1 Extract Coastline

US, Cubam Bahamas and Mexico

```{r maps}
shelf(maps)

gulf_coast_line <- map_data(
  "world",
  region = c("USA", "Cuba", "Bahamas", "Mexico")
)
```


## 7.2 Plot Data

```{r}
ggplot() +
  geom_sf(
    data = mesh_sf,
    aes(fill = height),
    color     = "grey30",
    linewidth = 0.05
  ) +
  geom_polygon(
    data = gulf_coast_line,
    aes(x = long, y = lat, group = group),
    fill      = "grey80",
    color     = "grey30",
    linewidth = 0.2,
    alpha     = 0.5
  ) +
  scale_fill_viridis_c(name = "Bathymetry") +
  coord_sf(
    xlim = c(-100, -78.0),
    ylim = c(17, 33),
    expand = FALSE
    ) +
  theme_minimal()
```

## 7.3 Further Crop to Make More Square
```{r}
keys_bbox <- st_bbox(
  c(
    xmin = -98,
    xmax = -80,
    ymin = 18,
    ymax = 31
  ),
  crs = st_crs(4326)
)

keys_box <- st_as_sfc(keys_bbox)

mesh_keys <- st_crop(mesh_sf, keys_bbox)
# plot(mesh_keys[,-1])

ggplot(mesh_keys) +
  geom_sf(
    aes(fill = height),
    color = "grey30",
    linewidth = 0.05
  ) +
  geom_polygon(
    data = gulf_coast_line,
    aes(x = long, y = lat, group = group),
    fill = "grey80",
    color = "grey30",
    linewidth = 0.2,
    alpha = 0.5
  ) +
  scale_fill_viridis_c(name = "Bathymetry") +
  coord_sf(
    xlim = c(-98, -80.0),
    ylim = c(18, 31),
    expand = FALSE) +
  theme_minimal()

```



